---
title: "Transforming Covariates"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Transforming Covariates}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

*Built using Zelig version `r packageVersion('Zelig')`*

```{r setup, include=FALSE}
knitr::opts_knit$set(
        stop_on_error = 2L
)
```

> NOTE: this vignette is under construction.

For convienience, you may wish to transform one or more of your covariates within the `zelig` model estimation call. For example, imagine we want to natural log transform the variable `dist` in a linear regression:

```{r}
library(Zelig)
library(dplyr)

z.log <- zelig(speed ~ log(dist), data = cars, model = 'ls', cite = FALSE)
```

Then we want to find quantities of interest when the `dist` variable is 26 to
56 (the first to third quartile). To do this we added the logged fitted values 
to `setx`, simulate quantites of interest, and plot the results:

```{r}
setx(z.log, dist = log(26:56)) %>%
    sim() %>%
    ci.plot(xlab = 'dist (log)') 
```


