---
title: "Gamma Survey"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{zelig-gammasurvey}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

*Built using Zelig version `r packageVersion("Zelig")`*

```{r setup, include=FALSE}
knitr::opts_knit$set(
    stop_on_error = 2L
)
knitr::opts_chunk$set(
    fig.height = 11,
    fig.width = 7
)

options(cite = FALSE)
```
---



Gamma Regression for Continuous, Positive Dependent Variables with `gamma.survey`.

Use the gamma regression model if you have a positive-valued dependent
variable such as the number of years a parliamentary cabinet endures, or
the seconds you can stay airborne while jumping. The gamma distribution
assumes that all waiting times are complete by the end of the study
(censoring is not allowed).

Syntax
------

With reference classes:


With the Zelig 4 compatibility wrappers:

```{r, eval = FALSE}
z.out <- zelig(Y ~ X1 + X2, model = "gamma.survey", weights = w, data = mydata)
x.out <- setx(z.out)
s.out <- sim(z.out, x = x.out, x1 = NULL)
```

Examples
------


```{r, eval = TRUE, echo = FALSE}
rm(list=ls(pattern="\\.out"))
suppressWarnings(suppressMessages(library(Zelig)))
set.seed(1234)
```

### Example 1: User has Existing Sample Weights

Attach sample data and variable names:

```{r, eval = TRUE}
data(api, package="survey")
```

In this example, we will estimate a model using
the percentages of students who receive subsidized
lunch and an indicator for whether schooling is
year-round to predict California public schools'
academic performance index scores:

```{r, eval = TRUE}
z.out1 <- zelig(api00 ~ meals + yr.rnd, model = "gamma.survey",
                weights = ~pw, data = apistrat)
summary(z.out1)
```

Set explanatory variables to their default (mean/mode) values, and set
a high (80th percentile) and low (20th percentile) value for "meals,"
the percentage of students who receive subsidized meals:

```{r, eval = TRUE}
x.low <- setx(z.out1, meals= quantile(apistrat$meals, 0.2))
x.high <- setx(z.out1, meals= quantile(apistrat$meals, 0.8))
```

Generate first differences for the effect of high versus low "meals"
on academic performance:

```{r, eval = TRUE}
s.out1 <- sim(z.out1, x=x.high, x1=x.low)
summary(s.out1)
```

Generate a second set of fitted values and a plot:

```{r Zelig-gammasurvey, dev=c("png", "pdf"), eval = TRUE, fig.cap = "Graphs of Quantities of Interest for Gamma Survey"}
plot(s.out1)
```

### Example 2: User has Details about Complex Survey Design (but not sample weights)

Suppose that the survey house that provided
the dataset excluded probability weights
but made other details about the survey
design available.  We can still estimate
a model without probability weights that takes
instead variables that identify each the stratum
and/or cluster from which each observation was
selected and the size of the finite sample from
which each observation was selected.

```{r, eval = TRUE}
z.out2 <- zelig(api00 ~ meals + yr.rnd, model = "gamma.survey",
                strata=~stype, fpc=~fpc, data = apistrat)
summary(z.out2)
```

Note that these results are identical to the results obtained
when pre-existing sampling weights were used.  When sampling
weights are omitted, Zelig estimates them automatically for
"gamma.survey" models based on the user-defined description
of sampling designs.  If no description is present, the default
assumption is equal probability sampling.

The methods `setx()`` and `sim()` can then be run on `z.out2` in the same fashion
described in Example 1.

### Example 3: User has Replicate Weights

Suppose that the survey house that published
these data withheld details about the survey
design and instead published replication weights.

For the purpose of illustration, create a set of
jk1 replicate weights:

```{r, eval = TRUE}
library("survey")
jk1reps <- jk1weights(psu=apistrat$dnum)
```

Estimate the model regressing api00 on the "meals"
"yr.rnd" variables.

```{r, eval = TRUE}
z.out3 <- zelig(api00 ~ meals + yr.rnd, model = "gamma.survey",
      data = apistrat, repweights=jk1reps$weights,
      type="JK1")
summary(z.out3)
```

Set the explanatory variable "meals" at high and low values

```{r, eval = TRUE}
x.low <- setx(z.out3, meals= quantile(apistrat$meals, 0.2))
x.high <- setx(z.out3, meals= quantile(apistrat$meals, 0.8))
```

Generate first differences for the effect of the high
versus low concentrations of poverty on school performance

```{r, eval = TRUE}
s.out3 <- sim(z.out3, x=x.high, x1=x.low)
summary(s.out3)
```

Generate a second set of fitted values and a plot:

```{r, eval = TRUE}
plot(s.out3)
```

The user should also refer to the gamma model demo, since
gamma.survey models can take many of the same options as
gamma models.


Model
------

-  The Gamma distribution with scale parameter $\alpha$ has a
   *stochastic component*:


$$
    \begin{aligned}
        Y &\sim& \textrm{Gamma}(y_i \mid \lambda_i, \alpha) \\
      f(y)  &=& \frac{1}{\alpha^{\lambda_i} \, \Gamma \lambda_i} \, y_i^{\lambda_i
        - 1} \exp -\left\{ \frac{y_i}{\alpha} \right\}\end{aligned}

   | for $\alpha, \lambda_i, y_i > 0$.

-  The *systematic component* is given by


$$
    \lambda_i = \frac{1}{x_i \beta}
$$

Quantities of Interest
------

-  The expected values (qi$ev) are simulations of the mean of the
   stochastic component given draws of $\alpha$ and $\beta$
   from their posteriors:


$$
    E(Y) = \alpha \lambda_i.
$$

-  The predicted values (qi$pr) are draws from the gamma distribution
   for each given set of parameters $(\alpha, \lambda_i)$.

-  If x1 is specified, sim() also returns the differences in the
   expected values (qi$fd),


$$
    E(Y \mid x_1) -
        E(Y \mid x)
$$

   .

-  In conditional prediction models, the average expected treatment
   effect (att.ev) for the treatment group is


$$
    \frac{1}{\sum_{i=1}^n t_i}\sum_{i:t_i=1}^n \left\{ Y_i(t_i=1) -
        E[Y_i(t_i=0)] \right\},
$$

   where $t_i$ is a binary explanatory variable defining the
   treatment ($t_i=1$) and control ($t_i=0$) groups.
   Variation in the simulations are due to uncertainty in simulating
   $E[Y_i(t_i=0)]$, the counterfactual expected value of
   $Y_i$ for observations in the treatment group, under the
   assumption that everything stays the same except that the treatment
   indicator is switched to $t_i=0$.

-  In conditional prediction models, the average predicted treatment
   effect (att.pr) for the treatment group is


$$
    \frac{1}{\sum_{i=1}^n t_i}\sum_{i:t_i=1}^n \left\{ Y_i(t_i=1) -
        \widehat{Y_i(t_i=0)} \right\},
$$

   where $t_i$ is a binary explanatory variable defining the
   treatment ($t_i=1$) and control ($t_i=0$) groups.
   Variation in the simulations are due to uncertainty in simulating
   $\widehat{Y_i(t_i=0)}$, the counterfactual predicted value of
   $Y_i$ for observations in the treatment group, under the
   assumption that everything stays the same except that the treatment
   indicator is switched to $t_i=0$.

Output Values
------

The Zelig object stores fields containing everything needed to
rerun the Zelig output, and all the results and simulations as they are generated.
In addition to the summary commands demonstrated above, some simply utility
functions (known as *getters*) provide easy access to the raw fields most
commonly of use for further investigation.

In the example above `z.out$get_coef()` returns the estimated coefficients, `z.out$get_vcov()` returns the estimated covariance matrix, and `z.out$get_predict()` provides predicted values for all observations in the dataset from the analysis.

See also
------

The gammasurvey model is part of the survey package by Thomas Lumley, which in turn depends heavily on `glm` package. Advanced users may
wish to refer to `help(svyglm)` and `help(family)`.

```{r, eval = TRUE, echo=FALSE, results = "asis"}
z5 <- zgammasurvey$new()
z5$references()
```
