---
title: "Zelig-lognorm"
output:
  html_document:
    number_sections: yes
    toc: yes
bibliography: zelig5.bib
---

<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Model: Zelig-lognorm}
-->

# Log-Normal Regression for Duration Dependent Variables {#lognorm}

The log-normal model describes an event’s duration, the dependent
variable, as a function of a set of explanatory variables. The
log-normal model may take time censored dependent variables, and allows
the hazard rate to increase and decrease.

## Syntax

```{r, eval = FALSE}
z.out <- zelig(Surv(Y, C) ~ X, model = "lognorm", data = mydata)
x.out <- setx(z.out)
s.out <- sim(z.out, x = x.out)
```

Log-normal models require that the dependent variable be in the form
`Surv(Y, C)`, where `Y` and `C` are
vectors of length $n$. For each observation $i$ in 1, …, $n$, the value
$y_i$ is the duration (lifetime, for example) of each subject, and the
associated $c_i$ is a binary variable such that $c_i = 1$ if the
duration is not censored (`*e.g.*`, the subject dies during
the study) or $c_i = 0$ if the duration is censored
(`*e.g.*`, the subject is still alive at the end of the
study). If $c_i$ is omitted, all Y are assumed to be completed; that is,
time defaults to 1 for all observations.

## Input Values

In addition to the standard inputs, `zelig()` takes the
following additional options for lognormal regression:

-   `robust`: defaults to `FALSE`. If
    `TRUE`, `zelig()` computes robust standard
    errors based on sandwich estimators (see @Huber81 and @White80)
    based on the options in `cluster`.

-   `cluster`: if `robust = TRUE`, you may select
    a variable to define groups of correlated observations. Let
    `x3` be a variable that consists of either discrete
    numeric values, character strings, or factors that define strata.
    Then
```{r, eval=FALSE}
z.out <- zelig(y ~ x1 + x2, robust = TRUE, cluster = "x3", 
               model = "exp", data = mydata)
```

    means that the observations can be correlated within the strata
    defined by the variable `x3`, and that robust standard
    errors should be calculated according to those clusters. If
    `robust = TRUE` but `cluster` is not
    specified, `zelig()` assumes that each observation falls
    into its own cluster.

## Example

Attach the sample data:

```{r echo = FALSE, warning = FALSE, message = FALSE}
suppressPackageStartupMessages(library(Zelig5, quietly = TRUE))
data(coalition)
```

```{r}
data(coalition)
```

Estimate the model:

```{r}
z.out <-zelig(Surv(duration, ciep12) ~ fract + numst2, model = "lognorm",
              data = coalition)
```

View the regression output:

```{r}
summary(z.out)
```

Set the baseline values (with the ruling coalition in the minority) and
the alternative values (with the ruling coalition in the majority) for
X:

```{r}
x.low <- setx(z.out, numst2 = 0)
x.high <- setx(z.out, numst2 = 1)
```


Simulate expected values (`qi$ev`) and first differences
(`qi$fd`):

```{r}
s.out <- sim(z.out, x = x.low, x1 = x.high)
summary(s.out)
# plot(s.out)
```

## Model

Let $Y_i^*$ be the survival time for observation $i$ with the density
function $f(y)$ and the corresponding distribution function
$F(t)=\int_{0}^t f(y) dy$. This variable might be censored for some
observations at a fixed time $y_c$ such that the fully observed
dependent variable, $Y_i$, is defined as

$$Y_i = \left\{ \begin{array}{ll}
      Y_i^* & \textrm{if }Y_i^* \leq y_c \\
      y_c & \textrm{if }Y_i^* > y_c \\
    \end{array} \right.$$

-   The *stochastic component* is described by the distribution of the
    partially observed variable, $Y^*$. For the lognormal model, there
    are two equivalent representations:

    $$\begin{aligned}
        Y_i^* \; \sim \; \textrm{LogNormal}(\mu_i, \sigma^2) & \textrm{ or
    } & \log(Y_i^*) \; \sim \; \textrm{Normal}(\mu_i, \sigma^2)\end{aligned}$$

    where the parameters $\mu_i$ and $\sigma^2$ are the mean and
    variance of the Normal distribution. (Note that the output from
    `zelig()` parameterizes `scale`$ = \sigma$.)

    In addition, survival models like the lognormal have three
    additional properties. The hazard function $h(t)$ measures the
    probability of not surviving past time $t$ given survival up to $t$.
    In general, the hazard function is equal to $f(t)/S(t)$ where the
    survival function $S(t) =
      1 - \int_{0}^t f(s) ds$ represents the fraction still surviving at
    time $t$. The cumulative hazard function $H(t)$ describes the
    probability of dying before time $t$. In general, $H(t)=
    \int_{0}^{t} h(s) ds = -\log S(t)$. In the case of the lognormal
    model,

    $$\begin{aligned}
    h(t) &=& \frac{1}{\sqrt{2 \pi} \, \sigma t \, S(t)}
    \exp\left\{-\frac{1}{2 \sigma^2} (\log \lambda t)^2\right\} \\
    S(t) &=& 1 - \Phi\left(\frac{1}{\sigma} \log \lambda t\right) \\
    H(t) &=& -\log \left\{ 1 - \Phi\left(\frac{1}{\sigma} \log \lambda t\right) \right\}\end{aligned}$$

    where $\Phi(\cdot)$ is the cumulative density function for the
    Normal distribution.

-   The *systematic component* is described as:

    $$\mu_i = x_i \beta .$$

## Quantities of Interest

-   The expected values (`qi$ev`) for the lognormal model
    are simulations of the expected duration:

    $$E(Y) =  \exp\left(\mu_i + \frac{1}{2}\sigma^2 \right),$$

    given draws of $\beta$ and $\sigma$ from their sampling
    distributions.

-   The predicted value is a draw from the log-normal distribution given
    simulations of the parameters $(\lambda_i, \sigma)$.

-   The first difference (`qi$fd`) is

    $$\textrm{FD} = E(Y \mid x_1) - E(Y \mid x).$$

-   In conditional prediction models, the average expected treatment
    effect (`att.ev`) for the treatment group is

    $$\frac{1}{\sum_{i=1}^n t_i}\sum_{i:t_i=1}^n \{ Y_i(t_i=1) - E[Y_i(t_i=0)] \},$$

    where $t_i$ is a binary explanatory variable defining the treatment
    ($t_i=1$) and control ($t_i=0$) groups. When $Y_i(t_i=1)$ is
    censored rather than observed, we replace it with a simulation from
    the model given available knowledge of the censoring process.
    Variation in the simulations is due to two factors: uncertainty in
    the imputation process for censored $y_i^*$ and uncertainty in
    simulating $E[Y_i(t_i=0)]$, the counterfactual expected value of
    $Y_i$ for observations in the treatment group, under the assumption
    that everything stays the same except that the treatment indicator
    is switched to $t_i=0$.

-   In conditional prediction models, the average predicted treatment
    effect (`att.pr`) for the treatment group is

    $$\frac{1}{\sum_{i=1}^n t_i} \sum_{i:t_i=1}^n \{  Y_i(t_i=1) -
    \widehat{Y_i(t_i=0)} \},$$

    where $t_i$ is a binary explanatory variable defining the treatment
    ($t_i=1$) and control ($t_i=0$) groups. When $Y_i(t_i=1)$ is
    censored rather than observed, we replace it with a simulation from
    the model given available knowledge of the censoring process.
    Variation in the simulations are due to two factors: uncertainty in
    the imputation process for censored $y_i^*$ and uncertainty in
    simulating $\widehat{Y_i(t_i=0)}$, the counterfactual predicted
    value of $Y_i$ for observations in the treatment group, under the
    assumption that everything stays the same except that the treatment
    indicator is switched to $t_i=0$.

## Output Values

The output of each Zelig command contains useful information which you
may view. For example, if you run
`z.out <- zelig(Surv(Y, C) ~ X, model = lognorm, data)`, then you may
examine the available information in `z.out` by using `names(z.out)`,
see the `coefficients` by using
`z.out$coefficients`, and a default summary of information
through `summary(z.out)`. Other elements available through the
`$` operator are listed below.

-   From the `zelig()` output object `z.out`, you
    may extract:

    -   `coefficients`: parameter estimates for the
        explanatory variables.

    -   `icoef`: parameter estimates for the intercept and
        $\sigma$.

    -   `var`: Variance-covariance matrix.

    -   `loglik`: Vector containing the log-likelihood for
        the model and intercept only (respectively).

    -   `linear.predictors`: the vector of $x_{i}\beta$.

    -   `df.residual`: the residual degrees of freedom.

    -   `df.null`: the residual degrees of freedom for the
        null model.

    -   `zelig.data`: the input data frame if `save.data
        = TRUE`.

-   Most of this may be conveniently summarized using
    `summary(z.out)`. From `summary(z.out)`, you
    may additionally extract:

    -   `table`: the parameter estimates with their
        associated standard errors, $p$-values, and $t$-statistics.

-   From the `sim()` output object `s.out`, you
    may extract quantities of interest arranged as matrices indexed by
    simulation $\times$ `x`-observation (for more than one
    `x`-observation). Available quantities are:

    -   `qi$ev`: the simulated expected values for the
        specified values of `x`.

    -   `qi$pr`: the simulated predicted values drawn from
        the distribution defined by $(\lambda_i, \sigma)$.

    -   `qi$fd`: the simulated first differences between the
        simulated expected values for `x` and
        `x1`.

    -   `qi$att.ev`: the simulated average expected
        treatment effect for the treated from conditional prediction
        models.

    -   `qi$att.pr`: the simulated average predicted
        treatment effect for the treated from conditional prediction
        models.

How to Cite {#how-to-cite .unnumbered}
-----------

See also {#see-also .unnumbered}
--------

The exponential function is part of the survival library by by Terry
Therneau, ported to R by Thomas Lumley. Advanced users may wish to refer
to `help(survfit)` in the survival library, and @VenRip02. Sample data
are from @KinAltBur90.
