---
title: "Probit Regression for Dependent Variables with Survey Weights"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{zelig-probitsurvey}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

*Built using Zelig version `r packageVersion("Zelig")`*

```{r setup, include=FALSE}
knitr::opts_knit$set(
    stop_on_error = 2L
)
knitr::opts_chunk$set(
    fig.height = 11,
    fig.width = 7
)

options(cite = FALSE)
```
---

Probit Regression for Dichotomous Dependent Variables with Survey Weights with `probit.survey`.

Use probit regression to model binary dependent variables specified as a
function of a set of explanatory variables.

Syntax
------

With the Zelig 4 compatibility wrappers:

```{r, eval = FALSE}
z.out <- zelig(Y ~ X1 + X2, model = "probit.survey", weights = w,
               data = mydata)
x.out <- setx(z.out)
s.out <- sim(z.out, x = x.out, x1 = NULL)
```

Examples
------

```{r, eval = TRUE, echo = FALSE}
rm(list=ls(pattern="\\.out"))
suppressWarnings(suppressMessages(library(Zelig)))
set.seed(1234)
```

### Example 1: User has Existing Sample Weights

Our example dataset comes from the survey package:

```{r, eval = TRUE}
data(api, package = "survey")
```

In this example, we will estimate a model using
the percentages of students who receive subsidized
lunch and the percentage who are new to a school
to predict whether each California public school
attends classes year round.  We first make a numeric
version of the variable in the example dataset, which
you may not need to do in another dataset.

```{r, eval = TRUE}
apistrat$yr.rnd.numeric <- as.numeric(apistrat$yr.rnd=="Yes")
z.out1 <- zelig(yr.rnd.numeric ~ meals + mobility,
                model = "probit.survey", weights = apistrat$pw,
                data = apistrat)
summary(z.out1)
```

Set explanatory variables to their default (mean/mode) values, and set
a high (80th percentile) and low (20th percentile) value for "meals,"
the percentage of students who receive subsidized meals:

```{r, eval = TRUE}
x.low <- setx(z.out1, meals = quantile(apistrat$meals, 0.2))
x.high <- setx(z.out1, meals = quantile(apistrat$meals, 0.8))
```

Generate first differences for the effect of high versus low "meals"
on the probability that a school will hold classes year round:

```{r, eval = TRUE}
s.out1 <- sim(z.out1, x=x.low, x1=x.high)
summary(s.out1)
```

Generate a second set of fitted values and a plot:

```{r Zelig-probitsurvey1, dev=c("png", "pdf"), eval = TRUE, fig.cap = "Graphs of Quantities of Interest for Probit Survey Model"}
plot(s.out1)
```

### Example 2: User has Details about Complex Survey Design (but not sample weights)

Suppose that the survey house that provided
the dataset excluded probability weights
but made other details about the survey
design available.  We can still estimate
a model without probability weights that takes
instead variables that identify each the stratum
and/or cluster from which each observation was
selected and the size of the finite sample from
which each observation was selected.

```{r, eval = TRUE}
z.out2 <- zelig(yr.rnd.numeric ~ meals + mobility, model = "probit.survey", strata=~stype, fpc=~fpc, data = apistrat)
summary(z.out2)
```

The coefficient estimates from this model are identical to
point estimates in the previous example, but the standard errors
are smaller.  When sampling weights are omitted, Zelig estimates
them automatically for "normal.survey" models based on the
user-defined description of sampling designs.  In addition,
when user-defined descriptions of the sampling design are
entered as inputs, variance estimates are better and standard
errors are consequently smaller.

The methods `setx()` and `sim()` can then be run on `z.out2` in the same fashion
described in Example 1.


### Example 3: User has Replicate Weights

Load data for a model using the number of out-of-hospital
cardiac arrests and the number of patients who arrive
alive in hospitals to predict whether each hospital
has been sued (an indicator variable artificially created
here for the purpose of illustration).

```{r, eval = TRUE}
data(scd, package = "survey")
scd$sued <- as.vector(c(0,0,0,1,1,1))
```

Again, for the purpose of illustration, create four Balanced
Repeated Replicate (BRR) weights:

```{r, eval = TRUE}
BRRrep <- 2*cbind(c(1,0,1,0,1,0), c(1,0,0,1,0,1),
                  c(0,1,1,0,0,1), c(0,1,0,1,1,0))
```

Estimate the model using Zelig:

```{r, eval = TRUE}
z.out3 <- zelig(formula=sued ~ arrests + alive ,
                model = "probit.survey",
                repweights = BRRrep, type = "BRR", data = scd)
summary(z.out3)
```

Set the explanatory variables at their means and set
arrests at its 20th and 80th quartiles

```{r, eval = TRUE}
x.low <- setx(z.out3, arrests = quantile(scd$arrests, .2))
x.high <- setx(z.out3, arrests = quantile(scd$arrests,.8))
```

Generate first differences for the effect of the minimum
versus the maximum number of individuals who arrive
alive on the probability that a hospital will be sued:

```{r, eval = TRUE}
s.out3 <- sim(z.out3, x = x.high, x1 = x.low)
summary(s.out3)
```

Generate a second set of fitted values and a plot:

```{r Zelig-probitsurvey3, dev=c("png", "pdf"), eval = TRUE, fig.cap = "Graphs of Quantities of Interest for Probit Survey Model"}
plot(s.out3)
```

The user should also refer to the probit model demo, since probit.survey models can take many of the same options as probit models.



Model
------

Let $Y_i$ be the observed binary dependent variable for
observation $i$ which takes the value of either 0 or 1.

-  The *stochastic component* is given by


$$
    Y_i \; \sim \; \textrm{Bernoulli}(\pi_i),
$$

   where $\pi_i=\Pr(Y_i=1)$.

-  The *systematic component* is


$$
    \pi_i \; = \; \Phi (x_i \beta)
$$

   where $\Phi(\mu)$ is the cumulative distribution function of
   the Normal distribution with mean 0 and unit variance.

Quantities of Interest
------

-  The expected value (qi$ev) is a simulation of predicted probability
   of success


$$
    E(Y) = \pi_i = \Phi(x_i
        \beta),
$$

   given a draw of $\beta$ from its sampling distribution.

-  The predicted value (qi$pr) is a draw from a Bernoulli distribution
   with mean $\pi_i$.

-  The first difference (qi$fd) in expected values is defined as


$$
    \textrm{FD} = \Pr(Y = 1 \mid x_1) - \Pr(Y = 1 \mid x).
$$

-  The risk ratio (qi$rr) is defined as


$$
    \textrm{RR} = \Pr(Y = 1 \mid x_1) / \Pr(Y = 1 \mid x).
$$

-  In conditional prediction models, the average expected treatment
   effect (att.ev) for the treatment group is


$$
    \frac{1}{\sum_{i=1}^n t_i}\sum_{i:t_i=1}^n \left\{ Y_i(t_i=1) -
        E[Y_i(t_i=0)] \right\},
$$

   where $t_i$ is a binary explanatory variable defining the
   treatment ($t_i=1$) and control ($t_i=0$) groups.
   Variation in the simulations are due to uncertainty in simulating
   $E[Y_i(t_i=0)]$, the counterfactual expected value of
   $Y_i$ for observations in the treatment group, under the
   assumption that everything stays the same except that the treatment
   indicator is switched to $t_i=0$.

-  In conditional prediction models, the average predicted treatment
   effect (att.pr) for the treatment group is


$$
    \frac{1}{\sum_{i=1}^n t_i}\sum_{i:t_i=1}^n \left\{ Y_i(t_i=1) -
        \widehat{Y_i(t_i=0)} \right\},
$$

   where $t_i$ is a binary explanatory variable defining the
   treatment ($t_i=1$) and control ($t_i=0$) groups.
   Variation in the simulations are due to uncertainty in simulating
   $\widehat{Y_i(t_i=0)}$, the counterfactual predicted value of
   $Y_i$ for observations in the treatment group, under the
   assumption that everything stays the same except that the treatment
   indicator is switched to $t_i=0$.

Output Values
------

The Zelig object stores fields containing everything needed to
rerun the Zelig output, and all the results and simulations as they are generated.
In addition to the summary commands demonstrated above, some simply utility
functions (known as *getters*) provide easy access to the raw fields most
commonly of use for further investigation.

In the example above `z.out$get_coef()` returns the estimated coefficients, `z.out$get_vcov()` returns the estimated covariance matrix, and `z.out$get_predict()` provides predicted values for all observations in the dataset from the analysis.

See also
------

The probitsurvey model is part of the survey package by Thomas Lumley, which in turn depends heavily on `glm` package. Advanced users may
wish to refer to `help(svyglm)` and `help(family)`.

```{r, eval = TRUE, echo=FALSE, results = "asis"}
z5 <- zprobitsurvey$new()
z5$references()
```
