---
title: "Normal Linear Regression"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{zelig-normal}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

*Built using Zelig version `r packageVersion("Zelig")`*

```{r setup, include=FALSE}
knitr::opts_knit$set(
    stop_on_error = 2L
)
knitr::opts_chunk$set(
    fig.height = 11,
    fig.width = 7
)

options(cite = FALSE)
```
---

Normal Regression for Continuous Dependent Variables with `normal`.

The Normal regression model is a close variant of the more standard
[least squares regression model](zelig_ls.html). Both models specify a
continuous dependent variable as a linear function of a set of explanatory
variables. The Normal model reports maximum likelihood (rather than
least squares) estimates. The two models differ only in their estimate
for the stochastic parameter $\sigma$.

Syntax
------


With the Zelig 4 compatibility wrappers:

```{r, eval = FALSE}
z.out <- zelig(Y ~ X1 + X2, model = "normal", weights = w,
               data = mydata)
x.out <- setx(z.out)
s.out <- sim(z.out, x = x.out)
```

Examples
------

```{r, eval = TRUE, echo = FALSE}
rm(list=ls(pattern="\\.out"))
suppressWarnings(suppressMessages(library(Zelig)))
set.seed(1234)
```

### Basic Example with First Differences

Attach sample data:

```{r, eval = TRUE}
data(macro)
```

Estimate model:

```{r, eval = TRUE}
z.out1 <- zelig(unem ~ gdp + capmob + trade, model = "normal",
                data = macro)
```

Summarize of regression coefficients:

```{r, eval = TRUE}
summary(z.out1)
```

Set explanatory variables to their default (mean/mode) values, with
high (80th percentile) and low (20th percentile) values for trade:

```{r, eval = TRUE}
x.high <- setx(z.out1, trade = quantile(macro$trade, 0.8))
x.low <- setx(z.out1, trade = quantile(macro$trade, 0.2))
```

Generate first differences for the effect of high versus low trade on GDP:

```{r, eval = TRUE}
s.out1 <- sim(z.out1, x = x.high, x1 = x.low)
```

```{r, eval = TRUE}
summary(s.out1)
```

A visual summary of quantities of interest:

```{r Zelig-normal, dev=c("png", "pdf"), eval = TRUE, fig.cap = "Graphs of Quantities of Interest for Zelig-normal"}
plot(s.out1)
```

Model
------

Let $Y_i$ be the continuous dependent variable for observation
$i$.

-  The *stochastic component* is described by a univariate normal model
   with a vector of means $\mu_i$ and scalar variance
   $\sigma^2$:


$$
    Y_i \; \sim \; \textrm{Normal}(\mu_i, \sigma^2).
$$

-  The *systematic component* is


$$
    \mu_i \;= \; x_i \beta,
$$

   where $x_i$ is the vector of $k$ explanatory variables
   and $\beta$ is the vector of coefficients.

Quantities of Interest
------

-  The expected value (qi$ev) is the mean of simulations from the the
   stochastic component,


$$
    E(Y) = \mu_i = x_i \beta,
$$

   given a draw of $\beta$ from its posterior.

-  The predicted value (qi$pr) is drawn from the distribution defined by
   the set of parameters $(\mu_i, \sigma)$.

-  The first difference (qi$fd) is:


$$
    \textrm{FD}\; = \;E(Y \mid x_1) -  E(Y \mid x)
$$

-  In conditional prediction models, the average expected treatment
   effect (att.ev) for the treatment group is


$$
    \frac{1}{\sum_{i=1}^n t_i}\sum_{i:t_i=1}^n \left\{ Y_i(t_i=1) -
        E[Y_i(t_i=0)] \right\},
$$

   where $t_i$ is a binary explanatory variable defining the
   treatment ($t_i=1$) and control ($t_i=0$) groups.
   Variation in the simulations are due to uncertainty in simulating
   $E[Y_i(t_i=0)]$, the counterfactual expected value of
   $Y_i$ for observations in the treatment group, under the
   assumption that everything stays the same except that the treatment
   indicator is switched to $t_i=0$.

-  In conditional prediction models, the average predicted treatment
   effect (att.pr) for the treatment group is


$$
    \frac{1}{\sum_{i=1}^n t_i}\sum_{i:t_i=1}^n \left\{ Y_i(t_i=1) -
        \widehat{Y_i(t_i=0)} \right\},
$$

   where $t_i$ is a binary explanatory variable defining the
   treatment ($t_i=1$) and control ($t_i=0$) groups.
   Variation in the simulations are due to uncertainty in simulating
   $\widehat{Y_i(t_i=0)}$, the counterfactual predicted value of
   $Y_i$ for observations in the treatment group, under the
   assumption that everything stays the same except that the treatment
   indicator is switched to $t_i=0$.

Output Values
------

The Zelig object stores fields containing everything needed to
rerun the Zelig output, and all the results and simulations as they are
generated. In addition to the summary commands demonstrated above, some simply
utility functions (known as *[getters](getters.html)*) provide easy access to
the raw fields most commonly of use for further investigation.

If the `zelig()` call output object is `z.out`, then `coef(z.out)` returns
the estimated coefficients, `vcov(z.out)` returns the estimated covariance
matrix, and `predict(z.out)` provides predicted values for all observations
in the dataset from the analysis.

See also
------

The normal model is part of the stats package by the R Core Team. Advanced users may
wish to refer to `help(glm)` and `help(family)`.

```{r, eval = TRUE, echo=FALSE, results = "asis"}
z5 <- znormal$new()
z5$references()
```
