---
title: "Normal GEE"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{zelig-normalgee}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

*Built using Zelig version `r packageVersion("Zelig")`*

```{r setup, include=FALSE}
knitr::opts_knit$set(
    stop_on_error = 2L
)
knitr::opts_chunk$set(
    fig.height = 11,
    fig.width = 7
)

options(cite = FALSE)
```
---

Generalized Estimating Equation for Normal Regression with `normal.gee`.

The GEE normal estimates the same model as the standard normal
regression. Unlike in normal regression, GEE normal allows for
dependence within clusters, such as in longitudinal data, although its
use is not limited to just panel data. The user must first specify a
"working" correlation matrix for the clusters, which models the
dependence of each observation with other observations in the same
cluster. The "working" correlation matrix is a $T \times T$ matrix
of correlations, where $T$ is the size of the largest cluster and
the elements of the matrix are correlations between within-cluster
observations. The appeal of GEE models is that it gives consistent
estimates of the parameters and consistent estimates of the standard
errors can be obtained using a robust "sandwich" estimator even if the
"working" correlation matrix is incorrectly specified. If the "working"
correlation matrix is correctly specified, GEE models will give more
efficient estimates of the parameters. GEE models measure
population-averaged effects as opposed to cluster-specific effects (See
).

Syntax
------



With the Zelig 4 compatibility wrappers:

```{r, eval = FALSE}
z.out <- zelig(Y ~ X1 + X2, model = "normal.gee",
                       id = "X3", weights = w, data = mydata)
x.out <- setx(z.out)
s.out <- sim(z.out, x = x.out)
```

where `id` is a variable which identifies the clusters. The data
should be sorted by `id` and should be ordered within each cluster
when appropriate.

Additional Inputs
------

-  `robust`: defaults to `TRUE`. If `TRUE`, consistent standard
   errors are estimated using a "sandwich" estimator.

Use the following arguments to specify the structure of the "working"
correlations within clusters:

-  `corstr`: defaults to "independence". It can take on the following
   arguments:

   -  Independence (`corstr = independence`): ${\rm cor}(y_{it}, y_{it'})=0$, $\forall t, t'$ with
      $t\ne t'$. It assumes that there is no correlation within
      the clusters and the model becomes equivalent to standard normal
      regression. The "working" correlation matrix is the identity
      matrix.

   -  Fixed (`corstr = fixed`): If selected, the user must define the
      "working" correlation matrix with the `R` argument rather than
      estimating it from the model.

   -  Stationary $m$ dependent (`corstr = stat_M_dep`):


$$
    {\rm cor}(y_{it}, y_{it'})=\left\{\begin{array}{ccc}
        \alpha_{|t-t'|} & {\rm if} & |t-t'|\le m \\ 0 & {\rm if}
               & |t-t'| > m
             \end{array}\right.
$$
      If (`corstr = stat_M_dep`), you must also specify `Mv` =
      $m$, where $m$ is the number of periods $t$ of
      dependence. Choose this option when the correlations are assumed
      to be the same for observations of the same $|t-t'|$ periods
      apart for $|t-t'| \leq m$.

      Sample "working" correlation for Stationary 2 dependence ($Mv=2$)


$$
    \left( \begin{array}{ccccc}
        1 & \alpha_1 & \alpha_2 & 0 & 0 \\
      \alpha_1 & 1 & \alpha_1 & \alpha_2 & 0 \\
      \alpha_2 & \alpha_1 & 1 & \alpha_1 & \alpha_2 \\
      0 & \alpha_2 & \alpha_1 & 1 & \alpha_1 \\
      0 & 0 & \alpha_2 & \alpha_1 & 1
      \end{array} \right)
$$

   -  Non-stationary $m$ dependent (`corstr = non_stat_M_dep`):


$$
    {\rm cor}(y_{it}, y_{it'})=\left\{\begin{array}{ccc}
        \alpha_{tt'} & {\rm if} & |t-t'|\le m \\ 0 & {\rm if}
               & |t-t'| > m
             \end{array}\right.
$$

      If (`corstr = non_stat_M_dep`), you must also specify `Mv` =
      $m$, where $m$ is the number of periods $t$ of
      dependence. This option relaxes the assumption that the
      correlations are the same for all observations of the same
      $|t-t'|$ periods apart.

      Sample "working" correlation for Non-stationary 2 dependence
      (Mv=2)


$$
    \left( \begin{array}{ccccc}
        1 & \alpha_{12} & \alpha_{13} & 0 & 0 \\
      \alpha_{12} & 1 & \alpha_{23} & \alpha_{24} & 0 \\
      \alpha_{13} & \alpha_{23} & 1 & \alpha_{34} & \alpha_{35} \\
      0 & \alpha_{24} & \alpha_{34} & 1 & \alpha_{45} \\
      0 & 0 & \alpha_{35} & \alpha_{45} & 1
      \end{array} \right)
$$

   -  Exchangeable (`corstr = exchangeable`):


$$
    {\rm cor}(y_{it}, y_{it'})=\alpha,
$$

      $\forall t, t'$ with $t\ne t'$. Choose this option if the correlations are
      assumed to be the same for all observations within the cluster.

      Sample "working" correlation for Exchangeable

$$
    \left( \begin{array}{ccccc}
        1 & \alpha & \alpha & \alpha & \alpha \\
      \alpha & 1 & \alpha & \alpha & \alpha \\
      \alpha & \alpha & 1 & \alpha & \alpha \\
      \alpha & \alpha & \alpha & 1 & \alpha \\
      \alpha & \alpha & \alpha & \alpha & 1
      \end{array} \right)
$$

   -  Stationary $m$\ th order autoregressive (`corstr = AR-M`):
      If (`corstr = AR-M`), you must also specify `Mv` = $m$,
      where $m$ is the number of periods $t$ of dependence.
      For example, the first order autoregressive model (AR-1) implies
      ${\rm cor}(y_{it}, y_{it'})=\alpha^{|t-t'|}, \forall t, t'$ with $t\ne t'$.
      In AR-1, observation 1 and observation 2 have a correlation of
      $\alpha$. Observation 2 and observation 3 also have a
      correlation of $\alpha$. Observation 1 and observation 3
      have a correlation of $\alpha^2$, which is a function of how
      1 and 2 are correlated ($\alpha$) multiplied by how 2 and 3
      are correlated ($\alpha$). Observation 1 and 4 have a
      correlation that is a function of the correlation between 1 and 2,
      2 and 3, and 3 and 4, and so forth.

      Sample "working" correlation for Stationary AR-1 (Mv=1)

$$
      \left( \begin{array}{ccccc}
      1 & \alpha & \alpha^2 & \alpha^3 & \alpha^4 \\
      \alpha & 1 & \alpha & \alpha^2 & \alpha^3 \\
      \alpha^2 & \alpha & 1 & \alpha & \alpha^2 \\
      \alpha^3 & \alpha^2 & \alpha & 1 & \alpha \\
      \alpha^4 & \alpha^3 & \alpha^2 & \alpha & 1
      \end{array} \right)
$$

   -  Unstructured (`corstr = unstructured`): ${\rm
      cor}(y_{it}, y_{it'})=\alpha_{tt'}$, $\forall t, t'$
      with $t\ne t'$. No constraints are placed on the
      correlations, which are then estimated from the data.

-  `Mv`: defaults to 1. It specifies the number of periods of
   correlation and only needs to be specified when `corstr` is
   `stat\_M\_dep`, `non\_stat\_M\_dep`, or `AR-M`.

-  `R`: defaults to `NULL`. It specifies a user-defined correlation
   matrix rather than estimating it from the data. The argument is used
   only when `corstr` is "fixed". The input is a $T \times T$
   matrix of correlations, where $T$ is the size of the largest
   cluster.

Examples
------

```{r, eval = TRUE, echo = FALSE}
rm(list=ls(pattern="\\.out"))
suppressWarnings(suppressMessages(library(Zelig)))
set.seed(1234)
```

### Example with AR-1 Dependence

Attaching the sample turnout dataset:

```{r, eval = TRUE}
data(macro)
```

Estimating model and presenting summary:

```{r, eval = TRUE}
z.out <- zelig(unem ~ gdp + capmob + trade, model =
                       "normal.gee", id = "country", data = macro, corstr = "AR-M")
summary(z.out)
```

Set explanatory variables to their default (mean/mode) values, with high (80th percentile) and low (20th percentile) values:

```{r, eval = TRUE}
x.high <- setx(z.out, trade = quantile(macro$trade, 0.8))
x.low <- setx(z.out, trade = quantile(macro$trade, 0.2))
```

Generate first differences for the effect of high versus low trade on GDP:

```{r, eval = TRUE}
s.out <- sim(z.out, x = x.high, x1 = x.low)
summary(s.out)
```

Generate a plot of quantities of interest:

```{r Zelig-normalgee, dev=c("png", "pdf"), eval = TRUE, fig.cap = "Graphs of Quantities of Interest for Normal GEE Model"}
plot(s.out)
```

The Model
------

Suppose we have a panel dataset, with $Y_{it}$ denoting the
continuous dependent variable for unit $i$ at time $t$.
$Y_{i}$ is a vector or cluster of correlated data where
$y_{it}$ is correlated with $y_{it^\prime}$ for some or all
$t, t^\prime$. Note that the model assumes correlations within
$i$ but independence across $i$.

-  The *stochastic component* is given by the joint and marginal
   distributions


$$
    \begin{aligned}
        Y_{i} &\sim& f(y_{i} \mid \mu_{i})\\
      Y_{it} &\sim& g(y_{it} \mid \mu_{it})
    \end{aligned}

$$

   where $f$ and $g$ are unspecified distributions with
   means $\mu_{i}$ and $\mu_{it}$. GEE models make no
   distributional assumptions and only require three specifications: a
   mean function, a variance function, and a correlation structure.

-  The *systematic component* is the *mean function*, given by:

$$
    \mu_{it} = x_{it} \beta
$$

   where $x_{it}$ is the vector of $k$ explanatory variables
   for unit $i$ at time $t$ and $\beta$ is the vector
   of coefficients.

-  The *variance function* is given by:


$$
    V_{it} = 1
$$

-  The *correlation structure* is defined by a $T \times T$
   "working" correlation matrix, where $T$ is the size of the
   largest cluster. Users must specify the structure of the "working"
   correlation matrix *a priori*. The "working" correlation matrix then
   enters the variance term for each $i$, given by:


$$
    V_{i} = \phi \, A_{i}^{\frac{1}{2}} R_{i}(\alpha) A_{i}^{\frac{1}{2}}
$$

   where $A_{i}$ is a $T \times T$ diagonal matrix with the
   variance function $V_{it} = 1$ as the $t$\ th diagonal
   element (in the case of GEE normal, $A_{i}$ is the identity
   matrix), $R_{i}(\alpha)$ is the "working" correlation matrix,
   and $\phi$ is a scale parameter. The parameters are then
   estimated via a quasi-likelihood approach.

-  In GEE models, if the mean is correctly specified, but the variance
   and correlation structure are incorrectly specified, then GEE models
   provide consistent estimates of the parameters and thus the mean
   function as well, while consistent estimates of the standard errors
   can be obtained via a robust "sandwich" estimator. Similarly, if the
   mean and variance are correctly specified but the correlation
   structure is incorrectly specified, the parameters can be estimated
   consistently and the standard errors can be estimated consistently
   with the sandwich estimator. If all three are specified correctly,
   then the estimates of the parameters are more efficient.

-  The robust "sandwich" estimator gives consistent estimates of the
   standard errors when the correlations are specified incorrectly only
   if the number of units $i$ is relatively large and the number
   of repeated periods $t$ is relatively small. Otherwise, one
   should use the "naïve" model-based standard errors, which assume that
   the specified correlations are close approximations to the true
   underlying correlations. See for more details.

Quantities of Interest
------

-  All quantities of interest are for marginal means rather than joint
   means.

-  The method of bootstrapping generally should not be used in GEE
   models. If you must bootstrap, bootstrapping should be done within
   clusters, which is not currently supported in Zelig. For conditional
   prediction models, data should be matched within clusters.

-  The expected values (qi$ev) for the GEE normal model is the mean of
   simulations from the stochastic component:


$$
    E(Y) =
        \mu_{c}= x_{c} \beta,
$$

   given draws of $\beta$ from its sampling distribution, where
   $x_{c}$ is a vector of values, one for each independent
   variable, chosen by the user.

-  The first difference (qi$fd) for the GEE normal model is defined as


$$
    \textrm{FD} = \Pr(Y = 1 \mid x_1) - \Pr(Y = 1 \mid x).
$$

-  In conditional prediction models, the average expected treatment
   effect (att.ev) for the treatment group is


$$
    \frac{1}{\sum_{i=1}^n \sum_{t=1}^T tr_{it}}\sum_{i:tr_{it}=1}^n \sum_{t:tr_{it}=1}^T \left\{ Y_{it}(tr_{it}=1) -
        E[Y_{it}(tr_{it}=0)] \right\},
$$

   where $tr_{it}$ is a binary explanatory variable defining the
   treatment ($tr_{it}=1$) and control ($tr_{it}=0$) groups.
   Variation in the simulations are due to uncertainty in simulating
   $E[Y_{it}(tr_{it}=0)]$, the counterfactual expected value of
   $Y_{it}$ for observations in the treatment group, under the
   assumption that everything stays the same except that the treatment
   indicator is switched to $tr_{it}=0$.

Output Values
------

The output of each Zelig command contains useful information which you
may view. For example, if you run
`z.out <- zelig(y ~ x, model = "normal.gee", id, data)`, then you may
examine the available information in `z.out` by using
`names(z.out)`, see the coefficients by using z.out$coefficients, and
a default summary of information through `summary(z.out)`. Other
elements available through the $ operator are listed below.

-  From the `zelig()` output object `z.out`, you may extract:

   -  coefficients: parameter estimates for the explanatory variables.

   -  residuals: the working residuals in the final iteration of the
      fit.

   -  fitted.values: the vector of fitted values for the systemic
      component, $\mu_{it}$.

   -  linear.predictors: the vector of $x_{it}\beta$

   -  max.id: the size of the largest cluster.

-  From summary(z.out), you may extract:

   -  coefficients: the parameter estimates with their associated
      standard errors, $p$-values, and $z$-statistics.

   -  working.correlation: the "working" correlation matrix

-  From the sim() output object s.out, you may extract quantities of
   interest arranged as matrices indexed by simulation $\times$
   x-observation (for more than one x-observation). Available quantities
   are:

   -  `qi$ev`: the simulated expected values for the specified values of
      x.

   -  `qi$fd`: the simulated first difference in the expected
      probabilities for the values specified in x and x1.

   -  `qi$att.ev`s: the simulated average expected treatment effect for the
      treated from conditional prediction models.

See also
------

The geeglm function is part of the geepack package by Søren Højsgaard,
Ulrich Halekoh and Jun Yan. Advanced users may wish to refer
to `help(geepack)` and `help(family)`.

```{r, eval = TRUE, echo=FALSE, results = "asis"}
z5 <- znormalgee$new()
z5$references()
```
