---
title: "Normal Survey Regression"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{zelig-normalsurvey}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

*Built using Zelig version `r packageVersion("Zelig")`*

```{r setup, include=FALSE}
knitr::opts_knit$set(
    stop_on_error = 2L
)
knitr::opts_chunk$set(
    fig.height = 11,
    fig.width = 7
)

options(cite = FALSE)
```
---

Normal Regression for Continuous Dependent Variables with Survey Weights with `normal.survey`.

The Normal regression model is a close variant of the more standard
least squares regression model (see [Normal Regress](zelig_normal.html)).
Both models specify a continuous
dependent variable as a linear function of a set of explanatory
variables. The Normal model reports maximum likelihood (rather than
least squares) estimates. The two models differ only in their estimate
for the stochastic parameter $\sigma$.

Syntax
------

With the Zelig 4 compatibility wrappers:

```{r, eval = FALSE}
z.out <- zelig(Y ~ X1 + X2, model = "normal.survey", weights = w,
               data = mydata)
x.out <- setx(z.out)
s.out <- sim(z.out, x = x.out)
```

Examples
------

```{r, eval = TRUE, echo = FALSE}
rm(list=ls(pattern="\\.out"))
suppressWarnings(suppressMessages(library(Zelig)))
set.seed(1234)
```

### Example 1: User has Existing Sample Weights

Attach sample data and variable names:

```{r, eval = TRUE}
data(api, package = "survey")
```

In this example, we will estimate a model using
the percentages of students who receive subsidized
lunch (`meals`) and an indicator for whether schooling is
year-round (`yr.rnd`) to predict California public schools'
academic performance index scores (`api00`). Sampling weights are stored in a
variable called `pw`.

```{r, eval = TRUE}
z.out1 <- zelig(api00 ~ meals + yr.rnd, model = "normal.survey",
                weights = ~pw, data = apistrat)
summary(z.out1)
```

Set explanatory variables to their default (mean/mode) values, and set
a high (80th percentile) and low (20th percentile) value for "meals,"
the percentage of students who receive subsidized meals:

```{r, eval = TRUE}
x.low <- setx(z.out1, meals = quantile(apistrat$meals, 0.2))
x.high <- setx(z.out1, meals = quantile(apistrat$meals, 0.8))
```

Generate first differences for the effect of high versus low "meals"
on academic performance:

```{r, eval = TRUE}
s.out1 <- sim(z.out1, x = x.high, x1 = x.low)
summary(s.out1)
```

Generate a second set of fitted values and a plot:

```{r Zelig-normalsurvey1, dev=c("png", "pdf"), eval = TRUE, fig.cap = "Graphs of Quantities of Interest for Normal Survey Model"}
plot(s.out1)
```

### Example 2: User has Details about Complex Survey Design (but not sample weights)

Suppose that the survey house that provided
the dataset excluded probability weights
but made other details about the survey
design available.  We can still estimate
a model without probability weights that takes
instead variables that identify each the stratum
and/or cluster from which each observation was
selected and the size of the finite sample from
which each observation was selected.

```{r, eval = TRUE}
z.out2 <- zelig(api00 ~ meals + yr.rnd, model = "normal.survey",
                strata = ~stype, fpc = ~fpc, data = apistrat)
summary(z.out2)
```

Note that these results are identical to the results obtained
when pre-existing sampling weights were used.  When sampling
weights are omitted, Zelig estimates them automatically for
"normal.survey" models based on the user-defined description
of sampling designs.  If no description is present, the default
assumption is equal probability sampling.

The methods `setx()` and `sim()` can then be run on `z.out2` in the same fashion
described above.

### Example 3: User has Replicate Weights #####

Load data for a model using the number of out-of-hospital
cardiac arrests to predict the number of patients who arrive
alive in hospitals:

```{r, eval = TRUE}
data(scd, package = "survey")
```

Create four Balanced Repeated Replicate (BRR) weights:

```{r, eval = TRUE}
BRRrep <- 2*cbind(c(1,0,1,0,1,0), c(1,0,0,1,0,1),
                c(0,1,1,0,0,1), c(0,1,0,1,1,0))
```

Estimate the model using Zelig:

```{r, eval = TRUE}
z.out3 <- zelig(formula=alive ~ arrests , model = "normal.survey",
                repweights = BRRrep, type = "BRR",
                data = scd, na.action = NULL)
summary(z.out3)
```

Set the explanatory variable at its minimum and maximum

```{r, eval = TRUE}
x.min <- setx(z.out3, arrests = min(scd$alive))
x.max <- setx(z.out3, arrests = max(scd$alive))
```

Generate first differences for the effect of the minimum
versus the maximum number of cardiac arrests on the number
of people who arrive alive:

```{r, eval = TRUE}
s.out3 <- sim(z.out3, x=x.max, x1=x.min)
summary(s.out3)
```

Generate a second set of fitted values and a plot:

```{r Zelig-normalsurvey3, dev=c("png", "pdf"), eval = TRUE, fig.cap = "Graphs of Quantities of Interest for Normal Survey Model"}
plot(s.out3)
```

The user should also refer to the normal model demo, since
normal.survey models can take many of the same options as
normal models.

Model
------

Let $Y_i$ be the continuous dependent variable for observation
$i$.

-  The *stochastic component* is described by a univariate normal model
   with a vector of means $\mu_i$ and scalar variance
   $\sigma^2$:


$$
    Y_i \; \sim \; \textrm{Normal}(\mu_i, \sigma^2).
$$

-  The *systematic component* is


$$
    \mu_i \;= \; x_i \beta,
$$

   where $x_i$ is the vector of $k$ explanatory variables
   and $\beta$ is the vector of coefficients.

Quantities of Interest
------

-  The expected value (qi$ev) is the mean of simulations from the the
   stochastic component,


$$
    E(Y) = \mu_i = x_i \beta,
$$

   given a draw of $\beta$ from its posterior.

-  The predicted value (qi$pr) is drawn from the distribution defined by
   the set of parameters $(\mu_i, \sigma)$.

-  The first difference (qi$fd) is:


$$
    \textrm{FD}\; = \;E(Y \mid x_1) -  E(Y \mid x)
$$

-  In conditional prediction models, the average expected treatment
   effect (att.ev) for the treatment group is


$$
    \frac{1}{\sum_{i=1}^n t_i}\sum_{i:t_i=1}^n \left\{ Y_i(t_i=1) -
        E[Y_i(t_i=0)] \right\},
$$

   where $t_i$ is a binary explanatory variable defining the
   treatment ($t_i=1$) and control ($t_i=0$) groups.
   Variation in the simulations are due to uncertainty in simulating
   $E[Y_i(t_i=0)]$, the counterfactual expected value of
   $Y_i$ for observations in the treatment group, under the
   assumption that everything stays the same except that the treatment
   indicator is switched to $t_i=0$.

-  In conditional prediction models, the average predicted treatment
   effect (att.pr) for the treatment group is


$$
    \frac{1}{\sum_{i=1}^n t_i}\sum_{i:t_i=1}^n \left\{ Y_i(t_i=1) -
        \widehat{Y_i(t_i=0)} \right\},
$$

   where $t_i$ is a binary explanatory variable defining the
   treatment ($t_i=1$) and control ($t_i=0$) groups.
   Variation in the simulations are due to uncertainty in simulating
   $\widehat{Y_i(t_i=0)}$, the counterfactual predicted value of
   $Y_i$ for observations in the treatment group, under the
   assumption that everything stays the same except that the treatment
   indicator is switched to $t_i=0$.

Output Values
------

The Zelig object stores fields containing everything needed to
rerun the Zelig output, and all the results and simulations as they are generated.
In addition to the summary commands demonstrated above, some simply utility
functions (known as *getters*) provide easy access to the raw fields most
commonly of use for further investigation.

In the example above `z.out$get_coef()` returns the estimated coefficients, `z.out$get_vcov()` returns the estimated covariance matrix, and `z.out$get_predict()` provides predicted values for all observations in the dataset from the analysis.

See also
------

The normalsurvey model is part of the survey package by Thomas Lumley, which in turn depends heavily on `glm` package. Advanced users may
wish to refer to `help(svyglm)` and `help(family)`.

```{r, eval = TRUE, echo=FALSE, results = "asis"}
z5 <- znormalsurvey$new()
z5$references()
```
