---
title: "Quantile Regression"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{zelig-quantile}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

*Built using Zelig version `r packageVersion("Zelig")`*

```{r setup, include=FALSE}
knitr::opts_knit$set(
    stop_on_error = 2L
)
knitr::opts_chunk$set(
    fig.height = 11,
    fig.width = 7
)

options(cite = FALSE)
```
---

Quantile Regression for Continuous Dependent Variables with `rq`.

Use a linear programming implementation of quantile regression to
estimate a linear predictor of the $\tau$\ th conditional quantile
of the population.

Syntax
------

With the Zelig 4 compatibility wrappers:

```{r, eval = FALSE}
z.out <- zelig(Y ~ X1 + X2, model = "rq", weights = w,
               data = mydata, tau = 0.5)
x.out <- setx(z.out)
s.out <- sim(z.out, x = x.out)
```

Additional Inputs
------

In addition to the standard inputs, `zelig()` takes the following additional
options for quantile regression:

-  `tau`: defaults to 0.5. Specifies the conditional quantile(s) that will
   be estimated. 0.5 corresponds to estimating the conditional median,
   0.25 and 0.75 correspond to the conditional quartiles, etc. tau
   vectors with length greater than 1 are not currently supported.
   If tau is set outside of the interval $[0,1]$, zelig returns the
   solution for all possible conditional quantiles given the data, but
   does not support inference on this fit (setx and sim will fail).

-  `se`: a string value that defaults to “nid”. Specifies the method by
   which the covariance matrix of coefficients is estimated during the
   sim stage of analysis. se can take the following values, which are
   passed to the summary.rq function from the quantreg package. These
   descriptions are copied from the summary.rq documentation.

   -  `"iid"` which presumes that the errors are iid and computes an
      estimate of the asymptotic covariance matrix as in KB(1978).

   -  `"nid"` which presumes local (in tau) linearity (in x) of the the
      conditional quantile functions and computes a Huber sandwich
      estimate using a local estimate of the sparsity.

   -  `"ker"` which uses a kernel estimate of the sandwich as proposed by
      Powell(1990).

-  `...`: additional options passed to rq when fitting the model. See
   documentation for rq in the quantreg package for more information.

Examples
------

### Basic Example with First Differences

Attach sample data, in this case a dataset pertaining to the
efficiency of plants that convert ammonia to nitric acid. The
dependent variable, stack.loss, is 10 times the percentage of ammonia
that escaped unconverted:

```{r, eval = TRUE, echo = FALSE}
rm(list=ls(pattern="\\.out"))
suppressWarnings(suppressMessages(library(Zelig)))
set.seed(1234)
```

```{r, eval = TRUE}
data(stackloss)
```

Estimate model:

```{r, eval = TRUE}
z.out1 <- zelig(stack.loss ~ Air.Flow + Water.Temp + Acid.Conc.,
                model = "rq", data = stackloss,
                tau = 0.5)
```

Summarize regression coefficients:

```{r, eval = TRUE}
summary(z.out1)
```

Set explanatory variables to their default (mean/mode) values, with
high (80th percentile) and low (20th percentile) values for the water
temperature variable (the variable that indiates the temperature of water in the plant’s cooling coils):

```{r, eval = TRUE}
x.high <- setx(z.out1, Water.Temp = quantile(stackloss$Water.Temp, 0.8))
x.low <- setx(z.out1, Water.Temp = quantile(stackloss$Water.Temp, 0.2))
```

Generate first differences for the effect of high versus low water temperature on stack loss:

```{r, eval = TRUE}
s.out1 <- sim(z.out1, x = x.high, x1 = x.low)
```

```{r, eval = TRUE}
summary(s.out1)
```

```{r Zelig-quantile1, dev=c("png", "pdf"), eval = TRUE, fig.cap = "Graphs of Quantities of Interest for Quantile Regression"}
plot(s.out1)
```

### Using Dummy Variables

We can estimate a model of unemployment as a function of macroeconomic
indicators and fixed effects for each country (see for help with dummy
variables). Note that you do not need to create dummy variables, as
the program will automatically parse the unique values in the selected
variable into discrete levels.

```{r, eval = TRUE}
data(macro)
z.out2 <- zelig(unem ~ gdp + trade + capmob + as.factor(country),
                model = "rq", tau = 0.5,
                data = macro)
```

Set values for the explanatory variables, using the default mean/mode
values, with country set to the United States and Japan, respectively:

```{r, eval = TRUE}
x.US <- setx(z.out2, country = "United States")
x.Japan <- setx(z.out2, country = "Japan")
```

Simulate quantities of interest:

```{r, eval = TRUE}
s.out2 <- sim(z.out2, x = x.US, x1 = x.Japan)
```

```{r, eval = TRUE}
summary(s.out2)
```

```{r Zelig-quantile2, dev=c("png", "pdf"), eval = TRUE, fig.cap = "Graphs of Quantities of Interest for Quantile Regression"}
plot(s.out2)
```

### Estimating Multiple Quantiles

Using the Engel dataset (from the `quantreg` package) on food expenditure as a function of income,
we can use the “quantile” model to estimate multiple conditional quantiles:

```{r, eval = FALSE}
data(engel, package="quantreg")
z.out3 <- zelig(foodexp ~ income, model = "rq",
                tau = seq(0.1, 0.9, by = 0.1), data = engel)
```

We can summarize the coefficient fits, or plot them to compare them to the least squares conditional mean estimator.

```{r, eval = FALSE}
summary(z.out3)
```

```{r Zelig-quantile3, dev=c("png", "pdf"), eval = FALSE, fig.cap = "Graphs of Quantities of Interest for Quantile Regression"}
plot(summary(z.out3))
```

Set the value of income to the top quartile and the bottom quartile of the income distribution for each fit:

```{r, eval = FALSE}
x.bottom <- setx(z.out3, income = quantile(engel$income, 0.25))
x.top <- setx(z.out3, income = quantile(engel$income, 0.75))
```

Simulate quantities of interest for each fit simultaneously:

```{r, eval = FALSE}
s.out3 <- sim(z.out3, x = x.bottom, x1 = x.top)
```

Summary

```{r, eval = FALSE}
summary(s.out3)
```

Model
------

The quantile estimator is best introduced by considering the sample
median estimator and comparing it to the sample mean estimator. To find
the mean of a sample, we solve for the quantity $\mu$ which
minimizes the sum squared residuals:

$$
    \begin{aligned}
        \mu &=& \arg\min_\mu \sum_i (y_i-\mu)^2\end{aligned}
$$

Estimating a quantile is similar, but we solve for $\xi$ which
minimizes the sum absolute residuals:

$$
    \begin{aligned}
        \xi &=& \arg\min_\xi \sum_i |y_i-\xi|
    \end{aligned}
$$

One can confirm the equivalence of these optimization problems and the
standard mean and median operators by taking the derivative with respect
to the argument and setting it to zero.

The relationship between quantile regression and ordinary least squares
regression is analogous to the relationship between the sample median
and the sample mean, except we are now solving for the conditional
median or conditional mean given covariates and a linear functional
form. The optimization problems for the sample mean and median are then
easily generalized to optimization problems for estimating *conditional*
means or medians by replacing $\mu$ or $\xi$ with a linear
combination of covariates $X'\beta$:

$$
    \begin{aligned}
        \hat\beta_\mathrm{mean} &=& \arg\min_\beta \sum_i (Y_i-X_i'\beta)^2 \nonumber \\
       \hat\beta_\mathrm{median} &=& \arg\min_\beta \sum_i |Y_i-X_i'\beta| \label{median}
    \end{aligned}
$$

Equation [median] can be generalized to provide any quantile of the
conditional distribution, not just the median. We do this by weighting
the aboslute value function asymmetrically in proportion to the
requested $\tau$\ th quantile:

$$
    \begin{aligned}
        \hat\beta_{\tau} &=& \arg\max_\beta \sum \rho(Y_i-X_i'\beta) \label{beta}\\
       \rho &=& \tau(1-I(Y-X_i'\beta > 0)) + (1-\tau)I(Y-X_i'\beta > 0) \nonumber
    \end{aligned}
$$

We call the asymmetric absolute value function a “check function”. This
optimization problem has no closed form solution and is solved via
linear programming.

Equation [beta] now lets us define a conditional quantile estimator.
Suppose that for a given set of covariates $x$, the response
variable $Y$ has as true conditional probability distribution
$f(\theta | x)$ where $f$ can be any probability density
function parametrized by a vector of parameters $\theta$. This
density function defines a value $Q_\tau(\theta|x)$, the true
$\tau$\ th population quantile given $x$. We can write our
conditional quantile estimator $\hat Q_\tau(\theta|x)$ as:

$$
    \begin{aligned}
        \hat Q_\tau(\theta|x) &=& x'\hat\beta_\tau
    \end{aligned}
$$

Where $\hat\beta$ is the vector that solves equation [beta].
Because we solve for the estimator $\hat Q_\tau$ without
constructing a likelihood function, it is not straightforward to specify
a systematic and stochastic component for conditional quantile
estimates. However, systematic and stochastic components do emerge
asymptotically in the large-\ $n$ limit. Asymptotically,
$\hat Q_\tau$ is normally distributed, and can be written with
*stochastic component*

$$
    \begin{aligned}
        \hat Q_\tau &\sim& \mathcal{N}(\mu, \sigma^2),
    \end{aligned}
$$

And *systematic components*

$$
    \begin{aligned}
        \mu &=& x'\hat\beta_\tau\\
       \sigma^2 &=& \frac{\tau(1-\tau)}{n f^2},
    \end{aligned}
$$

Where $n$ is the number of datapoints, and $f$ is the true
population density at the $\tau$\ th conditional quantile. Zelig
uses this asymptotic approximation of stochastic and systematic
components in simulation and numerically estimates the population
density to derive $\sigma^2$. The simulation results should thus
be treated with caution when using small datasets as both this
asymptotic approximation and the population density approximation can
break down.

Quantities of Interest
------

-  The expected value (`qi$ev`) is the mean of simulations from the
   stochastic component,


$$
    E(\hat Q_\tau) = x_i \beta_\tau,
$$

   given a draw of $\beta_\tau$ from its sampling distribution.
   Variation in the expected value distribution comes from estimation
   uncertainty of $\beta_\tau$.

-  The predicted value (`qi$pr`) is the result of a single draw from the
   stochastic component given a draw of $\beta_\tau$ from its
   sampling distribution. The distribution of predicted values should be
   centered around the same place as the expected values but have larger
   variance because it includes both estimation uncertainty and
   fundamental uncertainty.

-  This model does not support conditional prediction.

Output Values
------

The Zelig object stores fields containing everything needed to
rerun the Zelig output, and all the results and simulations as they are generated.
In addition to the summary commands demonstrated above, some simply utility
functions (known as *getters*) provide easy access to the raw fields most
commonly of use for further investigation.

In the example above `z.out$get_coef()` returns the estimated coefficients,
`z.out$get_vcov()` returns the estimated covariance matrix, and
`z.out$get_predict()` provides predicted values for all observations in the
dataset from the analysis.


See also
------

The quantile regression package quantreg by Richard Koenker. In
addition, advanced users may wish to refer to `help(rq)`,
`help(summary.rq)` and `help(rq.object)`.

```{r, eval = TRUE, echo=FALSE, results = "asis"}
z5 <- zquantile$new()
z5$references()
```
