---
title: "Ordered Probit"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{zeligchoice-oprobit}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

*Built using Zelig version `r packageVersion("Zelig")`*

```{r setup, include=FALSE}
knitr::opts_knit$set(
    stop_on_error = 2L
)
knitr::opts_chunk$set(
    fig.height = 11,
    fig.width = 7
)

options(cite = FALSE)
```
---

Ordinal Probit Regression for Ordered Categorical Dependent Variables with `oprobit` in ZeligChoice.

Use the ordinal probit regression model if your dependent variables are
ordered and categorical. They may take on either integer values or
character strings. For a Bayesian implementation of this model, see .

Syntax
------

First load packages:

```{r, eval = TRUE, message=FALSE}
library(zeligverse)
```

With the Zelig 4 compatibility wrappers:

```{r, eval = FALSE}
z.out <- zelig(as.factor(Y) ~ X1 + X23,
               model = "oprobit", data = mydata)
x.out <- setx(z.out)
s.out <- sim(z.out, x = x.out, x1 = NULL)
```

If `Y` takes discrete integer values, the `as.factor()` function will order
it automatically. If `Y` takes on values composed of character strings,
such as “strongly agree”, “agree”, and “disagree”, `as.factor()`` will
order the values in the order in which they appear in Y. You will need
to replace your dependent variable with a factored variable prior to
estimating the model through `zelig()`. See below for more information on
creating ordered factors.

Example
------

```{r, eval = TRUE, echo = FALSE}
rm(list=ls(pattern="\\.out"))
set.seed(1234)
```

### Creating An Ordered Dependent Variable

Load the sample data:

```{r, eval = TRUE}
data(sanction)
```

Create an ordered dependent variable:

```{r, eval = TRUE}
sanction$ncost <- factor(sanction$ncost, ordered = TRUE,
                         levels = c("net gain", "little effect", "modest lost", "major loss"))
```

Estimate the model:

```{r, eval = TRUE}
z.out <- zelig(ncost ~ mil + coop, model = "oprobit", data = sanction)
```

Summarize estimated paramters:

```{r, eval = TRUE}
summary(z.out)
```

Set the explanatory variables to their observed values:

```{r, eval = TRUE}
x.out <- setx(z.out)
```

Simulate fitted values given x.out and view the results:

```{r, eval = TRUE}
s.out <- sim(z.out, x = x.out)
summary(s.out)
```

```{r Zelig-oprobit1, dev=c("png", "pdf"), eval = TRUE, fig.cap = "Graphs of Quantities of Interest for Ordered Probit"}
plot(s.out)
```

### First Differences

Using the sample data `sanction`, let us estimate the empirical model and return the coefficients:

```{r, eval = TRUE}
z.out <- zelig(as.factor(cost) ~ mil + coop, model = "oprobit",
               data = sanction)
summary(z.out)
```

Set the explanatory variables to their means, with `coop` set to `1` (the lowest value) in the baseline case and set
to `4` (the highest value) in the alternative case:

```{r, eval = TRUE}
x.low <- setx(z.out, coop = 1)
x.high <- setx(z.out, coop = 4)
```

Generate simulated fitted values and first differences, and view the results:

```{r, eval = TRUE}
s.out2 <- sim(z.out, x = x.low, x1 = x.high)
summary(s.out2)
```

```{r Zelig-oprobit2, dev=c("png", "pdf"), eval = TRUE, fig.cap = "Graphs of Quantities of Interest for Ordered Probit"}
plot(s.out2)
```

Model
------

Let $Y_i$ be the ordered categorical dependent variable for
observation $i$ that takes one of the integer values from
$1$ to $J$ where $J$ is the total number of
categories.

-  The *stochastic component* is described by an unobserved continuous
   variable, $Y^*_i$, which follows the normal distribution with
   mean $\mu_i$ and unit variance


$$
    Y_i^* \; \sim \; N(\mu_i, 1).
$$

   The observation mechanism is


$$
    Y_i \; = \; j \quad {\rm if} \quad \tau_{j-1} \le Y_i^* \le \tau_j
        \quad {\rm for} \quad j=1,\dots,J.
$$

   where $\tau_k$ for $k=0,\dots,J$ is the threshold
   parameter with the following constraints; [\tau_l]( \tau_m$ for
   all [l](m$ and $\tau_0=-\infty$ and $\tau_J=\infty$.

   Given this observation mechanism, the probability for each category,
   is given by


$$
    \Pr(Y_i = j) \; = \; \Phi(\tau_{j} \mid \mu_i) - \Phi(\tau_{j-1} \mid
        \mu_i) \quad {\rm for} \quad j=1,\dots,J
$$

   where $\Phi(\mu_i)$ is the cumulative distribution function for
   the Normal distribution with mean $\mu_i$ and unit variance.

-  The *systematic component* is given by


$$
    \mu_i \; = \; x_i \beta
$$

   where $x_i$ is the vector of explanatory variables and
   $\beta$ is the vector of coefficients.

Quantities of Interest
------

-  The expected values (qi$ev) for the ordinal probit model are
   simulations of the predicted probabilities for each category:


$$
    E(Y_i = j) \; = \; \Pr(Y_i = j) \; = \; \Phi(\tau_{j} \mid \mu_i)
        - \Phi(\tau_{j-1} \mid  \mu_i) \quad {\rm for} \quad j=1,\dots,J,
$$

   given draws of $\beta$ from its posterior.

-  The predicted value (qi$pr) is the observed value of $Y_i$
   given the underlying standard normal distribution described by
   $\mu_i$.

-  The difference in each of the predicted probabilities (qi$fd) is
   given by


$$
    \Pr(Y=j \mid x_1) \;-\; \Pr(Y=j \mid x) \quad {\rm for} \quad
        j=1,\dots,J.
$$

-  In conditional prediction models, the average expected treatment
   effect (`qi$att.ev`) for the treatment group in category $j$
   is


$$
    \begin{aligned}
        \frac{1}{n_j}\sum_{i:t_{i}=1}^{n_j}[Y_{i}(t_{i}=1)-E[Y_{i}(t_{i}=0)]],\end{aligned}
$$

   where $t_{i}$ is a binary explanatory variable defining the
   treatment ($t_{i}=1$) and control ($t_{i}=0$) groups, and
   $n_j$ is the number of treated observations in category
   $j$.

-  In conditional prediction models, the average predicted treatment
   effect (`qi$att.pr`) for the treatment group in category $j$
   is


$$
    \begin{aligned}
        \frac{1}{n_j}\sum_{i:t_{i}=1}^{n_j}[Y_{i}(t_{i}=1)-\widehat{Y_{i}(t_{i}=0)}],\end{aligned}
$$

   where $t_{i}$ is a binary explanatory variable defining the
   treatment ($t_{i}=1$) and control ($t_{i}=0$) groups, and
   $n_j$ is the number of treated observations in category
   $j$.

Output Values
------

The output of each Zelig command contains useful information which you
may view. For example, if you run
`z.out <- zelig(y ~ x, model = oprobit, data)`, then you may examine
the available information in `z.out` by using `names(z.out)`, see
the coefficients by using z.out$coefficients, and a default summary of
information through `summary(z.out)`. Other elements available through
the $ operator are listed below.

-  From the `zelig()` output object `z.out`, you may extract:

   -  coefficients: the named vector of coefficients.

   -  fitted.values: an $n \times J$ matrix of the in-sample
      fitted values.

   -  predictors: an $n \times (J-1)$ matrix of the linear
      predictors $x_i \beta_j$.

   -  residuals: an $n \times (J-1)$ matrix of the residuals.

   -  zeta: a vector containing the estimated class boundaries.

   -  df.residual: the residual degrees of freedom.

   -  df.total: the total degrees of freedom.

   -  rss: the residual sum of squares.

   -  y: an $n \times J$ matrix of the dependent variables.

   -  zelig.data: the input data frame if save.data = TRUE.

-  From `summary(z.out)``, you may extract:

   -  coef3: a table of the coefficients with their associated standard
      errors and $t$-statistics.

   -  cov.unscaled: the variance-covariance matrix.

   -  pearson.resid: an $n \times (m-1)$ matrix of the Pearson
      residuals.

See also
------

The ordinal probit function is part of the VGAM package by Thomas Yee.
In addition, advanced users may wish to refer to `help(vglm)` in the
VGAM library.

```{r, eval = TRUE, echo=FALSE, results = "asis"}
z5 <- zoprobit$new()
z5$references()
```
