---
title: "Wakefield's Hierarchical Ecological Inference"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{zeligei-eihier}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

*Built using Zelig version `r packageVersion("Zelig")`*

```{r setup, include=FALSE}
knitr::opts_knit$set(
    stop_on_error = 2L
)
knitr::opts_chunk$set(
    fig.height = 11,
    fig.width = 7
)

options(cite = FALSE)
```
---

Wakefield's Hierarchical Ecological Inference Model with `eihier` in ZeligEI.

Syntax
------

The EI models accept several different formula syntaxes.  If $C1$ and $C2$ are the column totals, and $R1$ and $R2$ are the row totals, and $N=R1_i + R2_i = C1_i + C2_i$ is the total in unit $i$, then the formula can be expressed with just one row and one column, with the totals provided separately as:

```{r, eval = FALSE}
z.out <- zelig(C1 ~ R1, N = N, data = data)
```

The argument `N` can be either a numeric vector of the total in each *i*-th unit, or the character name of a variable in the dataset that contains these values.

Or with both rows and columns coupled together, and omitting the totals:

```{r, eval = FALSE}
z.out <- zelig(cbind(C1, C2) ~ cbind(R1, R2), data = data)
```

Additionally, if `C1`, `C2`, `R1`, `R2` are percentages rather than counts, then either formula method above is acceptable, however, `N` must always be provided.

First load packages:

```{r, eval = TRUE, message=FALSE}
library(zeligverse)
```

Here is an example of all the syntax for the analysis using the first syntax method, and the direct use of the reference classes:

```{r, eval = FALSE}
z5 <- zeihier$new()
z5$zelig(C1 ~ R1, N = myN, weights = w, data = myData)
z5$setx()
z5$sim()
```

With the Zelig 4 compatibility wrappers this looks like:

```{r, eval = FALSE}
z.out <- zelig(C1 ~ R1, N = N, model = "eihier", weights = w,
               data = myData)
x.out <- setx(z.out)
s.out <- sim(z.out, x = x.out)
```

### Additional Inputs

In addition, Zelig accepts the following additional inputs for `eihier` to monitor the convergence of the Markov chain:

* `burnin`: number of the initial MCMC iterations to be discarded (defaults to 5,000).
* `mcmc`: number of the MCMC iterations after burnin (defaults to 50,000).
* `thin`: thinning interval for the Markov chain. Only every `thin`-th draw from the Markov chain is kept. The value of mcmc must be divisible by this value. The default value is 1.
* `verbose`: defaults to FALSE. If TRUE, the progress of the sampler (every 10%) is printed to the screen.
* `seed`: seed for the random number generator. The default is NA which corresponds to a random seed of 12345.

The model also accepts the following additional arguments to specify prior parameters used in the model:

* `m0`: prior mean of $\mu_0$ (defaults to 0).
* `M0`: prior variance of $\mu_0$ (defaults to 2.287656).
* `m1`: prior mean of $\mu_1$ (defaults to 0).
* `M1`: prior variance of $\mu_1$ (defaults to 2.287656).
* `a0`: $a0/2$ is the shape parameter for the Inverse Gamma prior on $\sigma_{02}$. The default is 0.825.
* `b0`: $b0/2$ is the scale parameter for the Inverse Gamma prior on $\sigma_{02}$. The default is 0.0105.
* `a1`: $a1/2$ is the shape parameter for the Inverse Gamma prior on $\sigma_{12}$. The default is 0.825.
* `b1`: $b1/2$ is the scale parameter for the Inverse Gamma prior on $\sigma_{12}$. The default is 0.0105.

Users may wish to refer to `help(MCMCdynamicEI)` for more options.

Examples
------

```{r, eval = TRUE, echo = FALSE}
rm(list=ls(pattern="\\.out"))
set.seed(1234)
```

We'll use a dataset from the `ei` package, of black and non-black turnout in 141 precincts.

```{r, eval = TRUE}
library("ei", quietly=TRUE)
data(sample)
```

Here is the model estimated in Zelig.
```{r, eval = TRUE}
z.out <- zeihier$new()
z.out$zelig(t ~ x, N = "n", data = sample)
summary(z.out)
```

You can check for convergence before summarizing the estimates with three diagnostic tests.  See the section [Diagnostics for Zelig Models](diagnostics.html) for examples of the output with interpretation:

```{r, eval = FALSE}
z.out$geweke.diag()
z.out$heidel.diag()
z.out$raftery.diag()
```

See also
------

The Wakefield hierarchical model is part of the MCMCpack package by Andrew Martin, Kevin Quinn, and Jong Hee Park. Advanced users may wish to refer to
`help(MCMChierEI)` in the MCMCpack package.

```{r, eval = TRUE, echo=FALSE, results = "asis"}
z5 <- zeihier$new()
z5$references()
```
