---
title: "Weibull"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{zelig-weibull}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

*Built using Zelig version `r packageVersion("Zelig")`*

```{r setup, include=FALSE}
knitr::opts_knit$set(
    stop_on_error = 2L
)
knitr::opts_chunk$set(
    fig.height = 11,
    fig.width = 7
)

options(cite = FALSE)
```
---

Weibull Regression for Duration Dependent Variables with `weibull`.

Choose the Weibull regression model if the values in your dependent
variable are duration observations. The Weibull model relaxes the
exponential model’s assumption of constant hazard, and allows the
hazard rate to increase or decrease monotonically with respect to
elapsed time.

Syntax
------

With the Zelig 4 compatibility wrappers:

```{r, eval = FALSE}
z.out <- zelig(Surv(Y, C) ~ X, model = "weibull", weights = w,
               data = mydata)
x.out <- setx(z.out)
s.out <- sim(z.out, x = x.out)
```

Weibull models require that the dependent variable be in the form
Surv(Y, C), where Y and C are vectors of length $n$. For each
observation $i$ in 1, …, $n$, the value $y_i$ is the
duration (lifetime, for example), and the associated $c_i$ is a
binary variable such that $c_i = 1$ if the duration is not
censored (*e.g.*, the subject dies during the study) or $c_i = 0$
if the duration is censored (*e.g.*, the subject is still alive at the
end of the study). If $c_i$ is omitted, all Y are assumed to be
completed; that is, time defaults to 1 for all observations.

Input Values
------

In addition to the standard inputs, zelig() takes the following
additional options for weibull regression:

-  robust: defaults to FALSE. If TRUE, zelig() computes robust standard
   errors based on sandwich estimators (see and ) based on the options
   in cluster.

-  cluster: if robust = TRUE, you may select a variable to define groups
   of correlated observations. Let x3 be a variable that consists of
   either discrete numeric values, character strings, or factors that
   define strata. Then

```{r, eval = FALSE}
z.out <- zelig(y ~ x1 + x2, robust = TRUE, cluster = "x3",
               model = "exp", data = mydata)
```

means that the observations can be correlated within the strata defined by the variable x3, and that robust standard errors should be calculated according to those clusters. If $robust = TRUE$ but cluster is not specified, zelig() assumes that each observation falls into its own cluster.

Example
------

```{r, eval = TRUE, echo = FALSE}
rm(list=ls(pattern="\\.out"))
suppressWarnings(suppressMessages(library(Zelig)))
set.seed(1234)
```

Attach the sample data:

```{r, eval = TRUE}
data(coalition)
```

Estimate the model:

```{r, eval = TRUE}
z.out <- zelig(Surv(duration, ciep12) ~ fract + numst2,
               model = "weibull", data = coalition)
```

View the regression output:

```{r, eval = TRUE}
summary(z.out)
```

Set the baseline values (with the ruling coalition in the minority) and
the alternative values (with the ruling coalition in the majority) for
X:

```{r, eval = TRUE}
x.low <- setx(z.out, numst2 = 0)
x.high <- setx(z.out, numst2 = 1)
```

Simulate expected values and first differences:

```{r, eval = TRUE}
s.out <- sim(z.out, x = x.low, x1 = x.high)
```

```{r, eval = TRUE}
summary(s.out)
```

```{r Zelig-weibull, dev=c("png", "pdf"), eval = TRUE, fig.cap = "Graphs of Quantities of Interest for Zelig-weibull"}
plot(s.out)
```

Model
------

Let $Y_i^*$ be the survival time for observation $i$. This
variable might be censored for some observations at a fixed time
$y_c$ such that the fully observed dependent variable,
$Y_i$, is defined as

$$
    Y_i = \left\{ \begin{array}{ll}
        Y_i^* & \textrm{if }Y_i^* \leq y_c \\
         y_c & \textrm{if }Y_i^* > y_c
       \end{array} \right.
$$

-  The *stochastic component* is described by the distribution of the
   partially observed variable $Y^*$. We assume $Y_i^*$
   follows the Weibull distribution whose density function is given by


$$
    f(y_i^*\mid \lambda_i, \alpha) = \frac{\alpha}{\lambda_i^\alpha}
        y_i^{* \alpha-1} \exp \left\{ -\left( \frac{y_i^*}{\lambda_i}
      \right)^{\alpha} \right\}
$$

   for $y_i^* \ge 0$, the scale parameter $\lambda_i > 0$,
   and the shape parameter $\alpha > 0$. The mean of this
   distribution is $\lambda_i \Gamma(1 + 1 / \alpha)$. When $\alpha = 1$, the distribution
   reduces to the exponential distribution (see Section [exp]). (Note
   that the output from zelig() parameterizes scale $ = 1 / \alpha$.)

   In addition, survival models like the Weibull have three additional
   properties. The hazard function $h(t)$ measures the probability
   of not surviving past time $t$ given survival up to $t$.
   In general, the hazard function is equal to $f(t)/S(t)$ where
   the survival function $S(t) = 1 - \int_{0}^t f(s) ds$
   represents the fraction still surviving at time $t$. The
   cumulative hazard function $H(t)$ describes the probability of
   dying before time $t$. In general,
   $H(t)= \int_{0}^{t} h(s) ds = -\log S(t)$. In the case of the
   Weibull model,


$$
    \begin{aligned}
        h(t) &=& \frac{\alpha}{\lambda_i^{\alpha}} t^{\alpha - 1}  \\
      S(t) &=&  \exp \left\{ -\left( \frac{t}{\lambda_i} \right)^{\alpha} \right\} \\
      H(t) &=& \left( \frac{t}{\lambda_i} \right)^{\alpha}
    \end{aligned}
$$
   For the Weibull model, the hazard function $h(t)$ can increase
   or decrease monotonically over time.

-  The *systematic component* $\lambda_i$ is modeled as


$$
    \lambda_i = \exp(x_i \beta),
$$

   where $x_i$ is the vector of explanatory variables, and
   $\beta$ is the vector of coefficients.

Quantities of Interest
------

-  The expected values (qi$ev) for the Weibull model are simulations of
   the expected duration:


$$
    E(Y) = \lambda_i \, \Gamma (1 + \alpha^{-1}),
$$

   given draws of $\beta$ and $\alpha$ from their sampling
   distributions.

-  The predicted value (qi$pr) is drawn from a distribution defined by
   $(\lambda_i, \alpha)$.

-  The first difference (qi$fd) in expected value is


$$
    \textrm{FD} = E(Y \mid x_1) - E(Y \mid x).
$$

-  In conditional prediction models, the average expected treatment
   effect (att.ev) for the treatment group is


$$
    \frac{1}{\sum_{i=1}^n t_i}\sum_{i:t_i=1}^n \left\{ Y_i(t_i=1) -
        E[Y_i(t_i=0)] \right\},
$$

   where $t_i$ is a binary explanatory variable defining the
   treatment ($t_i=1$) and control ($t_i=0$) groups. When
   $Y_i(t_i=1)$ is censored rather than observed, we replace it
   with a simulation from the model given available knowledge of the
   censoring process. Variation in the simulations are due to
   uncertainty in simulating $E[Y_i(t_i=0)]$, the counterfactual
   expected value of $Y_i$ for observations in the treatment
   group, under the assumption that everything stays the same except
   that the treatment indicator is switched to $t_i=0$.

-  In conditional prediction models, the average predicted treatment
   effect (att.pr) for the treatment group is


$$
    \frac{1}{\sum_{i=1}^n t_i}\sum_{i:t_i=1}^n \left\{ Y_i(t_i=1) -
        \widehat{Y_i(t_i=0)} \right\},
$$

   where $t_i$ is a binary explanatory variable defining the
   treatment ($t_i=1$) and control ($t_i=0$) groups. When
   $Y_i(t_i=1)$ is censored rather than observed, we replace it
   with a simulation from the model given available knowledge of the
   censoring process. Variation in the simulations are due to
   uncertainty in simulating $\widehat{Y_i(t_i=0)}$, the
   counterfactual predicted value of $Y_i$ for observations in the
   treatment group, under the assumption that everything stays the same
   except that the treatment indicator is switched to $t_i=0$.

Output Values
------

The Zelig object stores fields containing everything needed to
rerun the Zelig output, and all the results and simulations as they are generated.
In addition to the summary commands demonstrated above, some simply utility
functions (known as *getters*) provide easy access to the raw fields most
commonly of use for further investigation.

In the example above `z.out$get_coef()` returns the estimated coefficients, `z.out$get_vcov()` returns the estimated covariance matrix, and `z.out$get_predict()` provides predicted values for all observations in the dataset from the analysis.

See also
------

The Weibull model is part of the survival library by Terry Therneau,
ported to R by Thomas Lumley. Advanced users may wish to refer to
`help(survfit)` in the survival library.

```{r, eval = TRUE, echo=FALSE, results = "asis"}
z5 <- zweibull$new()
z5$references()
```
