---
title: "Logistic Regression"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{zelig-logit}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

*Built using Zelig version `r packageVersion('Zelig')`*

```{r setup, include=FALSE}
knitr::opts_knit$set(
        stop_on_error = 2L
)
knitr::opts_chunk$set(
    fig.height = 11,
    fig.width = 7
)
```

Logistic Regression for Dichotomous Dependent Variables with `logit`.

Logistic regression specifies a dichotomous dependent variable as a function of a set of explanatory variables.

## Syntax


With the Zelig 4 compatibility wrappers:

```{r eval=FALSE}
z.out <- zelig(Y ~ X1 + X2, model = "logit", weights = w, data = mydata)
x.out <- setx(z.out)
s.out <- sim(z.out, x = x.out, x1 = NULL)
```

## Examples

### Basic example

Load Zelig and attach the sample turnout dataset:

```{r}
library(Zelig)
data(turnout)
```

Estimating parameter values for the logistic regression:

```{r}
z.out1 <- zelig(vote ~ age + race, model = "logit", data = turnout,
                cite = FALSE)
```

Summarize estimated paramters:

```{r}
summary(z.out1)
```

For `logit` models you can also include the argument `odds_ratios = TRUE` in the
`summary` call to return odds ratios estimates ($\mathrm{exp}(\beta)$):

```{r}
summary(z.out1, odds_ratios = TRUE)
```

Set values for the explanatory variables:

```{r}
x.out1 <- setx(z.out1, age = 36, race = "white")
```

Simulate quantities of interest from the posterior distribution:

```{r}
s.out1 <- sim(z.out1, x = x.out1)
summary(s.out1)
```

Show the results graphically:

```{r}
plot(s.out1)
```

## First differences

Estimating the risk difference (and risk ratio) between low education (25th
percentile) and high education (75th percentile) while all the other variables
held at their default values.

```{r}
z.out2 <- zelig(vote ~ race + educate, model = "logit", data = turnout,
                cite = FALSE)
x.high <- setx(z.out2, educate = quantile(turnout$educate, prob = 0.75))
x.low <- setx(z.out2, educate = quantile(turnout$educate, prob = 0.25))
s.out2 <- sim(z.out2, x = x.high, x1 = x.low)
summary(s.out2)
```

```{r}
plot(s.out2)
```

## Model Definition

Let $Y_i$ be the binary dependent variable for observation $i$ which takes the
value of either 0 or 1.

- The *stochastic component* is given by

$$\begin{aligned}
Y_i &\sim& \textrm{Bernoulli}(y_i \mid \pi_i) \\
    &=& \pi_i^{y_i} (1-\pi_i)^{1-y_i}\end{aligned}
$$

where $\pi_i=\Pr(Y_i=1)$.

- The *systematic component* is given by:

$$\pi_i \; = \; \frac{1}{1 + \exp(-x_i \beta)}.$$

where $x_i$ is the vector of $k$ explanatory variables for observation $i$
and $\beta$ is the vector of coefficients.

## Quantities of Interest

- The expected values for the logit model are simulations of the predicted probability of a success:

$$
E(Y) =
  \pi_i= \frac{1}{1 + \exp(-x_i \beta)},
$$

given draws of $\beta$ from its sampling distribution.

- The predicted values are draws from the Binomial distribution with mean equal to the simulated expected value $\pi_i$.

- The first difference for the logit model is defined as

$$\textrm{FD} = \Pr(Y = 1 \mid x_1) - \Pr(Y = 1 \mid x).$$

The risk ratio is defined as

$$\textrm{RR} = \Pr(Y = 1 \mid x_1) \ / \ \Pr(Y = 1 \mid x).$$

In conditional prediction models, the average expected treatment effect (att.ev) for the treatment group is

$$
\frac{1}{\sum_{i=1}^n t_i}\sum_{i:t_i=1}^n \left\{ Y_i(t_i=1) -
      E[Y_i(t_i=0)] \right\},
$$

where $t_i$ is a binary explanatory variable defining the treatment ($t_i=1$) and control ($t_i=0$) groups.
Variation in the simulations are due to uncertainty in simulating $E[Y_i(t_i=0)]$, the counterfactual
expected value of $Y_i$ for observations in the treatment group, under the assumption that everything
stays the same except that the treatment indicator is switched to $t_i=0$.

- In conditional prediction models, the average predicted treatment effect (att.pr) for the treatment group is

$$
\frac{1}{\sum_{i=1}^n t_i}\sum_{i:t_i=1}^n \left\{ Y_i(t_i=1) -
      \widehat{Y_i(t_i=0)}\right\},
$$

where $t_i$ is a binary explanatory variable defining the treatment ($t_i=1$) and control ($t_i=0$) groups. Variation in the simulations are due to uncertainty in simulating $\widehat{Y_i(t_i=0)}$, the counterfactual predicted value of $Y_i$ for observations in the treatment group, under the assumption that everything stays the same except that the treatment indicator is switched to $t_i=0$.

## Output values

The Zelig object stores fields containing everything needed to rerun the Zelig output, and all the results and simulations as they are generated. In addition to the summary functions demonstrated above, use standard R utility functions such as `coef`, `vcov`, `predict` to extract model estimates and `zelig_qi_to_df` to extract simulations.

## See also

The logit model is part of the stats R package. Advanced users may wish to refer to `help(glm)` and `help(family)`.
