---
title: "Tobit"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{zelig-tobit}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

*Built using Zelig version `r packageVersion("Zelig")`*

```{r setup, include=FALSE}
knitr::opts_knit$set(
    stop_on_error = 2L
)
knitr::opts_chunk$set(
    fig.height = 11,
    fig.width = 7
)

options(cite = FALSE)
```
---

Linear Regression for a Left-Censored Dependent Variable with `tobit`.

Tobit regression estimates a linear regression model for a left-censored
dependent variable, where the dependent variable is censored from below.
While the classical tobit model has values censored at 0, you may select
another censoring point. For other linear regression models with fully
observed dependent variables, see [Bayesian regression](zelig_normalbayes.html), [maximum
likelihood normal regression](zelig_normal.html), and [least squares](zelig_ls.html).

Syntax
------

With the Zelig 4 compatibility wrappers:

```{r, eval = FALSE}
z.out <- zelig(Y ~ X1 + X2, below = 0, above = Inf, model = "tobit", weights = w, data = mydata)
x.out <- setx(z.out)
s.out <- sim(z.out, x = x.out)
```

Inputs
------

zelig() accepts the following arguments to specify how the dependent
variable is censored.

-  `below`: (defaults to 0) The point at which the dependent variable
   is censored from below. If any values in the dependent variable are
   observed to be less than the censoring point, it is assumed that that
   particular observation is censored from below at the observed value.
   (See for a Bayesian implementation that supports both left and right
   censoring.)

-  robust: defaults to FALSE. If TRUE, zelig() computes robust standard
   errors based on sandwich estimators (see and ) and the options
   selected in cluster.

-  cluster: if robust = TRUE, you may select a variable to define groups
   of correlated observations. Let x3 be a variable that consists of
   either discrete numeric values, character strings, or factors that
   define strata. Then

```{r, eval=FALSE}
z.out <- zelig(y ~ x1 + x2, robust = TRUE, cluster = "x3",
                        model = "tobit", data = mydata)
```
   means that the observations can be correlated within the strata
   defined by the variable x3, and that robust standard errors should be
   calculated according to those clusters. If robust = TRUE but cluster
   is not specified, zelig() assumes that each observation falls into
   its own cluster.

Zelig users may wish to refer to `help(survreg)` for more information.

Examples
------

```{r, eval = TRUE, echo = FALSE}
rm(list=ls(pattern="\\.out"))
suppressWarnings(suppressMessages(library(Zelig)))
set.seed(1234)
```

### Basic Example

Attaching the sample dataset:

```{r, eval = TRUE}
data(tobin)
```

Estimating linear regression using `tobit`:

```{r, eval = TRUE}
z.out <- zelig(durable ~ age + quant, model = "tobit", data = tobin)
```

Summarize estimated paramters:

```{r, eval = TRUE}
summary(z.out)
```

Setting values for the explanatory variables to their sample averages:

```{r, eval = TRUE}
x.out <- setx(z.out)
```

Simulating quantities of interest from the posterior distribution given `x.out`.

```{r, eval = TRUE}
s.out1 <- sim(z.out, x = x.out)
```

```{r, eval = TRUE}
summary(s.out1)
```

### Simulating First Differences

Set explanatory variables to their default(mean/mode) values, with
high (80th percentile) and low (20th percentile) liquidity ratio
(`quant`):

```{r, eval = TRUE}
x.high <- setx(z.out, quant = quantile(tobin$quant, prob = 0.8))
x.low <- setx(z.out, quant = quantile(tobin$quant, prob = 0.2))
```

Estimating the first difference for the effect of high versus low
liquidity ratio on duration(\ `durable`):

```{r, eval = TRUE}
s.out2 <- sim(z.out, x = x.high, x1 = x.low)
```

```{r, eval = TRUE}
summary(s.out2)
```

```{r Zelig-tobit, dev=c("png", "pdf"), eval = TRUE, fig.cap = "Graphs of Quantities of Interest for Zelig-tobit"}
plot(s.out1)
```

Model
------

-  Let $Y_i^*$ be a latent dependent variable which is distributed
   with *stochastic* component


$$
    \begin{aligned}
        Y_i^* & \sim & \textrm{Normal}(\mu_i, \sigma^2) \\\end{aligned}
$$

   where $\mu_i$ is a vector means and $\sigma^2$ is a
   scalar variance parameter. $Y_i^*$ is not directly observed,
   however. Rather we observed $Y_i$ which is defined as:


$$
    Y_i = \left\{
        \begin{array}{lcl}
      Y_i^*  &\textrm{if} & c <Y_i^* \\
      c    &\textrm{if} & c \ge Y_i^*
      \end{array}\right.
$$

   where $c$ is the lower bound below which $Y_i^*$ is
   censored.

-  The *systematic component* is given by


$$
    \begin{aligned}
        \mu_{i} &=& x_{i} \beta,\end{aligned}
$$

   where $x_{i}$ is the vector of $k$ explanatory variables
   for observation $i$ and $\beta$ is the vector of
   coefficients.

Quantities of Interest
------

-  The expected values (`qi$ev`) for the tobit regression model are
   the same as the expected value of $Y*$:


$$
    E(Y^* | X) = \mu_{i} = x_{i} \beta
$$

-  The first difference (`qi$fd`) for the tobit regression model is
   defined as


$$
    \begin{aligned}
        \text{FD}=E(Y^* \mid x_{1}) - E(Y^* \mid x).\end{aligned}
$$

-  In conditional prediction models, the average expected treatment
   effect (`qi$att.ev`) for the treatment group is


$$
    \begin{aligned}
        \frac{1}{\sum t_{i}}\sum_{i:t_{i}=1}[E[Y^*_{i}(t_{i}=1)]-E[Y^*_{i}(t_{i}=0)]],\end{aligned}
$$

   where $t_{i}$ is a binary explanatory variable defining the
   treatment ($t_{i}=1$) and control ($t_{i}=0$) groups.

Output Values
------

The Zelig object stores fields containing everything needed to
rerun the Zelig output, and all the results and simulations as they are generated.
In addition to the summary commands demonstrated above, some simply utility
functions (known as *getters*) provide easy access to the raw fields most
commonly of use for further investigation.

In the example above `z.out$get_coef()` returns the estimated coefficients, `z.out$get_vcov()` returns the estimated covariance matrix, and `z.out$get_predict()` provides predicted values for all observations in the dataset from the analysis.

See also
------

The tobit function is part of the survival library by Terry Therneau,
ported to R by Thomas Lumley. Advanced users may wish to refer to
`help(survfit)` in the survival library.

```{r, eval = TRUE, echo=FALSE, results = "asis"}
z5 <- ztobit$new()
z5$references()
```
