---
title: "Gamma Regression"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{zelig-gamma}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

*Built using Zelig version `r packageVersion("Zelig")`*

```{r setup, include=FALSE}
knitr::opts_knit$set(
    stop_on_error = 2L
)
knitr::opts_chunk$set(
    fig.height = 11,
    fig.width = 7
)

options(cite = FALSE)
```
---

Gamma Regression for Continuous, Positive Dependent Variables with `gamma`.

Use the gamma regression model if you have a positive-valued dependent
variable such as the number of years a parliamentary cabinet endures, or
the seconds you can stay airborne while jumping. The gamma distribution
assumes that all waiting times are complete by the end of the study
(censoring is not allowed).

Syntax
------


With the Zelig 4 compatibility wrappers:

```{r, eval = FALSE}
z.out <- zelig(Y ~ X1 + X2, model = "gamma", weights = w, data = mydata)
x.out <- setx(z.out)
s.out <- sim(z.out, x = x.out, x1 = NULL)
```

Example
------

Attach the sample data:

```{r, eval = TRUE, echo = FALSE}
rm(list=ls(pattern="\\.out"))
suppressWarnings(suppressMessages(library(Zelig)))
set.seed(1234)
```

```{r, eval = TRUE}
data(coalition)
```

Estimate the model:

```{r, eval = TRUE}
z.out <- zelig(duration ~ fract + numst2, model = "gamma", data = coalition)
```

View the regression output:

```{r, eval = TRUE}
summary(z.out)
```

Set the baseline values (with the ruling coalition in the minority) and
the alternative values (with the ruling coalition in the majority) for
X:

```{r, eval = TRUE}
x.low <- setx(z.out, numst2 = 0)
x.high <- setx(z.out, numst2 = 1)
```

Simulate expected values (`qi$ev`) and first differences (`qi$fd`):

```{r, eval = TRUE}
s.out <- sim(z.out, x = x.low, x1 = x.high)
```

```{r, eval = TRUE}
summary(s.out)
```

```{r Zelig-gamma, dev=c("png", "pdf"), eval = TRUE, fig.cap = "Graphs of Quantities of Interest for Zelig-gamma"}
plot(s.out)
```

Model
------

-   The Gamma distribution with scale parameter $\alpha$ has a
   *stochastic component*:

$$
    \begin{aligned}
        Y &\sim& \textrm{Gamma}(y_i \mid \lambda_i, \alpha) \\
      f(y)  &=& \frac{1}{\alpha^{\lambda_i} \, \Gamma \lambda_i} \, y_i^{\lambda_i
        - 1} \exp -\left\{ \frac{y_i}{\alpha} \right\}\end{aligned}

   | for $\alpha, \lambda_i, y_i > 0$.

-   The *systematic component* is given by

$$
    \lambda_i = \frac{1}{x_i \beta}
$$

Quantities of Interest
------

-   The expected values (qi$ev) are simulations of the mean of the
   stochastic component given draws of $\alpha$ and $\beta$
   from their posteriors:


$$
    E(Y) = \alpha \lambda_i.
$$

-   The predicted values (qi$pr) are draws from the gamma distribution
   for each given set of parameters $(\alpha, \lambda_i)$.

-   If x1 is specified, sim() also returns the differences in the
   expected values (qi$fd),


$$
    E(Y \mid x_1) -
        E(Y \mid x)
$$

   .

-  In conditional prediction models, the average expected treatment
   effect (att.ev) for the treatment group is


$$
    \frac{1}{\sum_{i=1}^n t_i}\sum_{i:t_i=1}^n \left\{ Y_i(t_i=1) -
        E[Y_i(t_i=0)] \right\},
$$

   where $t_i$ is a binary explanatory variable defining the
   treatment ($t_i=1$) and control ($t_i=0$) groups.
   Variation in the simulations are due to uncertainty in simulating
   $E[Y_i(t_i=0)]$, the counterfactual expected value of
   $Y_i$ for observations in the treatment group, under the
   assumption that everything stays the same except that the treatment
   indicator is switched to $t_i=0$.

-  In conditional prediction models, the average predicted treatment
   effect (att.pr) for the treatment group is


$$
    \frac{1}{\sum_{i=1}^n t_i}\sum_{i:t_i=1}^n \left\{ Y_i(t_i=1) -
        \widehat{Y_i(t_i=0)} \right\},
$$

   where $t_i$ is a binary explanatory variable defining the
   treatment ($t_i=1$) and control ($t_i=0$) groups.
   Variation in the simulations are due to uncertainty in simulating
   $\widehat{Y_i(t_i=0)}$, the counterfactual predicted value of
   $Y_i$ for observations in the treatment group, under the
   assumption that everything stays the same except that the treatment
   indicator is switched to $t_i=0$.

Output Values
------

The Zelig object stores fields containing everything needed to
rerun the Zelig output, and all the results and simulations as they are generated.
In addition to the summary commands demonstrated above, some simply utility
functions (known as *getters*) provide easy access to the raw fields most
commonly of use for further investigation.

In the example above `z.out$get_coef()` returns the estimated coefficients, `z.out$get_vcov()` returns the estimated covariance matrix, and `z.out$get_predict()` provides predicted values for all observations in the dataset from the analysis.

See also
------

The gamma model is part of the stats package by the R Core Team. Advanced users may
wish to refer to `help(glm)` and `help(family)`.

```{r, eval = TRUE, echo=FALSE, results = "asis"}
z5 <- zgamma$new()
z5$references()
```
