---
title: "Zelig-gamma"
output:
  html_document:
    number_sections: yes
    toc: yes
bibliography: zelig5.bib
---

<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Model: Zelig-gamma}
-->

# Gamma Regression for Continuous, Positive Dependent Variables

Use the gamma regression model if you have a positive-valued dependent
variable such as the number of years a parliamentary cabinet endures, or
the seconds you can stay airborne while jumping. The gamma distribution
assumes that all waiting times are complete by the end of the study
(censoring is not allowed).

## Syntax

```{r, eval = FALSE}
z.out <- zelig(Y ~ X1 + X2, model = "gamma", data = mydata)
x.out <- setx(z.out)
s.out
```

## Additional Inputs

In addition to the standard inputs, `zelig()` takes the
following additional options for gamma regression:

-   `robust`: defaults to `FALSE`. If
    `TRUE` is selected, `zelig()` computes robust
    standard errors via the `sandwich` package (see
    @Zeileis04). The default type of robust standard error is
    heteroskedastic and autocorrelation consistent (HAC), and assumes
    that observations are ordered by time index.

    In addition, `robust` may be a list with the following
    options:

    -   `method`: Choose from

        -   `“vcovHAC”`: (default if `robust =
            TRUE`) HAC standard errors.

        -   `“kernHAC”`: HAC standard errors using the
            weights given in @Andrews91.

        -   `“weave”`: HAC standard errors using the weights
            given in @LumHea99.

    -   `order.by`: defaults to `NULL` (the
        observations are chronologically ordered as in the original
        data). Optionally, you may specify a vector of weights (either
        as `order.by = z`, where `z` exists
        outside the data frame; or as `order.by = \~z`, where
        `z` is a variable in the data frame). The
        observations are chronologically ordered by the size of
        `z`.

    -   `...`: additional options passed to the functions
        specified in `method`. See the `sandwich`
        library and @Zeileis04 for more options.

## Example

Attach the sample data:

```{r echo = FALSE, warning = FALSE, message = FALSE}
suppressPackageStartupMessages(library(Zelig5, quietly = TRUE))
data(coalition)
```

```{r}
data(coalition)
```

Estimate the model:

```{r}
z.out <- zelig(duration ~ fract + numst2, model = 'gamma', data = coalition)
```

View the regression output:

```{r}
summary(z.out)
```

Set the baseline values (with the ruling coalition in the minority) and
the alternative values (with the ruling coalition in the majority) for
X:

```{r}
x.low <- setx(z.out, numst2 = 0)
x.high <- setx(z.out, numst2 = 1)
```

Simulate expected values (`qi$ev`) and first differences
(`qi$fd`):

```{r}
s.out <- sim(z.out, x = x.low, x1 = x.high)
summary(s.out)
# plot(s.out)
```

## Model

-   The Gamma distribution with scale parameter $\alpha$ has a
    *stochastic component*:

    $$\begin{aligned}
    Y &\sim& \textrm{Gamma}(y_i \mid \lambda_i, \alpha) \\
    f(y)  &=& \frac{1}{\alpha^{\lambda_i} \, \Gamma \lambda_i} \, y_i^{\lambda_i
      - 1} \exp -\left\{ \frac{y_i}{\alpha} \right\}\end{aligned}$$

    for $\alpha, \lambda_i, y_i > 0$.\

-   The *systematic component* is given by

    $$\lambda_i = \frac{1}{x_i \beta}$$

## Quantities of Interest

-   The expected values (`qi$ev`) are simulations of the
    mean of the stochastic component given draws of $\alpha$ and $\beta$
    from their posteriors: $$E(Y) = \alpha \lambda_i.$$

-   The predicted values (`qi$pr`) are draws from the gamma
    distribution for each given set of parameters $(\alpha, \lambda_i)$.

-   If `x1` is specified, `sim()` also returns the
    differences in the expected values (`qi$fd`),
    $$E(Y \mid x_1) -
      E(Y \mid x)$$.

-   In conditional prediction models, the average expected treatment
    effect (`att.ev`) for the treatment group is

    $$\frac{1}{\sum_{i=1}^n t_i}\sum_{i:t_i=1}^n \left\{ Y_i(t_i=1) -
          E[Y_i(t_i=0)] \right\},$$

    where $t_i$ is a binary explanatory variable defining the treatment
    ($t_i=1$) and control ($t_i=0$) groups. Variation in the simulations
    are due to uncertainty in simulating $E[Y_i(t_i=0)]$, the
    counterfactual expected value of $Y_i$ for observations in the
    treatment group, under the assumption that everything stays the same
    except that the treatment indicator is switched to $t_i=0$.

-   In conditional prediction models, the average predicted treatment
    effect (`att.pr`) for the treatment group is

    $$\frac{1}{\sum_{i=1}^n t_i}\sum_{i:t_i=1}^n \left\{ Y_i(t_i=1) -
          \widehat{Y_i(t_i=0)} \right\},$$

    where $t_i$ is a binary explanatory variable defining the treatment
    ($t_i=1$) and control ($t_i=0$) groups. Variation in the simulations
    are due to uncertainty in simulating $\widehat{Y_i(t_i=0)}$, the
    counterfactual predicted value of $Y_i$ for observations in the
    treatment group, under the assumption that everything stays the same
    except that the treatment indicator is switched to $t_i=0$.

## Output Values

The output of each Zelig command contains useful information which you
may view. For example, if you run
`z.out <- zelig(y ~ x, model = gamma, data)`, then you may examine the
available information in `z.out` by using `names(z.out)`, see the
`coefficients` by using `z.out$coefficients`, and
a default summary of information through `summary(z.out)`. Other
elements available through the `$` operator are listed
below.

-   From the `zelig()` output object `z.out`, you
    may extract:

    -   `coefficients`: parameter estimates for the
        explanatory variables.

    -   `residuals`: the working residuals in the final
        iteration of the IWLS fit.

    -   `fitted.values`: the vector of fitted values.

    -   `linear.predictors`: the vector of $x_{i}\beta$.

    -   `aic`: Akaike’s Information Criterion (minus twice
        the maximized log-likelihood plus twice the number of
        coefficients).

    -   `df.residual`: the residual degrees of freedom.

    -   `df.null`: the residual degrees of freedom for the
        null model.

    -   `zelig.data`: the input data frame if `save.data
        = TRUE`.

-   From `summary(z.out)`, you may extract:

    -   `coefficients`: the parameter estimates with their
        associated standard errors, $p$-values, and $t$-statistics.

    -   `cov.scaled`: a $k \times k$ matrix of scaled
        covariances.

    -   `cov.unscaled`: a $k \times k$ matrix of unscaled
        covariances.

-   From the `sim()` output object `s.out`, you
    may extract quantities of interest arranged as matrices indexed by
    simulation $\times$ `x`-observation (for more than one
    `x`-observation). Available quantities are:

    -   `qi$ev`: the simulated expected values for the
        specified values of `x`.

    -   `qi$pr`: the simulated predicted values drawn from a
        distribution defined by $(\alpha, \lambda_i)$.

    -   `qi$fd`: the simulated first difference in the
        expected values for the specified values in `x` and
        `x1`.

    -   `qi$att.ev`: the simulated average expected
        treatment effect for the treated from conditional prediction
        models.

    -   `qi$att.pr`: the simulated average predicted
        treatment effect for the treated from conditional prediction
        models.

How to Cite {#how-to-cite .unnumbered}
-----------

See also {#see-also .unnumbered}
--------

The gamma model is part of the stats package by @VenRip02. Advanced
users may wish to refer to `help(glm)` and `help(family)`, as well as
@McCNel89. Robust standard errors are implemented via the sandwich
package by @Zeileis04. Sample data are from @KinTomWit00.
