---
title: "Least Squares Regression"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{zelig-ls}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

*Built using Zelig version `r packageVersion('Zelig')`*

```{r setup, include=FALSE}
knitr::opts_knit$set(
        stop_on_error = 2L
)
knitr::opts_chunk$set(
    fig.height = 11,
    fig.width = 7
)
```

Least Squares Regression for Continuous Dependent Variables with `ls`.

Use least squares regression analysis to estimate the best linear
predictor for the specified dependent variables.

## Syntax


With the Zelig 4 compatibility wrappers:

```{r, eval = FALSE}
z.out <- zelig(Y ~ X1 + X2, model = "ls", weights = w, data = mydata)
x.out <- setx(z.out)
s.out <- sim(z.out, x = x.out)
```

For more details about the difference between Zelig's reference classes and
compatibility wrappers see the [Quickstart Vignette](zelig5_vs_zelig4.html).

The following examples typically use the compatibility wrappers.

## Examples

### Simple first differences

Load Zelig and attach example data frame:

```{r}
library(Zelig)
data(macro)
```

Estimate model:

```{r message=FALSE}
z.out1 <- zelig(unem ~ gdp + capmob + trade, model = "ls", data = macro,
                cite = FALSE)
```

Summarize regression coefficients:

```{r, eval = TRUE}
summary(z.out1)
```

Set explanatory variables to their default (mean/mode) values, with
high (80th percentile) and low (20th percentile) values for the trade
variable:

```{r, eval = TRUE}
x.high <- setx(z.out1, trade = quantile(macro$trade, 0.8))
x.low <- setx(z.out1, trade = quantile(macro$trade, 0.2))
```

Simulate first differences for the effect of high versus low trade on GDP:

```{r, eval = TRUE}
s.out1 <- sim(z.out1, x = x.high, x1 = x.low)

summary(s.out1)
```

Plot the simulated distributions:

```{r}
plot(s.out1)
```

### Categorical variables, including fixed effects

Estimate a model with fixed effects for each country (see for help
with dummy variables). Note that you do not need to create dummy
variables, as the program will automatically parse the unique values
in the selected variable into discrete levels.

```{r}
z.out2 <- zelig(unem ~ gdp + trade + capmob + as.factor(country), model = "ls",
                data = macro, cite = FALSE)
```

Set values for the explanatory variables, using the default mean/mode
values, with country set to the United States and Japan,
respectively:

```{r}
x.US <- setx(z.out2, country = "United States")
x.Japan <- setx(z.out2, country = "Japan")
```

Simulate quantities of interest:

```{r}
s.out2 <- sim(z.out2, x = x.US, x1 = x.Japan)
```

and plot the results:

```{r}
plot(s.out2)
```

## Model Definition

-  The *stochastic component* is described by a density with mean
   $\mu_i$ and the common variance $\sigma^2$

$$
    Y_i \; \sim \; f(y_i \mid \mu_i, \sigma^2).
$$

-  The *systematic component* models the conditional mean as

$$
   \mu_i =  x_i \beta,
$$

   where $x_i$ is the vector of covariates, and $\beta$ is
   the vector of coefficients.

   The least squares estimator is the best linear predictor of a
   dependent variable given $x_i$, and minimizes the sum of
   squared residuals, $\sum_{i=1}^n (Y_i-x_i \beta)^2$.

## Quantities of Interest Definition

-  The expected value (`qi$ev`) is the mean of simulations from the
   stochastic component,

$$
   E(Y) = x_i \beta,
$$

  given a draw of $\beta$ from its sampling distribution.

-  In conditional prediction models, the average expected treatment
  effect (`att.ev`) for the treatment group is

$$
   \frac{1}{\sum_{i=1}^n t_i}\sum_{i:t_i=1}^n \left\{ Y_i(t_i=1) -
           E[Y_i(t_i=0)] \right\},
$$

   where $t_i$ is a binary explanatory variable defining the
   treatment ($t_i=1$) and control ($t_i=0$) groups.
   Variation in the simulations are due to uncertainty in simulating
   $E[Y_i(t_i=0)]$, the counterfactual expected value of
   $Y_i$ for observations in the treatment group, under the
   assumption that everything stays the same except that the treatment
   indicator is switched to :math: $t_i=0$.

## Output Values

The output of each Zelig function call contains useful information which you
may view. For example, using the `z.out1` object we estimated above,
then you may see the coefficients by using `get_coef`:

```{r}
z.out1$get_coef()
```

or the variance-covariance matrix with `get_vcov`:

```{r}
z.out1$get_vcov()
```

For a default summary of information use either

```{r}
z.out1$summarize()
```

or as we saw before using the wrapper, e.g. `summary(z.out1)`.

-  From the `zelig()` output object `z.out`, you may use the following
'getter' functions to extract:

  - `get_pvalue` and `get_se`: p-values and standard errors, respectively.

  -  `get_residuals`: residuals.

  -  `get_fitted`: fitted values.

  -  `get_df_residual`: the residual degrees of freedom.

  -  `get_model_data`: the input data frame.

-  You can extract the entire original fitted model object with
`from_zelig_model`. For example:

```{r eval=FALSE}
ls.original <- from_zelig_model(z.out1)
```

You can then use `summary` and the component selector `$` to extract:

  -  `coefficients`: the parameter estimates with their associated
      standard errors, $p$-values, and $t$-statistics.

      $$\hat{\beta} \; = \; \left(\sum_{i=1}^n x_i' x_i\right)^{-1} \sum x_i y_i$$

  -  `sigma`: the square root of the estimate variance of the random
      error $e$:

      $$\hat{\sigma} \; = \; \frac{\sum (Y_i-x_i\hat{\beta})^2}{n-k}$$

  -  `r.squared`: the fraction of the variance explained by the model:

      $$R^2 \; = \; 1 - \frac{\sum (Y_i-x_i\hat{\beta})^2}{\sum (y_i -
                  \bar{y})^2}$$

  -  `adj.r.squared`: the above $R^2$ statistic, penalizing for an
      increased number of explanatory variables.

  -  `cov.unscaled`: a $k \times k$ matrix of unscaled covariances.

For example:

```{r}
ls.original <- from_zelig_model(z.out1)
summary(ls.original)$adj.r.squared
```
