---
title: "Factor Bayes"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{zelig-factorbayes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

*Built using Zelig version `r packageVersion("Zelig")`*

```{r setup, include=FALSE}
knitr::opts_knit$set(
    stop_on_error = 2L
)
knitr::opts_chunk$set(
    fig.height = 11,
    fig.width = 7
)

options(cite = FALSE)
```
---

Bayesian Factor Analysis with `factor.bayes`.

Given some unobserved explanatory variables and observed dependent
variables, the Normal theory factor analysis model estimates the latent
factors. The model is implemented using a Markov Chain Monte Carlo
algorithm (Gibbs sampling with data augmentation). <!--For factor analysis
with ordinal dependent variables, see ordered factor analysis (), and
for a mix of types of dependent variables, see the mixed factor analysis
model ().-->

Syntax
------


With the Zelig 4 compatibility wrappers:

```{r, eval = FALSE}
z.out <- zelig(cbind(Y1 ,Y2, Y3) ~ NULL, factors = 2,
               model = "factor.bayes", weights = w, data = mydata)
```

Inputs
------

zelig() takes the following functions for factor.bayes:

-  `Y1`, `Y2`, and `Y3`: variables of interest in factor analysis
   (manifest variables), assumed to be normally distributed. The model
   requires a minimum of three manifest variables.

-  `factors`: number of the factors to be fitted (defaults to 2).

Additional Inputs
------

In addition, zelig() accepts the following additional arguments for
model specification:

-  `lambda.constraints`: list containing the equality or inequality
   constraints on the factor loadings. Choose from one of the following
   forms:

   -  `varname = list()``: by default, no constraints are imposed.

   -  `varname = list(d, c)`: constrains the $d$\ th loading for
      the variable named `varname` to be equal to `c`.

   -  `varname = list(d, +)`: constrains the $d$\ th loading for
      the variable named `varname` to be positive;

   -  `varname = list(d, -)`: constrains the $d$\ th loading for
      the variable named `varname` to be negative.

-  `std.var`: defaults to FALSE (manifest variables are rescaled to
   zero mean, but retain observed variance). If `TRUE`, the manifest
   variables are rescaled to be mean zero and unit variance.

In addition, zelig() accepts the following additional inputs for
bayes.factor:

-  `burnin`: number of the initial MCMC iterations to be discarded
   (defaults to 1,000).

-  `mcmc`: number of the MCMC iterations after burnin (defaults to
   20,000).

-  `thin`: thinning interval for the Markov chain. Only every
   `thin`-th draw from the Markov chain is kept. The value of `mcmc`
   must be divisible by this value. The default value is 1.

-  `verbose`: defaults to FALSE. If `TRUE`, the progress of the
   sampler (every $10\%$) is printed to the screen.

-  `seed`: seed for the random number generator. The default is `NA`
   which corresponds to a random seed 12345.

-  `Lambda.start`: starting values of the factor loading matrix
   $\Lambda$, either a scalar (all unconstrained loadings are set
   to that value), or a matrix with compatible dimensions. The default
   is `NA`, where the start value are set to be 0 for unconstrained
   factor loadings, and 0.5 or $-$\ 0.5 for constrained factor
   loadings (depending on the nature of the constraints).

-  `Psi.start`: starting values for the uniquenesses, either a scalar
   (the starting values for all diagonal elements of $\Psi$ are
   set to be this value), or a vector with length equal to the number of
   manifest variables. In the latter case, the starting values of the
   diagonal elements of $\Psi$ take the values of `Psi.start`.
   The default value is `NA` where the starting values of the all the
   uniquenesses are set to be 0.5.

-  `store.lambda`: defaults to TRUE, which stores the posterior draws
   of the factor loadings.

-  `store.scores`: defaults to FALSE. If TRUE, stores the posterior
   draws of the factor scores. (Storing factor scores may take large
   amount of memory for a large number of draws or observations.)

The model also accepts the following additional arguments to specify
prior parameters:

-  `l0`: mean of the Normal prior for the factor loadings, either a
   scalar or a matrix with the same dimensions as $\Lambda$. If a
   scalar value, that value will be the prior mean for all the factor
   loadings. Defaults to 0.

-  `L0`: precision parameter of the Normal prior for the factor
   loadings, either a scalar or a matrix with the same dimensions as
   $\Lambda$. If `L0` takes a scalar value, then the precision
   matrix will be a diagonal matrix with the diagonal elements set to
   that value. The default value is 0, which leads to an improper prior.

-  `a0`: the shape parameter of the Inverse Gamma prior for the
   uniquenesses is `a0/2`. It can take a scalar value or a vector. The
   default value is 0.001.

-  `b0`: the scale parameter of the Inverse Gamma prior for the
   uniquenesses is `b0/2`. It can take a scalar value or a vector. The
   default value is 0.001.

Zelig users may wish to refer to `help(MCMCfactanal)` for more
information.

Example
------

```{r, eval = TRUE, echo = FALSE}
rm(list=ls(pattern="\\.out"))
suppressWarnings(suppressMessages(library(Zelig)))
set.seed(1234)
```

Attaching the sample dataset:

```{r, eval = TRUE}
data(swiss)
names(swiss) <- c("Fert", "Agr", "Exam", "Educ", "Cath", "InfMort")
```

Factor analysis:

```{r, eval = TRUE}
z.out <- zelig(~ Agr + Exam + Educ + Cath + InfMort,
               model = "factor.bayes", data = swiss,
               factors = 2, verbose = FALSE,
               a0 = 1, b0 = 0.15, burnin = 500, mcmc = 5000)
```

Checking for convergence before summarizing the estimates.  See the section [Diagnostics for Zelig Models](diagnostic.html) for examples of the output with interpretation:

```{r, eval = TRUE}
z.out$geweke.diag()
```

Since the algorithm did not converge, we now add some constraints on
$\Lambda$ to optimize the algorithm:

```{r, eval = TRUE}
z.out <- zelig(~ Agr + Exam + Educ + Cath + InfMort,
               model = "factor.bayes", data = swiss, factors = 2,
               lambda.constraints =
                   list(Exam = list(1,"+"),
                        Exam = list(2,"-"),
                        Educ = c(2, 0),
                        InfMort = c(1, 0)),
               verbose = FALSE, a0 = 1, b0 = 0.15,
               burnin = 500, mcmc = 5000)
z.out$geweke.diag()
z.out$heidel.diag()
z.out$raftery.diag()
summary(z.out)
```

Model
------

Suppose for observation $i$ we observe $K$ variables and
hypothesize that there are $d$ underlying factors such that:

$$
    \begin{aligned}
        Y_i = \Lambda \phi_i+\epsilon_i\end{aligned}
$$

where $Y_{i}$ is the vector of $K$ manifest variables for
observation $i$. $\Lambda$ is the $K \times d$ factor
loading matrix and $\phi_i$ is the $d$-vector of latent
factor scores. Both $\Lambda$ and $\phi$ need to be
estimated.

-  The *stochastic component* is given by:


$$
    \begin{aligned}
        \epsilon_{i}  \sim \textrm{Normal}(0, \Psi).\end{aligned}
$$

   where $\Psi$ is a diagonal, positive definite matrix. The
   diagonal elements of $\Psi$ are referred to as uniquenesses.

-  The *systematic component* is given by


$$
    \begin{aligned}
        \mu_i = E(Y_i) = \Lambda\phi_i\end{aligned}
$$

-  The independent conjugate *prior* for each $\Lambda_{ij}$ is
   given by


$$
    \begin{aligned}
        \Lambda_{ij} \sim \textrm{Normal}(l_{0_{ij}}, L_{0_{ij}}^{-1})
      \textrm{ for } i=1,\ldots, k; \quad j=1,\ldots, d. \end{aligned}

-  The independent conjugate *prior* for each $\Psi_{ii}$ is given
   by


$$
    \begin{aligned}
        \Psi_{ii} \sim \textrm{InverseGamma}(\frac{a_0}{2}, \frac{b_0}{2}), \textrm{ for }
      i = 1, \ldots, k.\end{aligned}

-  The *prior* for $\phi_i$ is


$$
    \begin{aligned}
        \phi_i &\sim& \textrm{Normal}(0, I_d), \textrm{ for } i = 1, \ldots, n.\end{aligned}
$$

   where $I_d$ is a $ d\times d $ identity matrix.

Output Values
------

The output of each `zelig()` call contains useful information which you
may view. For example, if you run:

```{r, eval = FALSE}
z.out <- zelig(cbind(Y1, Y2, Y3), model = "factor.bayes", data)
```

then you may examine the available information in `z.out` by using
`names(z.out)`, see the draws from the posterior distribution of the
`coefficients` by using `z.out$coefficients`, and view a default
summary of information through `summary(z.out)`. Other elements
available through the `$` operator are listed below.

-  From the `zelig()` output object `z.out`, you may extract:

   -  `coefficients`: draws from the posterior distributions of the
      estimated factor loadings and the uniquenesses. If
      `store.scores = TRUE`, the estimated factors scores are also
      contained in `coefficients`.

   -  `data`: the name of the input data frame.

   -  `seed`: the random seed used in the model.

-  Since there are no explanatory variables, the `sim()` procedure is
   not applicable for factor analysis models.

See also
------

Bayesian factor analysis is part of the MCMCpack library by Andrew D.
Martin and Kevin M. Quinn. The convergence diagnostics are part of the
CODA library by Martyn Plummer, Nicky Best, Kate Cowles, and Karen Vines.

```{r, eval = TRUE, echo=FALSE, results = "asis"}
z.out$references()
```
