---
title: "Poisson"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{zelig-poisson}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

*Built using Zelig version `r packageVersion("Zelig")`*

```{r setup, include=FALSE}
knitr::opts_knit$set(
    stop_on_error = 2L
)
knitr::opts_chunk$set(
    fig.height = 11,
    fig.width = 7
)

options(cite = FALSE)
```
---

Poisson Regression for Event Count Dependent Variables with `poisson`.

Use the Poisson regression model if the observations of your dependent
variable represents the number of independent events that occur during a
fixed period of time (see the negative binomial model, , for
over-dispersed event counts.) For a Bayesian implementation of this
model, see .

Syntax
------


With the Zelig 4 compatibility wrappers:

```{r, eval = FALSE}
z.out <- zelig(Y ~ X1 + X2, model = "poisson", weights = w, data = mydata)
x.out <- setx(z.out)
s.out <- sim(z.out, x = x.out)
```

Example
------

```{r, eval = TRUE, echo = FALSE}
rm(list=ls(pattern="\\.out"))
suppressWarnings(suppressMessages(library(Zelig)))
set.seed(1234)
```

Load sample data:

```{r, eval = TRUE}
data(sanction)
```

Estimate Poisson model:

```{r, eval = TRUE}
z.out <- zelig(num ~ target + coop, model = "poisson", data = sanction)
```

```{r, eval = TRUE}
summary(z.out)
```

Set values for the explanatory variables to their default mean values, while changing the $coop$ variable across its range:

```{r, eval = TRUE}
x.low <- setx(z.out, coop = 1)
x.high <- setx1(z.out, coop = 4)
```

Simulate fitted values:

```{r, eval = TRUE}
s.out <- sim(z.out, x = x.low, x1=x.high)
summary(s.out)
```

```{r Zelig-poisson, dev=c("png", "pdf"), eval = TRUE, fig.cap = "Graphs of Quantities of Interest for Zelig-poisson"}
plot(s.out)
```

Model
------

Let $Y_i$ be the number of independent events that occur during a
fixed time period. This variable can take any non-negative integer.

-  The Poisson distribution has *stochastic component*


$$
    Y_i \; \sim \; \textrm{Poisson}(\lambda_i),
$$

   where $\lambda_i$ is the mean and variance parameter.

-  The *systematic component* is


$$
    \lambda_i \; = \; \exp(x_i \beta),
$$

   where $x_i$ is the vector of explanatory variables, and
   $\beta$ is the vector of coefficients.

Quantities of Interest
------

-  The expected value (qi$ev) is the mean of simulations from the
   stochastic component,


$$
    E(Y) = \lambda_i =  \exp(x_i
        \beta),
$$

   given draws of $\beta$ from its sampling distribution.

-  The predicted value (qi$pr) is a random draw from the poisson
   distribution defined by mean $\lambda_i$.

-  The first difference in the expected values (qi$fd) is given by:


$$
    \textrm{FD} \; = \; E(Y | x_1) - E(Y \mid x)
$$

-  In conditional prediction models, the average expected treatment
   effect (att.ev) for the treatment group is


$$
    \frac{1}{\sum_{i=1}^n t_i}\sum_{i:t_i=1}^n \left\{ Y_i(t_i=1) -
        E[Y_i(t_i=0)] \right\},
$$

   where $t_i$ is a binary explanatory variable defining the
   treatment ($t_i=1$) and control ($t_i=0$) groups.
   Variation in the simulations are due to uncertainty in simulating
   $E[Y_i(t_i=0)]$, the counterfactual expected value of
   $Y_i$ for observations in the treatment group, under the
   assumption that everything stays the same except that the treatment
   indicator is switched to $t_i=0$.

-  In conditional prediction models, the average predicted treatment
   effect (att.pr) for the treatment group is


$$
    \frac{1}{\sum_{i=1}^n t_i}\sum_{i:t_i=1}^n \left\{ Y_i(t_i=1) -
        \widehat{Y_i(t_i=0)} \right\},
$$

   where $t_i$ is a binary explanatory variable defining the
   treatment ($t_i=1$) and control ($t_i=0$) groups.
   Variation in the simulations are due to uncertainty in simulating
   $\widehat{Y_i(t_i=0)}$, the counterfactual predicted value of
   $Y_i$ for observations in the treatment group, under the
   assumption that everything stays the same except that the treatment
   indicator is switched to $t_i=0$.

Output Values
------

The Zelig object stores fields containing everything needed to
rerun the Zelig output, and all the results and simulations as they are generated.
In addition to the summary commands demonstrated above, some simply utility
functions (known as *getters*) provide easy access to the raw fields most
commonly of use for further investigation.

In the example above `z.out$get_coef()` returns the estimated coefficients, `z.out$get_vcov()` returns the estimated covariance matrix, and `z.out$get_predict()` provides predicted values for all observations in the dataset from the analysis.

See also
------

The poisson model is part of the stats package by the R Core Team. Advanced users may
wish to refer to `help(glm)` and `help(family)`.

```{r, eval = TRUE, echo=FALSE, results = "asis"}
z5 <- zpoisson$new()
z5$references()
```
