---
title: "Poisson GEE"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{zelig-poissongee}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

*Built using Zelig version `r packageVersion("Zelig")`*

```{r setup, include=FALSE}
knitr::opts_knit$set(
    stop_on_error = 2L
)
knitr::opts_chunk$set(
    fig.height = 11,
    fig.width = 7
)

options(cite = FALSE)
```
---

Generalized Estimating Equation for Poisson Regression with `poisson.gee`.

The GEE poisson estimates the same model as the standard poisson
regression (appropriate when your dependent variable represents the
number of independent events that occur during a fixed period of time).
Unlike in poisson regression, GEE poisson allows for dependence within
clusters, such as in longitudinal data, although its use is not limited
to just panel data. The user must first specify a "working" correlation
matrix for the clusters, which models the dependence of each observation
with other observations in the same cluster. The "working" correlation
matrix is a $T \times T$ matrix of correlations, where $T$
is the size of the largest cluster and the elements of the matrix are
correlations between within-cluster observations. The appeal of GEE
models is that it gives consistent estimates of the parameters and
consistent estimates of the standard errors can be obtained using a
robust "sandwich" estimator even if the "working" correlation matrix is
incorrectly specified. If the "working" correlation matrix is correctly
specified, GEE models will give more efficient estimates of the
parameters. GEE models measure population-averaged effects as opposed to
cluster-specific effects.

Syntax
------


With the Zelig 4 compatibility wrappers:

```{r, eval = FALSE}
z.out <- zelig(Y ~ X1 + X2, model = "poisson.gee",
               id = "X3", weights = w, data = mydata)
x.out <- setx(z.out)
s.out <- sim(z.out, x = x.out)
```

where `id` is a variable which identifies the clusters. The data
should be sorted by `id` and should be ordered within each cluster
when appropriate.

Additional Inputs
------

Use the following arguments to specify the structure of the "working"
correlations within clusters:

- `corstr`: character string specifying the correlation structure:
  "independence", "exchangeable", "ar1", "unstructured" and
  "userdefined"

- See `geeglm` in package `geepack` for other function arguments.

Examples
------

```{r, eval = TRUE, echo = FALSE}	
rm(list=ls(pattern="\\.out"))
suppressWarnings(suppressMessages(library(Zelig)))
set.seed(1234)
``` 

### Example with Exchangeable Dependence

Attaching the sample turnout dataset:

```{r, eval = TRUE}
data(sanction)
```

Variable identifying clusters

```{r, eval = TRUE}
sanction$cluster <- c(rep(c(1:15), 5), rep(c(16), 3))
```

Sorting by cluster

```{r, eval = TRUE}
sorted.sanction <- sanction[order(sanction$cluster),]
```

Estimating model and presenting summary:

```{r, eval = TRUE}
z.out <- zelig(num ~ target + coop, model = "poisson.gee",
               id = "cluster", data = sorted.sanction)
summary(z.out)
```

Set explanatory variables to their default values:

```{r, eval = TRUE}
x.out <- setx(z.out)
```

Simulate quantities of interest

```{r, eval = TRUE}
s.out <- sim(z.out, x = x.out)
summary(s.out)
```

Generate a plot of quantities of interest:

```{r Zelig-poissongee, dev=c("png", "pdf"), eval = TRUE, fig.cap = "Graphs of Quantities of Interest for Zelig-poissongee"}
plot(s.out)
```

The Model
------

Suppose we have a panel dataset, with $Y_{it}$ denoting the
dependent variable of the number of independent events for a fixed
period of time for unit $i$ at time $t$. $Y_{i}$ is a
vector or cluster of correlated data where $y_{it}$ is correlated
with $y_{it^\prime}$ for some or all $t, t^\prime$. Note
that the model assumes correlations within $i$ but independence
across $i$.

-  The *stochastic component* is given by the joint and marginal
   distributions

   
$$
    \begin{aligned}
        Y_{i} &\sim& f(y_{i} \mid \lambda_{i})\\
      Y_{it} &\sim& g(y_{it} \mid \lambda_{it})
    \end{aligned}
$$

   where $f$ and $g$ are unspecified distributions with
   means $\lambda_{i}$ and $\lambda_{it}$. GEE models make
   no distributional assumptions and only require three specifications:
   a mean function, a variance function, and a correlation structure.

-  The *systematic component* is the *mean function*, given by:

   
$$
    \lambda_{it}  = \mathrm{exp}(x_{it} \beta)
$$

   where $x_{it}$ is the vector of $k$ explanatory variables
   for unit $i$ at time $t$ and $\beta$ is the vector
   of coefficients.

-  The *variance function* is given by:

   
$$
    V_{it} = \lambda_{it}
$$

-  The *correlation structure* is defined by a $T \times T$
   "working" correlation matrix, where $T$ is the size of the
   largest cluster. Users must specify the structure of the "working"
   correlation matrix *a priori*. The "working" correlation matrix then
   enters the variance term for each $i$, given by:

   
$$
    V_{i} = \phi \, A_{i}^{\frac{1}{2}} R_{i}(\alpha) A_{i}^{\frac{1}{2}}
$$

   where $A_{i}$ is a $T \times T$ diagonal matrix with the
   variance function $V_{it} = \lambda_{it}$ as the $t$\ th
   diagonal element, $R_{i}(\alpha)$ is the "working" correlation
   matrix, and $\phi$ is a scale parameter. The parameters are
   then estimated via a quasi-likelihood approach.

-  In GEE models, if the mean is correctly specified, but the variance
   and correlation structure are incorrectly specified, then GEE models
   provide consistent estimates of the parameters and thus the mean
   function as well, while consistent estimates of the standard errors
   can be obtained via a robust "sandwich" estimator. Similarly, if the
   mean and variance are correctly specified but the correlation
   structure is incorrectly specified, the parameters can be estimated
   consistently and the standard errors can be estimated consistently
   with the sandwich estimator. If all three are specified correctly,
   then the estimates of the parameters are more efficient.

-  The robust "sandwich" estimator gives consistent estimates of the
   standard errors when the correlations are specified incorrectly only
   if the number of units $i$ is relatively large and the number
   of repeated periods $t$ is relatively small. Otherwise, one
   should use the "naïve" model-based standard errors, which assume that
   the specified correlations are close approximations to the true
   underlying correlations. See for more details.

Quantities of Interest
------

-  All quantities of interest are for marginal means rather than joint
   means.

-  The method of bootstrapping generally should not be used in GEE
   models. If you must bootstrap, bootstrapping should be done within
   clusters, which is not currently supported in Zelig. For conditional
   prediction models, data should be matched within clusters.

-  The expected values (qi$ev) for the GEE poisson model is the mean of
   simulations from the stochastic component:

   
$$
    E(Y) =
        \lambda_{c} = \mathrm{exp}(x_{c} \beta),
$$

   given draws of $\beta$ from its sampling distribution, where
   $x_{c}$ is a vector of values, one for each independent
   variable, chosen by the user.

-  The first difference (qi$fd) for the GEE poisson model is defined as

   
$$
    \textrm{FD} = \Pr(Y = 1 \mid x_1) - \Pr(Y = 1 \mid x).
$$

-  In conditional prediction models, the average expected treatment
   effect (att.ev) for the treatment group is

   
$$
    \frac{1}{\sum_{i=1}^n \sum_{t=1}^T tr_{it}}\sum_{i:tr_{it}=1}^n \sum_{t:tr_{it}=1}^T \left\{ Y_{it}(tr_{it}=1) -
        E[Y_{it}(tr_{it}=0)] \right\},
$$

   where $tr_{it}$ is a binary explanatory variable defining the
   treatment ($tr_{it}=1$) and control ($tr_{it}=0$) groups.
   Variation in the simulations are due to uncertainty in simulating
   $E[Y_{it}(tr_{it}=0)]$, the counterfactual expected value of
   $Y_{it}$ for observations in the treatment group, under the
   assumption that everything stays the same except that the treatment
   indicator is switched to $tr_{it}=0$.

Output Values
------

The output of each Zelig command contains useful information which you
may view. For example, if you run
`z.out <- zelig(y ~ x, model = poisson.gee, id, data)`, then you may
examine the available information in `z.out` by using
`names(z.out)`, see the coefficients by using z.out$coefficients, and
a default summary of information through `summary(z.out)`. Other
elements available through the `$` operator are listed below.

From the `zelig()` output object `z.out`, you may extract:

   -  coefficients: parameter estimates for the explanatory variables.

   -  residuals: the working residuals in the final iteration of the
      fit.

   -  fitted.values: the vector of fitted values for the systemic
      component, $\lambda_{it}$.

   -  linear.predictors: the vector of $x_{it}\beta$

   -  max.id: the size of the largest cluster.

From `summary(z.out)`, you may extract:

   - coefficients: the parameter estimates with their associated
      standard errors, $p$-values, and $z$-statistics.

   - working.correlation: the "working" correlation matrix

-  From the `sim()` output object s.out, you may extract quantities of
   interest arranged as matrices indexed by simulation $\times$
   x-observation (for more than one x-observation). Available quantities
   are:

   -  `qi$ev`: the simulated expected values for the specified values of
      x.

   -  `qi$fd`: the simulated first difference in the expected
      probabilities for the values specified in `x` and `x1`.

   -  `qi$att.ev`: the simulated average expected treatment effect for the
      treated from conditional prediction models.

See also
------

The `geeglm` function is part of the geepack package by Søren Højsgaard,
Ulrich Halekoh and Jun Yan. Advanced users may wish to refer
to `help(geepack)` and `help(family)`.

```{r, eval = TRUE, echo=FALSE, results = "asis"} 
z5 <- zpoissongee$new()
z5$references()
```

