---
title: "Mutlinomial Logit"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{zeligchoice-mlogit}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

*Built using Zelig version `r packageVersion("Zelig")`*

```{r setup, include=FALSE}
knitr::opts_knit$set(
    stop_on_error = 2L
)
knitr::opts_chunk$set(
    fig.height = 11,
    fig.width = 7
)

options(cite = FALSE)
```
---

Multinomial Logistic Regression for Dependent Variables with Unordered Categorical Values with `mlogit` in ZeligChoice.

Use the multinomial logit distribution to model unordered categorical
variables. The dependent variable may be in the format of either
character strings or integer values. See [Multinomial Bayesian Logistic Regression](zelig_mlogitbayes.html) for a Bayesian version of this
model.

Syntax
------

First load packages:

```{r, eval = TRUE, message=FALSE}
library(zeligverse)
```

With reference classes:

```{r, eval = FALSE}
z5 <- zmlogit$new()
z5$zelig(as.factor(Y) ~ X1 + X2, data = mydata)
z5$setx()
z5$sim()
```

With the Zelig 4 compatibility wrappers:

```{r, eval = FALSE}
z.out <- zelig(as.factor(Y) ~ X1 + X23,
               model = "mlogit", data = mydata)
x.out <- setx(z.out)
s.out <- sim(z.out, x = x.out, x1 = NULL)

```

where Y above is supposed to be a factor variable with levels
apples,bananas,oranges. By default, oranges is the last level and
omitted. (You cannot specify a different base level at this time.) For
$J$ equations, there must be $J + 1$ levels.

Examples
------

```{r, eval = TRUE, echo = FALSE}
rm(list=ls(pattern="\\.out"))
set.seed(1234)
```

Load the sample data:

```{r, eval = TRUE}
data(mexico)
```

Estimate the empirical model:

```{r, eval = TRUE}
z.out1 <- zelig(as.factor(vote88) ~ pristr + othcok + othsocok,
                model = "mlogit", data = mexico)
```

Summarize estimated paramters:

```{r, eval = TRUE}
summary(z.out1)
```

Set the explanatory variables to their default values, with $pristr$
(for the strength of the PRI) equal to 1 (weak) in the baseline
values, and equal to 3 (strong) in the alternative values:

```{r, eval = TRUE}
x.weak <- setx(z.out1, pristr = 1)
x.strong <- setx(z.out1, pristr = 3)
```

Generate simulated predicted probabilities `qi$ev` and differences in the predicted probabilities `qi$fd`:

```{r, eval = TRUE}
s.out.mlogit <- sim(z.out1, x = x.strong, x1 = x.weak)
summary(s.out.mlogit)
```

```{r Zelig-mlogit1, dev=c("png", "pdf"), eval = TRUE, fig.cap = "Graphs of Quantities of Interest for Multinomial Logit"}
plot(s.out.mlogit)
```

Model
------

Let $Y_i$ be the unordered categorical dependent variable that
takes one of the values from 1 to $J$, where $J$ is the
total number of categories.

-  The stochastic component is given by


$$
    Y_i \; \sim \; \textrm{Multinomial}(y_{i} \mid \pi_{ij}),
$$

   where $\pi_{ij}=\Pr(Y_i=j)$ for $j=1,\dots,J$.

-  The systemic component is given by:


$$
    \pi_{ij}\; = \; \frac{\exp(x_{i}\beta_{j})}{\sum^{J}_{k = 1}
        \exp(x_{i}\beta_{k})},
$$

   where $x_i$ is the vector of explanatory variables for
   observation $i$, and $\beta_j$ is the vector of
   coefficients for category $j$.

Quantities of Interest
------

-  The expected value (qi$ev) is the predicted probability for each
   category:


$$
    E(Y) \; = \; \pi_{ij}\; = \; \frac{\exp(x_{i}\beta_{j})}{\sum^{J}_{k = 1}
        \exp(x_{i}\beta_{k})}.
$$

-  The predicted value (qi$pr) is a draw from the multinomial
   distribution defined by the predicted probabilities.

-  The first difference in predicted probabilities (qi$fd), for each
   category is given by:


$$
    \textrm{FD}_j = \Pr(Y=j \mid x_1) - \Pr(Y=j \mid x) \quad {\rm for}
        \quad j=1,\dots,J.
$$

-  In conditional prediction models, the average expected treatment
   effect (att.ev) for the treatment group is


$$
    \frac{1}{n_j}\sum_{i:t_i=1}^{n_j} \left\{ Y_i(t_i=1) -
        E[Y_i(t_i=0)] \right\},
$$

   where $t_{i}$ is a binary explanatory variable defining the
   treatment ($t_{i}=1$) and control ($t_{i}=0$) groups, and
   $n_j$ is the number of treated observations in category
   $j$.

-  In conditional prediction models, the average predicted treatment
   effect (att.pr) for the treatment group is


$$
    \frac{1}{n_j}\sum_{i:t_i=1}^{n_j} \left\{ Y_i(t_i=1) -
        \widehat{Y_i(t_i=0)} \right\},
$$

   where $t_{i}$ is a binary explanatory variable defining the
   treatment ($t_{i}=1$) and control ($t_{i}=0$) groups, and
   $n_j$ is the number of treated observations in category
   $j$.

Output Values
------

The output of each Zelig command contains useful information which you
may view. For example, if you run
`z.out <- zelig(y ~ x, model = mlogit, data)`, then you may examine
the available information in `z.out` by using `names(z.out)`, see
the coefficients by using z.out$coefficients, and a default summary of
information through `summary(z.out)`. Other elements available through
the `$` operator are listed below.

-  From the `zelig()`` output object z.out, you may extract:

   -  coefficients: the named vector of coefficients.

   -  fitted.values: an $n \times J$ matrix of the in-sample
      fitted values.

   -  predictors: an $n \times (J-1)$ matrix of the linear
      predictors $x_i \beta_j$.

   -  residuals: an $n \times (J-1)$ matrix of the residuals.

   -  df.residual: the residual degrees of freedom.

   -  df.total: the total degrees of freedom.

   -  rss: the residual sum of squares.

   -  y: an $n \times J$ matrix of the dependent variables.

   -  zelig.data: the input data frame if save.data = TRUE.

-  From summary(z.out), you may extract:

   -  coef3: a table of the coefficients with their associated standard
      errors and $t$-statistics.

   -  cov.unscaled: the variance-covariance matrix.

   -  pearson.resid: an $n \times (m-1)$ matrix of the Pearson
      residuals.

See also
------

The multinomial logit function is part of the VGAM package by Thomas
Yee. In addition, advanced users may wish to refer to `help(vglm)`
in the VGAM library.

```{r, eval = TRUE, echo=FALSE, results = "asis"}
z5 <- zmlogit$new()
z5$references()
```
