---
title: "Normal Bayesian Linear Regression"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{zelig-normalbayes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

*Built using Zelig version `r packageVersion("Zelig")`*

```{r setup, include=FALSE}
knitr::opts_knit$set(
    stop_on_error = 2L
)
knitr::opts_chunk$set(
    fig.height = 11,
    fig.width = 7
)

options(cite = FALSE)
```
---

Bayesian Normal Linear Regression with `normal.bayes`.

Use Bayesian regression to specify a continuous dependent variable as a
linear function of specified explanatory variables. The model is
implemented using a Gibbs sampler. See for the maximum-likelihood
implementation or for the ordinary least squares variation.

Syntax
------


With the Zelig 4 compatibility wrappers:

```{r, eval = FALSE}
z.out <- zelig(Y ~ X1 + X2, model = "normal.bayes", weights = w,
               data = mydata)
x.out <- setx(z.out)
s.out <- sim(z.out, x = x.out)
```

Additional Inputs
------

Use the following arguments to monitor the convergence of the Markov
chain:

-  `burnin`: number of the initial MCMC iterations to be discarded
   (defaults to 1,000).

-  `mcmc`: number of the MCMC iterations after burnin (defaults to
   10,000).

-  `thin`: thinning interval for the Markov chain. Only every
   `thin`-th draw from the Markov chain is kept. The value of `mcmc`
   must be divisible by this value. The default value is 1.

-  `verbose`: defaults to FALSE. If `TRUE`, the progress of the
   sampler (every $10\%$) is printed to the screen.

-  `seed`: seed for the random number generator. The default is
   `NA`, which corresponds to a random seed of 12345.

-  `beta.start`: starting values for the Markov chain, either a scalar
   or vector with length equal to the number of estimated coefficients.
   The default is `NA`, which uses the least squares estimates as the
   starting values.

Use the following arguments to specify the model’s priors:

-  `b0`: prior mean for the coefficients, either a numeric vector or a
   scalar. If a scalar, that value will be the prior mean for all the
   coefficients. The default is 0.

-  `B0`: prior precision parameter for the coefficients, either a
   square matrix (with the dimensions equal to the number of the
   coefficients) or a scalar. If a scalar, that value times an identity
   matrix will be the prior precision parameter. The default is 0, which
   leads to an improper prior.

-  `c0`: `c0/2` is the shape parameter for the Inverse Gamma prior
   on the variance of the disturbance terms.

-  `d0`: `d0/2` is the scale parameter for the Inverse Gamma prior
   on the variance of the disturbance terms.

Zelig users may wish to refer to `help(MCMCregress)` for more
information.

Examples
------

```{r, eval = TRUE, echo = FALSE}
rm(list=ls(pattern="\\.out"))
suppressWarnings(suppressMessages(library(Zelig)))
set.seed(1234)
```

### Basic Example

Attaching the sample dataset:

```{r, eval = TRUE}
data(macro)
```

Estimating linear regression using `normal.bayes`:

```{r, eval = TRUE}
z.out <- zelig(unem ~ gdp + capmob + trade, model = "normal.bayes",
               data = macro, verbose = FALSE)
```

You can check for convergence before summarizing the estimates with three diagnostic tests.  See the section [Diagnostics for Zelig Models](diagnostics.html) for examples of the output with interpretation:

```{r, eval = FALSE}
z.out$geweke.diag()
z.out$heidel.diag()
z.out$raftery.diag()
```

```{r, eval = TRUE}
summary(z.out)
```

Setting values for the explanatory variables to their sample averages:

```{r, eval = TRUE}
x.out <- setx(z.out)
```

Simulating quantities of interest from the posterior distribution given `x.out`:

```{r, eval = TRUE}
s.out1 <- sim(z.out, x = x.out)
summary(s.out1)
```

### Simulating First Differences

Set explanatory variables to their default(mean/mode) values, with high (80th percentile) and low (20th percentile) trade on GDP:

```{r, eval = TRUE}
x.high <- setx(z.out, trade = quantile(macro$trade, prob = 0.8))
x.low <- setx(z.out, trade = quantile(macro$trade, prob = 0.2))
```

Estimating the first difference for the effect of high versus low trade on unemployment rate:

```{r, eval = TRUE}
s.out2 <- sim(z.out, x = x.high, x1 = x.low)
summary(s.out2)
```

Model
------

-  The *stochastic component* is given by


$$
    \begin{aligned}
        \epsilon_{i}  &  \sim & \textrm{Normal}(0, \sigma^2)
    \end{aligned}
$$

   where $\epsilon_{i}=Y_i-\mu_i$.

-  The *systematic component* is given by


$$
    \begin{aligned}
        \mu_{i}= x_{i} \beta,
    \end{aligned}
$$

   where $x_{i}$ is the vector of $k$ explanatory variables
   for observation $i$ and $\beta$ is the vector of
   coefficients.

-  The *semi-conjugate priors* for $\beta$ and $\sigma^2$
   are given by


$$
    \begin{aligned}
        \beta & \sim & \textrm{Normal}_k \left( b_{0},B_{0}^{-1}\right) \\
      \sigma^{2} & \sim & {\rm InverseGamma}\left( \frac{c_0}{2},
      \frac{d_0}{2} \right)
    \end{aligned}
$$

   where $b_{0}$ is the vector of means for the $k$
   explanatory variables, $B_{0}$ is the $k\times k$
   precision matrix (the inverse of a variance-covariance matrix), and
   $c_0/2$ and $d_0/2$ are the shape and scale parameters
   for $\sigma^{2}$. Note that $\beta$ and $\sigma^2$
   are assumed to be *a priori* independent.

Quantities of Interest
------

-  The expected values (`qi$ev`) for the linear regression model are
   calculated as following:

$$
    \begin{aligned}
        E(Y) = x_{i} \beta,
    \end{aligned}
$$

   given posterior draws of $\beta$ based on the MCMC iterations.

-  The first difference (`qi$fd`) for the linear regression model is
   defined as


$$
    \begin{aligned}
        \text{FD}=E(Y\mid X_{1})-E(Y\mid X).
    \end{aligned}
$$

-  In conditional prediction models, the average expected treatment
   effect (`qi$att.ev`) for the treatment group is


$$
    \begin{aligned}
        \frac{1}{\sum_{i=1}^n t_{i}}\sum_{i:t_{i}=1} \{
      Y_{i}(t_{i}=1)-E[Y_{i}(t_{i}=0)] \},
    \end{aligned}
$$

   where $t_{i}$ is a binary explanatory variable defining the
   treatment ($t_{i}=1$) and control ($t_{i}=0$) groups.

-  In conditional prediction models, the average predicted treatment
   effect (att.pr) for the treatment group is


$$
    \frac{1}{\sum_{i=1}^n t_i}\sum_{i:t_i=1}^n \left\{ Y_i(t_i=1) -
        \widehat{Y_i(t_i=0)} \right\},
$$

   where $t_{i}$ is a binary explanatory variable defining the
   treatment ($t_{i}=1$) and control ($t_{i}=0$) groups.

Output Values
------

The output of each Zelig command contains useful information which you
may view. For example, if you run:

```{r, eval = FALSE}
z.out <- zelig(y ~ x, model = "normal.bayes", data)
```

then you may examine the available information in `z.out` by using
`names(z.out)`, see the draws from the posterior distribution of the
`coefficients` by using `z.out$coefficients`, and view a default
summary of information through `summary(z.out)`. Other elements
available through the `$` operator are listed below.

-  From the `zelig()` output object `z.out`, you may extract:

   -  `coefficients`: draws from the posterior distributions of the
      estimated parameters. The first $k$ columns contain the
      posterior draws of the coefficients $\beta$, and the last
      column contains the posterior draws of the variance
      $\sigma^2$.

   -  `zelig.data`: the input data frame if save.data = TRUE.

   -  `seed`: the random seed used in the model.

-  From the `sim()` output object `s.out`:

   -  `qi$ev`: the simulated expected values for the specified values
      of `x`.

   -  `qi$fd`: the simulated first difference in the expected values
      for the values specified in `x` and `x1`.

   -  `qi$att.ev`: the simulated average expected treatment effect for
      the treated from conditional prediction models.

See also
------

Bayesian normal regression is part of the MCMCpack library by Andrew D.
Martin and Kevin M. Quinn . The convergence diagnostics are part of the
CODA library by Martyn Plummer, Nicky Best, Kate Cowles, and Karen Vines.

```{r, eval = TRUE, echo=FALSE, results = "asis"}
z5 <- znormalbayes$new()
z5$references()
```
