---
title: "Tobit Bayes"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{zelig-tobitbayes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

*Built using Zelig version `r packageVersion("Zelig")`*

```{r setup, include=FALSE}
knitr::opts_knit$set(
    stop_on_error = 2L
)
knitr::opts_chunk$set(
    fig.height = 11,
    fig.width = 7
)

options(cite = FALSE)
```
---

Bayesian Tobit Regression with `tobit.bayes`.

Bayesian tobit regression estimates a linear regression model with a
censored dependent variable using a Gibbs sampler. The dependent
variable may be censored from below and/or from above. For other linear
regression models with fully observed dependent variables, see Bayesian
regression, maximum likelihood normal regression, or least squares.

Syntax
------

With the Zelig 4 compatibility wrappers:

```{r, eval = FALSE}
z.out <- zelig(Y ~ X1 + X2, below = 0, above = Inf,
               model = "tobit.bayes", weights = w, data = mydata)
x.out <- setx(z.out)
s.out <- sim(z.out, x = x.out)
```

Inputs
------

zelig() accepts the following arguments to specify how the dependent
variable is censored.

-  `below`: point at which the dependent variable is censored from
   below. If the dependent variable is only censored from above, set
   `below = -Inf`. The default value is 0.

-  `above`: point at which the dependent variable is censored from
   above. If the dependent variable is only censored from below, set
   `above = Inf`. The default value is `Inf`.

Additional Inputs
------

Use the following arguments to monitor the convergence of the Markov
chain:

-  `burnin`: number of the initial MCMC iterations to be discarded
   (defaults to 1,000).

-  `mcmc`: number of the MCMC iterations after burnin (defaults to
   10,000).

-  `thin`: thinning interval for the Markov chain. Only every
   `thin`-th draw from the Markov chain is kept. The value of `mcmc`
   must be divisible by this value. The default value is 1.

-  `verbose`: defaults to FALSE. If `TRUE`, the progress of the
   sampler (every $10\%$) is printed to the screen.

-  `seed`: seed for the random number generator. The default is `NA`
   which corresponds to a random seed of 12345.

-  `beta.start`: starting values for the Markov chain, either a scalar
   or vector with length equal to the number of estimated coefficients.
   The default is `NA`, such that the least squares estimates are used
   as the starting values.

Use the following parameters to specify the model’s priors:

-  `b0`: prior mean for the coefficients, either a numeric vector or a
   scalar. If a scalar, that value will be the prior mean for all
   coefficients. The default is 0.

-  `B0`: prior precision parameter for the coefficients, either a
   square matrix (with the dimensions equal to the number of the
   coefficients) or a scalar. If a scalar, that value times an identity
   matrix will be the prior precision parameter. The default is 0, which
   leads to an improper prior.

-  `c0`: `c0/2` is the shape parameter for the Inverse Gamma prior
   on the variance of the disturbance terms.

-  `d0`: `d0/2` is the scale parameter for the Inverse Gamma prior
   on the variance of the disturbance terms.

Zelig users may wish to refer to `help(MCMCtobit)` for more
information.

Examples
------

```{r, eval = TRUE, echo = FALSE}
rm(list=ls(pattern="\\.out"))
suppressWarnings(suppressMessages(library(Zelig)))
set.seed(1234)
```

### Basic Example

Attaching the sample dataset:

```{r, eval = TRUE}
data(tobin)
```

Estimating linear regression using `tobit.bayes`:

```{r, eval = TRUE}
z.out <- zelig(durable ~ age + quant, model = "tobit.bayes",
               data = tobin, verbose = FALSE)
```

You can check for convergence before summarizing the estimates with three diagnostic tests.  See the section [Diagnostics for Zelig Models](diagnostics.html) for examples of the output with interpretation:

```{r, eval = FALSE}
z.out$geweke.diag()
z.out$heidel.diag()
z.out$raftery.diag()
```

```{r, eval = TRUE}
summary(z.out)
```

Setting values for the explanatory variables to their sample averages:

```{r, eval = TRUE}
x.out <- setx(z.out)
```

Simulating quantities of interest from the posterior distribution given `x.out`.

```{r, eval = TRUE}
s.out1 <- sim(z.out, x = x.out)
```

```{r, eval = TRUE}
summary(s.out1)
```

### Simulating First Differences

Set explanatory variables to their default(mean/mode) values, with
high (80th percentile) and low (20th percentile) liquidity ratio (`quant`):

```{r, eval = TRUE}
x.high <- setx(z.out, quant = quantile(tobin$quant, prob = 0.8))
x.low <- setx(z.out, quant = quantile(tobin$quant, prob = 0.2))
```

Estimating the first difference for the effect of high versus low liquidity ratio on duration(`durable`):

```{r, eval = TRUE}
s.out2 <- sim(z.out, x = x.high, x1 = x.low)
```

```{r, eval = TRUE}
summary(s.out2)
```

Model
------

Let $Y_i^*$ be the dependent variable which is not directly
observed. Instead, we observe $Y_i$ which is defined as following:

$$
    Y_i = \left\{
        \begin{array}{lcl}
   Y_i^*  &\textrm{if} & c_1<Y_i^*<c_2 \\
   c_1    &\textrm{if} & c_1 \ge Y_i^* \\
   c_2    &\textrm{if} & c_2 \le Y_i^*
   \end{array}\right.
$$

where $c_1$ is the lower bound below which $Y_i^*$ is
censored, and $c_2$ is the upper bound above which $Y_i^*$
is censored.

-  The *stochastic component* is given by


$$
    \begin{aligned}
        \epsilon_{i}  &  \sim & \textrm{Normal}(0, \sigma^2)\end{aligned}
$$

   where $\epsilon_{i}=Y^*_i-\mu_i$.

-  The *systematic component* is given by


$$
    \begin{aligned}
        \mu_{i}= x_{i} \beta,\end{aligned}
$$

   where $x_{i}$ is the vector of $k$ explanatory variables
   for observation $i$ and $\beta$ is the vector of
   coefficients.

-  The *semi-conjugate priors* for $\beta$ and $\sigma^2$
   are given by


$$
    \begin{aligned}
        \beta & \sim & \textrm{Normal}_k \left( b_{0},B_{0}^{-1}\right) \\
      \sigma^{2} & \sim & \textrm{InverseGamma} \left( \frac{c_0}{2}, \frac{d_0}{2}
      \right) \end{aligned}
$$

   where $b_{0}$ is the vector of means for the $k$
   explanatory variables, $B_{0}$ is the $k\times k$
   precision matrix (the inverse of a variance-covariance matrix), and
   $c_0/2$ and $d_0/2$ are the shape and scale parameters
   for $\sigma^{2}$. Note that $\beta$ and $\sigma^2$
   are assumed *a priori* independent.

Quantities of Interest
------

-  The expected values (`qi$ev`) for the tobit regression model is
   calculated as following. Let


$$
    \begin{aligned}
        \Phi_1 &=& \Phi\left(\frac{(c_1 - x \beta)}{\sigma}\right) \\
      \Phi_2 &=& \Phi\left(\frac{(c_2 - x \beta)}{\sigma}\right) \\
      \phi_1 &=& \phi\left(\frac{(c_1 - x \beta)}{\sigma}\right) \\
      \phi_2 &=& \phi\left(\frac{(c_2 - x \beta)}{\sigma}\right) \end{aligned}
$$

   where $\Phi(\cdot)$ is the (cumulative) Normal density function
   and $\phi(\cdot)$ is the Normal probability density function of
   the standard normal distribution. Then the expected values are


$$
    \begin{aligned}
        E(Y|x) &=& P(Y^* \le c_1|x) c_1+P(c_1<Y^*<c_2|x) E(Y^* \mid c_1<Y^*<c_2, x)+P(Y^* \ge c_2) c_2 \\
         &=& \Phi_{1}c_1 + x \beta(\Phi_{2}-\Phi_{1}) + \sigma (\phi_1 -\phi_2) + (1-\Phi_2) c_2,\end{aligned}
$$

-  The first difference (`qi$fd`) for the tobit regression model is
   defined as


$$
    \begin{aligned}
        \text{FD}=E(Y\mid x_{1})-E(Y\mid x).\end{aligned}
$$

-  In conditional prediction models, the average expected treatment
   effect (`qi$att.ev`) for the treatment group is


$$
    \begin{aligned}
        \frac{1}{\sum t_{i}}\sum_{i:t_{i}=1}[Y_{i}(t_{i}=1)-E[Y_{i}(t_{i}=0)]],\end{aligned}
$$

   where $t_{i}$ is a binary explanatory variable defining the
   treatment ($t_{i}=1$) and control ($t_{i}=0$) groups.

Output Values
------

The Zelig object stores fields containing everything needed to
rerun the Zelig output, and all the results and simulations as they are generated.
In addition to the summary commands demonstrated above, some simply utility
functions (known as *getters*) provide easy access to the raw fields most
commonly of use for further investigation.

In the example above `z.out$get_coef()` returns the estimated coefficients, `z.out$get_vcov()` returns the estimated covariance matrix, and `z.out$get_predict()` provides predicted values for all observations in the dataset from the analysis.

See also
------

Bayesian tobit regression is part of the MCMCpack library by Andrew D.
Martin and Kevin M. Quinn. The convergence diagnostics are part of the
CODA library by Martyn Plummer, Nicky Best, Kate Cowles, and Karen Vines.

```{r, eval = TRUE, echo=FALSE, results = "asis"}
z5 <- ztobitbayes$new()
z5$references()
```
