---
title: "Comparing Zelig 5 and Zelig 4 Syntax"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Zelig 5 vs Zelig 4}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

*Built using Zelig version `r packageVersion('Zelig')`*

```{r setup, include=FALSE}
knitr::opts_knit$set(
        stop_on_error = 2L
)
```

Zelig 5 uses [reference classes](http://adv-r.had.co.nz/R5.html) which work a bit differently from what you may expect in R. The big difference is that they are "mutable", i.e. assigning values to them does not overwrite the objects previous contents. 

Zelig 5 does contain wrappers (largely) allowing you to use Zelig 4 syntax if you'd like. Here is an example workflow with Zelig 5:

```{r zelig-5-pseudo, eval=FALSE}
z5 <- zls$new()
z5$zelig(Y ~ X1 + X ~ X, weights = w, data = mydata)
z5$setx()
z5$sim()
z5$graph()
```

Here is the same set of operations using the Zelig 4 wrappers:

```{r zelig-4-pseudo, eval=FALSE}
z.out <- zelig(Y ~ X1 + X2, model = "ls", weights = w, data = mydata)
x.out <- setx(z.out)
s.out <- sim(z.out, x = x.out)
plot(s.out)
```

Note that all of the output objects from the Zelig 4 wrappers are Zelig 5 reference class objects, so you can mix and match which syntax you like.