---
title: "Instrumental Variable (2 Stage Least Squares) Regression"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{zelig-ivreg}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

*Built using Zelig version `r packageVersion('Zelig')`*

```{r setup, include=FALSE}
knitr::opts_knit$set(
        stop_on_error = 2L
)
knitr::opts_chunk$set(
    fig.height = 11,
    fig.width = 7
)
```

Instrumental Variable Regression (2 Stage Least Squares) for Continuous
Dependent Variables with `ivreg`.

Syntax
------

With Zelig 5 reference classes:

```{r, eval = FALSE}
z5 <- zls$new()
z5$zelig(Y ~ X1 + X2 | Z1 + Z2 + Z3, weights = w, data = mydata)
z5$setx()
z5$sim()
```

Note that `|` separates the regressors (e.g. `X1` and `X2`) from the instruments
(e.g. `Z1`, `Z2`, `Z3`).

With the Zelig 4 compatibility wrappers:

```{r, eval = FALSE}
z.out <- zelig(Y ~ X1 + X2 | Z1 + Z2 + Z3, model = "ivreg", weights = w,
               data = mydata)
x.out <- setx(z.out)
s.out <- sim(z.out, x = x.out)
```

Examples
--------

Load Zelig and other packages used in the example:

```{r, message=FALSE}
library(Zelig)
library(dplyr) # for the pipe operator %>%
```

Before estimating the model, let's attach the example data and make some
transformations to it. Notably, `zivreg` does not currently support logging
regressors in the `zelig` call. To get around this, we simply find the logged
values before estimating the model:

```{r}
# load and transform data
data("CigarettesSW")
CigarettesSW$rprice <- with(CigarettesSW, price/cpi)
CigarettesSW$rincome <- with(CigarettesSW, income/population/cpi)
CigarettesSW$tdiff <- with(CigarettesSW, (taxs - tax)/cpi)

# log second stage independent variables, as logging internally for ivreg is
# not currently supported
CigarettesSW$log_rprice <- log(CigarettesSW$rprice)
CigarettesSW$log_rincome <- log(CigarettesSW$rincome)
```

Now we can estimate the model and summarize the coefficients:

```{r}
z.out1 <- zelig(log(packs) ~ log_rprice + log_rincome |
                    log_rincome + tdiff + I(tax/cpi),
                    data = CigarettesSW, subset = year == "1995",
                    model = "ivreg")

summary(z.out1)
```

We can then simulate and plot the expected value of the log of the `packs`
response variable across the range of the observed `rprice` regressor values:

```{r}
z.out1 %>% setx(log_rprice = range(log(CigarettesSW$rprice))) %>%
    sim() %>%
    plot()
```

To conduct standard diagnostic tests on the estimated model using functionality
in the [AER](https://CRAN.R-project.org/package=AER) package, exatract the
fitted model object with the `from_zelig_model()` function.


See also
--------

`ivreg` is from Christian Kleiber and Achim Zeileis (2008). Applied Econometrics
with R. New York: Springer-Verlag. ISBN 978-0-387-77316-2. URL <https://CRAN.R-project.org/package=AER>. For more information use
`?AER::ivreg`.

Greene, W. H. (1993) Econometric Analysis, 2nd ed., Macmillan.
