---
title: "Poisson Survey Regression"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{zelig-poissonsurvey}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

*Built using Zelig version `r packageVersion("Zelig")`*

```{r setup, include=FALSE}
knitr::opts_knit$set(
    stop_on_error = 2L
)
knitr::opts_chunk$set(
    fig.height = 11,
    fig.width = 7
)

options(cite = FALSE)
```
---

Poisson Regression for Event Count Dependent Variables with Survey Weights with `poisson.survey`.

Use the Poisson regression model if the observations of your dependent
variable represents the number of independent events that occur during a
fixed period of time (see the [negative binomial model](zelig_negbin.html) for
over-dispersed event counts.) For a Bayesian implementation of this
model, see [Zelig Poisson Bayes](zelig_poissonbayes.html).

Syntax
------

With the Zelig 4 compatibility wrappers:

```{r, eval = FALSE}
z.out <- zelig(Y ~ X1 + X2, model = "poisson.survey",
               weights = w, data = mydata)
x.out <- setx(z.out)
s.out <- sim(z.out, x = x.out)
```

Example
------

```{r, eval = TRUE, echo = FALSE}
rm(list=ls(pattern="\\.out"))
suppressWarnings(suppressMessages(library(Zelig)))
set.seed(1234)
```

### Example 1: User has Existing Sample Weights

Attach sample data:

```{r, eval = TRUE}
data(api, package="survey")
```

In this example, we will estimate a model using
each school's academic performance in 2000 and an
indicator for year-round schools to predict the
number of students who enrolled in each California school.

```{r, eval = TRUE}
z.out1 <- zelig(enroll ~ api99 + yr.rnd ,
                model = "poisson.survey", data = apistrat)
summary(z.out1)
```

Set explanatory variables to their default (mean/mode) values, and set
a high (80th percentile) and low (20th percentile) value for the
measure of academic performance, "api00":

```{r, eval = TRUE}
x.low <- setx(z.out1, api99= quantile(apistrat$api99, 0.2))
x.high <- setx(z.out1, api99= quantile(apistrat$api99, 0.8))
```

Generate first differences for the effect of high versus low "meals"
on the probability that a school will hold classes year round:

```{r, eval = TRUE}
s.out1 <- sim(z.out1, x=x.low, x1=x.high)
summary(s.out1)
```

Generate a second set of fitted values and a plot:

```{r Zelig-poissonsurvey1, dev=c("png", "pdf"), eval = TRUE, fig.cap = "Graphs of Quantities of Interest for Poisson Survey Model"}
plot(s.out1)
```


### Example 2: User has Details about Complex Survey Design (but not sample weights)

Suppose that the survey house that provided
the dataset excluded probability weights
but made other details about the survey
design available.  We can still estimate
a model without probability weights that takes
instead variables that identify each the stratum
and/or cluster from which each observation was
selected and the size of the finite sample from
which each observation was selected.

```{r, eval = TRUE}
z.out2 <- zelig(enroll ~ api99 + yr.rnd ,
                model = "poisson.survey", data = apistrat,
                strata = ~stype, fpc = ~fpc)
summary(z.out2)
```

The coefficient estimates from this model are identical to
point estimates in the previous example, but the standard errors
are smaller.  When sampling weights are omitted, Zelig estimates
them automatically for "normal.survey" models based on the
user-defined description of sampling designs.  In addition,
when user-defined descriptions of the sampling design are
entered as inputs, variance estimates are better and standard
errors are consequently smaller.

The methods setx() and sim() can then be run on z.out2 in the same fashion
described in Example 1.


### Example 3: User has Replicate Weights

Load data for a model using the number of out-of-hospital
cardiac arrests to predict the number of patients who arrive
alive in hospitals.

```{r, eval = TRUE}
data(scd, package="survey")
```

For the purpose of illustration, create four Balanced
Repeated Replicate (BRR) weights:

```{r, eval = TRUE}
BRRrep<-2*cbind(c(1,0,1,0,1,0), c(1,0,0,1,0,1), c(0,1,1,0,0,1), c(0,1,0,1,1,0))
```

Estimate the model using Zelig:

```{r, eval = TRUE}
z.out3 <- zelig(alive ~ arrests , model = "poisson.survey",
                repweights = BRRrep, type = "BRR", data = scd)
summary(z.out3)
```

Set the explanatory variables at their means and set
arrests at its 20th and 80th quartiles

```{r, eval = TRUE}
x.low <- setx(z.out3, arrests = quantile(scd$arrests, .2))
x.high <- setx(z.out3, arrests = quantile(scd$arrests,.8))
```

Generate first differences for the effect of the minimum
versus the maximum number of individuals who arrive
alive on the probability that a hospital will be sued:

```{r, eval = TRUE}
s.out3 <- sim(z.out3, x=x.high, x1=x.low)
summary(s.out3)
```

Generate a second set of fitted values and a plot:

```{r Zelig-poissonsurvey3, dev=c("png", "pdf"), eval = TRUE, fig.cap = "Graphs of Quantities of Interest for Poisson Survey Model"}
plot(s.out3)
```

The user should also refer to the poisson model demo, since
poisson.survey models can take many of the same options as
poisson models.


Model
------

Let $Y_i$ be the number of independent events that occur during a
fixed time period. This variable can take any non-negative integer.

-  The Poisson distribution has *stochastic component*


$$
    Y_i \; \sim \; \textrm{Poisson}(\lambda_i),
$$

   where $\lambda_i$ is the mean and variance parameter.

-  The *systematic component* is


$$
    \lambda_i \; = \; \exp(x_i \beta),
$$

   where $x_i$ is the vector of explanatory variables, and
   $\beta$ is the vector of coefficients.

Quantities of Interest
------

-  The expected value (qi$ev) is the mean of simulations from the
   stochastic component,


$$
    E(Y) = \lambda_i =  \exp(x_i
        \beta),
$$

   given draws of $\beta$ from its sampling distribution.

-  The predicted value (qi$pr) is a random draw from the poisson
   distribution defined by mean $\lambda_i$.

-  The first difference in the expected values (qi$fd) is given by:


$$
    \textrm{FD} \; = \; E(Y | x_1) - E(Y \mid x)
$$

-  In conditional prediction models, the average expected treatment
   effect (att.ev) for the treatment group is


$$
    \frac{1}{\sum_{i=1}^n t_i}\sum_{i:t_i=1}^n \left\{ Y_i(t_i=1) -
        E[Y_i(t_i=0)] \right\},
$$

   where $t_i$ is a binary explanatory variable defining the
   treatment ($t_i=1$) and control ($t_i=0$) groups.
   Variation in the simulations are due to uncertainty in simulating
   $E[Y_i(t_i=0)]$, the counterfactual expected value of
   $Y_i$ for observations in the treatment group, under the
   assumption that everything stays the same except that the treatment
   indicator is switched to $t_i=0$.

-  In conditional prediction models, the average predicted treatment
   effect (att.pr) for the treatment group is


$$
    \frac{1}{\sum_{i=1}^n t_i}\sum_{i:t_i=1}^n \left\{ Y_i(t_i=1) -
        \widehat{Y_i(t_i=0)} \right\},
$$

   where $t_i$ is a binary explanatory variable defining the
   treatment ($t_i=1$) and control ($t_i=0$) groups.
   Variation in the simulations are due to uncertainty in simulating
   $\widehat{Y_i(t_i=0)}$, the counterfactual predicted value of
   $Y_i$ for observations in the treatment group, under the
   assumption that everything stays the same except that the treatment
   indicator is switched to $t_i=0$.

Output Values
------

The Zelig object stores fields containing everything needed to
rerun the Zelig output, and all the results and simulations as they are generated.
In addition to the summary commands demonstrated above, some simply utility
functions (known as *getters*) provide easy access to the raw fields most
commonly of use for further investigation.

In the example above `z.out$get_coef()` returns the estimated coefficients, `z.out$get_vcov()` returns the estimated covariance matrix, and `z.out$get_predict()` provides predicted values for all observations in the dataset from the analysis.

See also
------

The poissonsurvey model is part of the survey package by Thomas Lumley, which in turn depends heavily on `glm` package. Advanced users may
wish to refer to `help(svyglm)` and `help(family)`.

```{r, eval = TRUE, echo=FALSE, results = "asis"}
z5 <- zpoissonsurvey$new()
z5$references()
```
