---
title: "Negative Binomial"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{zelig-negbin}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

*Built using Zelig version `r packageVersion("Zelig")`*

```{r setup, include=FALSE}
knitr::opts_knit$set(
    stop_on_error = 2L
)
knitr::opts_chunk$set(
    fig.height = 11,
    fig.width = 7
)

options(cite = FALSE)
```
---



Negative Binomial Regression for Event Count Dependent Variables with `negbin`.

Use the negative binomial regression if you have a count of events for
each observation of your dependent variable. The negative binomial model
is frequently used to estimate over-dispersed event count models.

Syntax
------

With the Zelig 4 compatibility wrappers:

```{r, eval = FALSE}
z.out <- zelig(Y ~ X1 + X2, model = "negbin", weights = w, data = mydata)
x.out <- setx(z.out)
s.out <- sim(z.out, x = x.out)
```

Example
------

```{r, eval = TRUE, echo = FALSE}
rm(list=ls(pattern="\\.out"))
suppressWarnings(suppressMessages(library(Zelig)))
set.seed(1234)
```

Load sample data:

```{r, eval = TRUE}
data(sanction)
```

Estimate the model:

```{r, eval = TRUE}
z.out <- zelig(num ~ target + coop, model = "negbin", data = sanction)
```

```{r, eval = TRUE}
summary(z.out)
```

Set values for the explanatory variables to their default mean values:

```{r, eval = TRUE}
x.out <- setx(z.out)
```

Simulate fitted values:

```{r, eval = TRUE}
s.out <- sim(z.out, x = x.out)
```

```{r, eval = TRUE}
summary(s.out)
```

```{r Zelig-negbin, dev=c("png", "pdf"), eval = TRUE, fig.cap = "Graphs of Quantities of Interest for Zelig-negbin"}
plot(s.out)
```

Model
------

Let $Y_i$ be the number of independent events that occur during a
fixed time period. This variable can take any non-negative integer
value.

-  The negative binomial distribution is derived by letting the mean of
   the Poisson distribution vary according to a fixed parameter
   $\zeta$ given by the Gamma distribution. The *stochastic
   component* is given by


$$
    \begin{aligned}
        Y_i \mid \zeta_i & \sim & \textrm{Poisson}(\zeta_i \mu_i),\\
           \zeta_i & \sim & \frac{1}{\theta}\textrm{Gamma}(\theta).
         \end{aligned}

   The marginal distribution of $Y_i$ is then the negative
   binomial with mean $\mu_i$ and variance
   $\mu_i + \mu_i^2/\theta$:


$$
    \begin{aligned}
        Y_i & \sim & \textrm{NegBin}(\mu_i, \theta), \\
             & = & \frac{\Gamma (\theta + y_i)}{y! \, \Gamma(\theta)}
                   \frac{\mu_i^{y_i} \, \theta^{\theta}}{(\mu_i + \theta)^{\theta + y_i}},
         \end{aligned}

   where $\theta$ is the systematic parameter of the Gamma
   distribution modeling $\zeta_i$.

-  The *systematic component* is given by


$$
    \mu_i = \exp(x_i \beta)
$$

   where $x_i$ is the vector of $k$ explanatory variables
   and $\beta$ is the vector of coefficients.

Quantities of Interest
------

-  The expected values (qi$ev) are simulations of the mean of the
   stochastic component. Thus,


$$
    E(Y) = \mu_i = \exp(x_i
        \beta),
$$

   given simulations of $\beta$.

-  The predicted value (qi$pr) drawn from the distribution defined by
   the set of parameters $(\mu_i, \theta)$.

-  The first difference (qi$fd) is


$$
    \textrm{FD} \; = \; E(Y | x_1) - E(Y \mid x)
$$

-  In conditional prediction models, the average expected treatment
   effect (att.ev) for the treatment group is


$$
    \frac{1}{\sum_{i=1}^n t_i}\sum_{i:t_i=1}^n \left\{ Y_i(t_i=1) -
        E[Y_i(t_i=0)] \right\},
$$

   where $t_i$ is a binary explanatory variable defining the
   treatment ($t_i=1$) and control ($t_i=0$) groups.
   Variation in the simulations are due to uncertainty in simulating
   $E[Y_i(t_i=0)]$, the counterfactual expected value of
   $Y_i$ for observations in the treatment group, under the
   assumption that everything stays the same except that the treatment
   indicator is switched to $t_i=0$.

-  In conditional prediction models, the average predicted treatment
   effect (att.pr) for the treatment group is


$$
    \frac{1}{\sum_{i=1}^n t_i}\sum_{i:t_i=1}^n \left\{ Y_i(t_i=1) -
        \widehat{Y_i(t_i=0)} \right\},
$$

   where $t_i$ is a binary explanatory variable defining the
   treatment ($t_i=1$) and control ($t_i=0$) groups.
   Variation in the simulations are due to uncertainty in simulating
   $\widehat{Y_i(t_i=0)}$, the counterfactual predicted value of
   $Y_i$ for observations in the treatment group, under the
   assumption that everything stays the same except that the treatment
   indicator is switched to $t_i=0$.

Output Values
------

The Zelig object stores fields containing everything needed to
rerun the Zelig output, and all the results and simulations as they are generated.
In addition to the summary commands demonstrated above, some simply utility
functions (known as *getters*) provide easy access to the raw fields most
commonly of use for further investigation.

In the example above `z.out$get_coef()` returns the estimated coefficients, `z.out$get_vcov()` returns the estimated covariance matrix, and `z.out$get_predict()` provides predicted values for all observations in the dataset from the analysis.

See also
------

The negative binomial model is part of the MASS package by William N.
Venable and Brian D. Ripley . Advanced users may wish to refer to
`help(glm.nb)`.

```{r, eval = TRUE, echo=FALSE, results = "asis"}
z5 <- znegbin$new()
z5$references()
```
