---
title: "Maximum Likelihood Ecological Inference"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{zeligei-eiml}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

*Built using Zelig version `r packageVersion("Zelig")`*

```{r setup, include=FALSE}
knitr::opts_knit$set(
    stop_on_error = 2L
)
knitr::opts_chunk$set(
    fig.height = 11,
    fig.width = 7
)

options(cite = FALSE)
```
---

Ecological Inference Model by Maximum Likelihood with `eiml` in ZeligEI.

Syntax
------

The EI models accept several different formula syntaxes.  If $C1$ and $C2$ are the column totals, and $R1$ and $R2$ are the row totals, and $N=R1_i + R2_i = C1_i + C2_i$ is the total in unit $i$, then the formula can be expressed with just one row and one column, with the totals provided separately as:

```{r, eval = FALSE}
z.out <- zelig(C1 ~ R1, N = N, data = data)
```

The argument `N` can be either a numeric vector of the total in each *i*-th unit, or the character name of a variable in the dataset that contains these values.

Or with both rows and columns coupled together, and omitting the totals:

```{r, eval = FALSE}
z.out <- zelig(cbind(C1,C2) ~ cbind(R1,R2), data = data)
```

Additionally, if `C1`, `C2`, `R1`, `R2` are percentages rather than counts, then either formula method above is acceptable, however, `N` must always be provided.

First load packages:

```{r, eval = TRUE, message=FALSE}
library(zeligverse)
```

Here is an example of all the syntax for the analysis using the first syntax method, and the direct use of the reference classes:

```{r, eval = FALSE}
z5 <- zeiml$new()
z5$zelig(C1 ~ R1, N = myN, weights = w, data = myData)
z5$setx()
z5$sim()
```

With the Zelig 4 compatibility wrappers this looks like:

```{r, eval = FALSE}
z.out <- zelig(C1 ~ R1, N=N, model = "eiml", weights = w, data = myData)
x.out <- setx(z.out)
s.out <- sim(z.out, x = x.out)
```

Examples
------

```{r, eval = TRUE, echo = FALSE}
rm(list=ls(pattern="\\.out"))
set.seed(1234)
```

We'll use a dataset from the `ei` package, of black and non-black turnout in 141 precincts.

```{r, eval = TRUE}
library("ei", quietly=TRUE)
data(sample)
```

Here is the model estimated in Zelig.
```{r, eval = TRUE}
z.out <- zeiml$new()
z.out$zelig(t ~ x, N = "n", data = sample)
summary(z.out)
```


See also
------

This model is part of the ei package by Gary King and Molly Roberts. Advanced users may wish to refer to
`help(ei)` in the ei package.

```{r, eval = TRUE, echo=FALSE, results = "asis"}

z5 <- zeiml$new()
z5$references()
```
