---
title: "Mutlinomial Logit"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{zeligchoice-mlogit}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

*Built using Zelig version `r packageVersion("Zelig")`*

```{r setup, include=FALSE}
knitr::opts_knit$set(
    stop_on_error = 2L
)
knitr::opts_chunk$set(
    fig.height = 11,
    fig.width = 7
)
```
---

Multinomial Logistic Regression for Dependent Variables with Unordered Categorical Values with `mlogit` in ZeligChoice.

Use the multinomial logit distribution to model unordered categorical
variables. The dependent variable may be in the format of either
character strings or integer values. See [Multinomial Bayesian Logistic Regression](zelig_mlogitbayes.html) for a Bayesian version of this
model.

Syntax
------

First load packages:

```{r, eval = TRUE, message=FALSE}
library(zeligverse)
```

With the Zelig 4 compatibility wrappers:

```{r, eval = FALSE}
z.out <- zelig(as.factor(Y) ~ X1 + X23,
               model = "mlogit", data = mydata)
x.out <- setx(z.out)
s.out <- sim(z.out, x = x.out, x1 = NULL)

```

where Y above is supposed to be a factor variable with levels
apples,bananas,oranges. By default, oranges is the last level and
omitted. (You cannot specify a different base level at this time.) For
$J$ equations, there must be $J + 1$ levels.

Examples
------

```{r, eval = TRUE, echo = FALSE}
rm(list=ls(pattern="\\.out"))
set.seed(1234)
```

Load the sample data:

```{r, eval = TRUE}
data(mexico)
```

Estimate the empirical model:

```{r, eval = TRUE}
z.out1 <- zelig(as.factor(vote88) ~ pristr + othcok + othsocok,
                model = "mlogit", data = mexico, cite = FALSE)
```

Summarize estimated paramters:

```{r, eval = TRUE}
summary(z.out1)
```

Set the explanatory variables to their default values, with $pristr$
(for the strength of the PRI) equal to 1 (weak) in the baseline
values, and equal to 3 (strong) in the alternative values:

```{r, eval = TRUE}
x.weak <- setx(z.out1, pristr = 1)
x.strong <- setx(z.out1, pristr = 3)
```

Generate simulated predicted probabilities `qi$ev` and differences in the predicted probabilities `qi$fd`:

```{r, eval = TRUE}
s.out.mlogit <- sim(z.out1, x = x.strong, x1 = x.weak)
summary(s.out.mlogit)
```

```{r Zelig-mlogit1, dev=c("png", "pdf"), eval = TRUE, fig.cap = "Graphs of Quantities of Interest for Multinomial Logit"}
plot(s.out.mlogit)
```

Model
------

Let $Y_i$ be the unordered categorical dependent variable that
takes one of the values from 1 to $J$, where $J$ is the
total number of categories.

-  The stochastic component is given by


$$
    Y_i \; \sim \; \textrm{Multinomial}(y_{i} \mid \pi_{ij}),
$$

   where $\pi_{ij}=\Pr(Y_i=j)$ for $j=1,\dots,J$.

-  The systemic component is given by:


$$
    \pi_{ij}\; = \; \frac{\exp(x_{i}\beta_{j})}{\sum^{J}_{k = 1}
        \exp(x_{i}\beta_{k})},
$$

   where $x_i$ is the vector of explanatory variables for
   observation $i$, and $\beta_j$ is the vector of
   coefficients for category $j$.

Quantities of Interest
------

-  The expected value (qi$ev) is the predicted probability for each
   category:


$$
    E(Y) \; = \; \pi_{ij}\; = \; \frac{\exp(x_{i}\beta_{j})}{\sum^{J}_{k = 1}
        \exp(x_{i}\beta_{k})}.
$$

-  The predicted value (qi$pr) is a draw from the multinomial
   distribution defined by the predicted probabilities.

-  The first difference in predicted probabilities (qi$fd), for each
   category is given by:


$$
    \textrm{FD}_j = \Pr(Y=j \mid x_1) - \Pr(Y=j \mid x) \quad {\rm for}
        \quad j=1,\dots,J.
$$

-  In conditional prediction models, the average expected treatment
   effect (att.ev) for the treatment group is


$$
    \frac{1}{n_j}\sum_{i:t_i=1}^{n_j} \left\{ Y_i(t_i=1) -
        E[Y_i(t_i=0)] \right\},
$$

   where $t_{i}$ is a binary explanatory variable defining the
   treatment ($t_{i}=1$) and control ($t_{i}=0$) groups, and
   $n_j$ is the number of treated observations in category
   $j$.

-  In conditional prediction models, the average predicted treatment
   effect (att.pr) for the treatment group is


$$
    \frac{1}{n_j}\sum_{i:t_i=1}^{n_j} \left\{ Y_i(t_i=1) -
        \widehat{Y_i(t_i=0)} \right\},
$$

   where $t_{i}$ is a binary explanatory variable defining the
   treatment ($t_{i}=1$) and control ($t_{i}=0$) groups, and
   $n_j$ is the number of treated observations in category
   $j$.

Output Values
------

The Zelig object stores fields containing everything needed to
rerun the Zelig output, and all the results and simulations as they are
generated. In addition to the summary commands demonstrated above, some simply
utility functions (known as *[getters](getters.html)*) provide easy access to
the raw fields most commonly of use for further investigation.

If the `zelig()` call output object is `z.out`, then `coef(z.out)` returns
the estimated coefficients, `vcov(z.out)` returns the estimated covariance
matrix, and `predict(z.out)` provides predicted values for all observations
in the dataset from the analysis.

See also
------

The multinomial logit function is part of the VGAM package by Thomas
Yee. In addition, advanced users may wish to refer to `help(vglm)`
in the VGAM library.

```{r, eval = TRUE, echo=FALSE, results = "asis"}
z5 <- zmlogit$new()
z5$references()
```
