---
title: "Ordered Logistic Regression"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{zeligchoice-ologit}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

*Built using Zelig version `r packageVersion("Zelig")`*

```{r setup, include=FALSE}
knitr::opts_knit$set(
    stop_on_error = 2L
)
knitr::opts_chunk$set(
    fig.height = 11,
    fig.width = 7
)

options(cite = FALSE)
```
---

Ordinal Logistic Regression for Ordered Categorical Dependent Variables with `ologit` in ZeligChoice.

Use the ordinal logit regression model if your dependent variable is
ordered and categorical, either in the form of integer values or
character strings.

Syntax
------

First load packages:

```{r, eval = TRUE, message=FALSE}
library(zeligverse)
```

With the Zelig 4 compatibility wrappers:

```{r, eval = FALSE}
z.out <- zelig(as.factor(Y) ~ X1 + X23,
               model = "ologit", data = mydata)
x.out <- setx(z.out)
s.out <- sim(z.out, x = x.out, x1 = NULL)
```

If `Y` takes discrete integer values, the `as.factor()` command will order
automatically order the values. If Y takes on values composed of
character strings, such as “strongly agree”, “agree”, and “disagree”,
`as.factor()` will order the values in the order in which they appear in
`Y`. You will need to replace your dependent variable with a factored
variable prior to estimating the model through `zelig()`. See below for more details.

Example
------

```{r, eval = TRUE, echo = FALSE}
rm(list=ls(pattern="\\.out"))
set.seed(1234)
```

### Creating An Ordered Dependent Variable

Load the sample data:

```{r, eval = TRUE}
data(sanction)
```

Create an ordered dependent variable:

```{r, eval = TRUE}
sanction$ncost <- factor(sanction$ncost, ordered = TRUE,
                         levels = c("net gain", "little effect", "modest loss", "major loss"))
```

Estimate the model:

```{r, eval = TRUE}
z.out <- zelig(ncost ~ mil + coop, model = "ologit",
               data = sanction)
```

Summarize estimated paramters:

```{r, eval = TRUE}
summary(z.out)
```

Set the explanatory variables to their observed values:

```{r, eval = TRUE}
x.out <- setx(z.out)
```

Simulate fitted values given x.out and view the results:

```{r, eval = TRUE}
s.out <- sim(z.out, x = x.out)
```

```{r, eval = TRUE}
summary(s.out)
```

```{r Zelig-ologit1, dev=c("png", "pdf"), eval = TRUE, fig.cap = "Graphs of Quantities of Interest for Ordered Logit"}
plot(s.out)
```

### First Differences

Using the sample data `sanction`, estimate the empirical model and returning the coefficients:

```{r, eval = TRUE}
z.out <- zelig(as.factor(cost) ~ mil + coop, model = "ologit",
               data = sanction)
```

```{r, eval = TRUE}
summary(z.out)
```

Set the explanatory variables to their means, with coop set to 1 (the lowest value) in the baseline case and set
to 4 (the highest value) in the alternative case:

```{r, eval = TRUE}
x.low <- setx(z.out, coop = 1)
x.high <- setx(z.out, coop = 4)
```

Generate simulated fitted values and first differences, and view the results:

```{r, eval = TRUE}
s.out2 <- sim(z.out, x = x.low, x1 = x.high)
summary(s.out2)
```

```{r Zelig-ologit2, dev=c("png", "pdf"), eval = TRUE, fig.cap = "Graphs of Quantities of Interest for Ordered Logit"}
plot(s.out2)
```

Model
------

Let $Y_i$ be the ordered categorical dependent variable for
observation $i$ that takes one of the integer values from
$1$ to $J$ where $J$ is the total number of
categories.

-  The *stochastic component* begins with an unobserved continuous
   variable, $Y^*_i$, which follows the standard logistic
   distribution with a parameter $\mu_i$,


$$
    Y_i^* \; \sim \; \textrm{Logit}(y_i^* \mid \mu_i),
$$

   to which we add an observation mechanism


$$
    Y_i \; = \; j \quad {\rm if} \quad \tau_{j-1} \le Y_i^* \le \tau_j
        \quad {\rm for} \quad j=1,\dots,J.
$$

   where $\tau_l$ (for $l=0,\dots,J$) are the threshold
   parameters with [\tau_l]( \tau_m[ for all $l](m$ and
   $\tau_0=-\infty$ and $\tau_J=\infty$.

-  The *systematic component* has the following form, given the
   parameters $\tau_j$ and $\beta$, and the explanatory
   variables $x_i$:


$$
    \Pr(Y \le j) \; = \; \Pr(Y^* \le \tau_j) \; = \frac{\exp(\tau_j -
        x_i \beta)}{1+\exp(\tau_j -x_i \beta)},
$$

   which implies:


$$
    \pi_{j}  \; = \; \frac{\exp(\tau_j - x_i \beta)}{1 + \exp(\tau_j -
        x_i \beta)} - \frac{\exp(\tau_{j-1} - x_i \beta)}{1 +
            \exp(\tau_{j-1} - x_i \beta)}.
$$

Quantities of Interest
------

-  The expected values (qi$ev) for the ordinal logit model are
   simulations of the predicted probabilities for each category:


$$
    E(Y = j) \; = \; \pi_{j} \; = \; \frac{\exp(\tau_j - x_i \beta)}
        {1 + \exp(\tau_j - x_i \beta)} - \frac{\exp(\tau_{j-1} - x_i \beta)}{1 +
       \exp(\tau_{j-1} - x_i \beta)},
$$

   given a draw of $\beta$ from its sampling distribution.

-  The predicted value (qi$pr) is drawn from the logit distribution
   described by $\mu_i$, and observed as one of $J$ discrete
   outcomes.

-  The difference in each of the predicted probabilities (qi$fd) is
   given by


$$
    \Pr(Y=j \mid x_1) \;-\; \Pr(Y=j \mid x) \quad {\rm for} \quad
        j=1,\dots,J.
$$

-  In conditional prediction models, the average expected treatment
   effect (att.ev) for the treatment group is


$$
    \frac{1}{n_j}\sum_{i:t_i=1}^{n_j} \left\{ Y_i(t_i=1) -
        E[Y_i(t_i=0)] \right\},
$$

   where $t_{i}$ is a binary explanatory variable defining the
   treatment ($t_{i}=1$) and control ($t_{i}=0$) groups, and
   $n_j$ is the number of treated observations in category
   $j$.

-  In conditional prediction models, the average predicted treatment
   effect (att.pr) for the treatment group is


$$
    \frac{1}{n_j}\sum_{i:t_i=1}^{n_j} \left\{ Y_i(t_i=1) -
        \widehat{Y_i(t_i=0)} \right\},
$$

   where $t_{i}$ is a binary explanatory variable defining the
   treatment ($t_{i}=1$) and control ($t_{i}=0$) groups, and
   $n_j$ is the number of treated observations in category
   $j$.

Output Values
------

The output of each Zelig command contains useful information which you
may view. For example, if you run
`z.out <- zelig(y ~ x, model = ologit, data)`, then you may examine
the available information in `z.out` by using `names(z.out)`, see
the coefficients by using z.out$coefficients, and a default summary of
information through `summary(z.out)`. Other elements available through
the $ operator are listed below.

-  From the zelig() output object z.out, you may extract:

   -  coefficients: parameter estimates for the explanatory variables.

   -  zeta: a vector containing the estimated class boundaries
      $\tau_j$.

   -  deviance: the residual deviance.

   -  fitted.values: the $n \times J$ matrix of in-sample fitted
      values.

   -  df.residual: the residual degrees of freedom.

   -  edf: the effective degrees of freedom.

   -  Hessian: the Hessian matrix.

   -  zelig.data: the input data frame if save.data = TRUE.

-  From summary(z.out), you may extract:

   -  coefficients: the parameter estimates with their associated
      standard errors, and $t$-statistics.

See also
------

The ordinal logit model is part of the MASS package by William N.
Venable and Brian D. Ripley. Advanced users may wish to refer to
`help(polr)`.

```{r, eval = TRUE, echo=FALSE, results = "asis"}
z5 <- zologit$new()
z5$references()
```
