---
title: "Multinomial Dirichlet model for Ecological Inference"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{zeligei-eirxc}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

*Built using Zelig version `r packageVersion("Zelig")`*

```{r setup, include=FALSE}
knitr::opts_knit$set(
    stop_on_error = 2L
)
knitr::opts_chunk$set(
    fig.height = 11,
    fig.width = 7
)

options(cite = FALSE)
```
---

Multinomial Dirichlet model for Ecological Inference in RxC tables with `rxc` using ZeligEI.

Syntax
------

Unlike other EI models in Zelig, in the $RxC$ case, all the row and column variables need to specified.  Let's assume $C1$ through $CC$ are the column totals, and $R1$ through $RR$ are the row totals, and $N=R1_i + R2_i + \ldots + RR_i = C1_i + C2_i + \ldots + CC_i$ is the total in unit $i$.  In the case with three row variables and three column variables, the syntax is:

```{r, eval = FALSE}
z.out <- zelig(cbind(C1,C2,C3) ~ cbind(R1,R2,R3), data=data)
```

Additionally, if `C1`, `C2`, `R1`, `R2` are percentages rather than counts, then you must also provide `N` the unit totals as:

```{r, eval = FALSE}
z.out <- zelig( cbind(C1,C2,C3) ~ cbind(R1,R2,R3), N=N, data=data)
```

The argument `N` can be either a numeric vector of the total in each *i*-th unit, or the character name of a variable in the dataset that contains these values.

First load packages:

```{r, eval = TRUE, message=FALSE}
library(zeligverse)
```

Here is an example of all the syntax for the analysis using the first syntax method, and the direct use of the reference classes:

```{r, eval = FALSE}
z5 <- zeirxc$new()
z5$zelig(cbind(C1, C2, C3) ~ cbind(R1, R2, R3), N = myN,
        weights = w, data = myData)
z5$setx()
z5$sim()
```

With the Zelig 4 compatibility wrappers this looks like:

```{r, eval = FALSE}
z.out <- zelig(cbind(C1, C2, C3) ~ cbind(R1, R2, R3), N = myN,
               model = "rxc", weights = w, data = myData)
x.out <- setx(z.out)
s.out <- sim(z.out, x = x.out)
```

Examples
------

```{r, eval = TRUE, echo = FALSE}
rm(list=ls(pattern="\\.out"))
set.seed(1234)
```

We'll use a dataset from the `eiPack` package, of registration data for White, Black, and Native American voters, and party voteshare, in 277 precincts in eight counties of south-eastern North Carolina in 2001.
```{r, eval = TRUE}
library("eiPack", quietly=TRUE)
data(senc)
```

Here is the model estimated in Zelig.
```{r, eval = TRUE}
z.out <- zeirxc$new()
z.out$zelig(cbind(dem, rep, non) ~ cbind(black, white, natam),
            N = "total", data = senc)
summary(z.out)
```

You can check for convergence before summarizing the estimates with three diagnostic tests.  See the section [Diagnostics for Zelig Models](diagnostics.html) for more examples of these tests, and interpretation of their meaning.  For brevity we only show the first set of parameters, the $\alpha$'s, as there are twice as many $\beta$'s as there are precincts:

```{r, eval = TRUE}
gd <- z.out$geweke.diag()
print(gd[[1]])
hd <- z.out$heidel.diag()
print(hd[[1]])
rd <-z.out$raftery.diag()
print(rd[[1]])
```



See also
------

This model is part of the eiPack package by Olivia Lau, Ryan T. Moore and Michael Kellerman. Advanced users may wish to refer to
`help(ei.MD.bayes)` in the eiPack package.

```{r, eval = TRUE, echo=FALSE, results = "asis"}
z5 <- zeirxc$new()
z5$references()
```
