---
title: "Bivariate Probit"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{zeligchoice-bprobit}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

*Built using Zelig version `r packageVersion("Zelig")`*

```{r setup, include=FALSE}
knitr::opts_knit$set(
    stop_on_error = 2L
)
knitr::opts_chunk$set(
    fig.height = 11,
    fig.width = 7
)

options(cite = FALSE)
```
---

Bivariate Probit Regression for Two Dichotomous Dependent Variables with `bprobit` from ZeligChoice.

Use the bivariate probit regression model if you have two binary
dependent variables $(Y_1, Y_2)$, and wish to model them jointly
as a function of some explanatory variables. Each pair of dependent
variables $(Y_{i1}, Y_{i2})$ has four potential outcomes,
$(Y_{i1}=1,
Y_{i2}=1)$, $(Y_{i1}=1, Y_{i2}=0)$, $(Y_{i1}=0, Y_{i2}=1)$,
and $(Y_{i1}=0, Y_{i2}=0)$. The joint probability for each of
these four outcomes is modeled with three systematic components: the
marginal Pr\ $(Y_{i1} = 1)$ and Pr\ $(Y_{i2} = 1)$, and the
correlation parameter $\rho$ for the two marginal distributions.
Each of these systematic components may be modeled as functions of
(possibly different) sets of explanatory variables.

Syntax
------

First load packages:

```{r, eval = TRUE, message=FALSE}
library(zeligverse)
```

With the Zelig 4 compatibility wrappers:

```{r, eval = FALSE}
z.out <- zelig(cbind(Y1, Y2) ~ X1 + X2 + X3,
               model = "bprobit", data = mydata)
x.out <- setx(z.out)
s.out <- sim(z.out, x = x.out, x1 = NULL)
```

Input Values
------

In every bivariate probit specification, there are three equations which
correspond to each dependent variable ($Y_1$, $Y_2$), and
the correlation parameter $\rho$. Since the correlation parameter
does not correspond to one of the dependent variables, the model
estimates $\rho$ as a constant by default. Hence, only two
formulas (for $\mu_1$ and $\mu_2$) are required. If the
explanatory variables for $\mu_1$ and $\mu_2$ are the same
and effects are estimated separately for each parameter, you may use the
following short hand:

```{r, eval = FALSE}
fml <- list(cbind(Y1,Y2) ~ X1 + X2)
```

which has the same meaning as:

```{r, eval = FALSE}
fml <- list(mu1 = Y1 ~ X1 + X2, + mu2 = Y2 ~ X1 + X2)
```

**Anticipated feature, not currently enabled:**

You may use the function `tag()` to constrain variables across equations.
The `tag()` function takes a variable and a label for the effect
parameter. Below, the constrained effect of x3 in both equations is
called the age parameter:

```{r, eval = FALSE}
fml <- list(mu1 = y1 ~ x1 + tag(x3, “age”), + mu2 = y2 ~ x2 +
tag(x3, “age”))
```

You may also constrain different variables across different equations to
have the same effect.

Examples
------

```{r, eval = TRUE, echo = FALSE}
rm(list=ls(pattern="\\.out"))
set.seed(1234)
```

### Basic Example [basic.bp]

Load the data and estimate the model:

```{r, eval = TRUE}
data(sanction)
```

```{r, eval = TRUE}
z.out1 <- zelig(cbind(import, export) ~ coop + cost + target,
                model = "bprobit", data = sanction)
summary(z.out1)
```

By default, `zelig()` estimates two effect parameters for each
explanatory variable in addition to the correlation coefficient; this
formulation is parametrically independent (estimating unconstrained
effects for each explanatory variable), but stochastically dependent
because the models share a correlation parameter. Generate baseline
values for the explanatory variables (with cost set to 1, net gain to
sender) and alternative values (with cost set to 4, major loss to
sender):

```{r, eval = TRUE}
x.low <- setx(z.out1, cost = 1)
x.high <- setx(z.out1, cost = 4)
```

Simulate fitted values and first differences:

```{r, eval = TRUE}
s.out1 <- sim(z.out1, x = x.low, x1 = x.high)
```

```{r, eval = TRUE}
summary(s.out1)
```

```{r Zelig-bprobit1, dev=c("png", "pdf"), eval = TRUE, fig.cap = "Graphs of Quantities of Interest for Bivariate Probit"}
plot(s.out1)
```



Model
------

For each observation, define two binary dependent variables, $Y_1$
and $Y_2$, each of which take the value of either 0 or 1 (in the
following, we suppress the observation index $i$). We model the
joint outcome $(Y_1$, $Y_2)$ using two marginal
probabilities for each dependent variable, and the correlation
parameter, which describes how the two dependent variables are related.

-  The *stochastic component* is described by two latent (unobserved)
   continuous variables which follow the bivariate Normal distribution:


$$
    \begin{aligned}
        \left ( \begin{array}{c}
            Y_1^* \\
            Y_2^*
          \end{array}
        \right ) &\sim &
        N_2 \left \{ \left (
            \begin{array}{c}
              \mu_1 \\ \mu_2
            \end{array} \right ), \left( \begin{array}{cc}
                       1 & \rho \\
                       \rho & 1
                       \end{array} \right) \right\},
    \end{aligned}
$$

   where $\mu_j$ is a mean for $Y_j^*$ and $\rho$ is a
   scalar correlation parameter. The following observation mechanism
   links the observed dependent variables, $Y_j$, with these
   latent variables


$$
    \begin{aligned}
        Y_j & = & \left \{ \begin{array}{cc}
                         1 & {\rm if} \; Y_j^* \ge 0, \\
                         0 & {\rm otherwise.}
                         \end{array}
                         \right.
    \end{aligned}
$$

-  The *systemic components* for each observation are


$$
    \begin{aligned}
        \mu_j & = & x_{j} \beta_j \quad {\rm for} \quad j=1,2, \\
          \rho & = & \frac{\exp(x_3 \beta_3) - 1}{\exp(x_3 \beta_3) + 1}.
    \end{aligned}
$$

Quantities of Interest
------

For $n$ simulations, expected values form an $n \times 4$
matrix.

-  The expected values (`qi$ev`) for the binomial probit model are the
   predicted joint probabilities. Simulations of $\beta_1$,
   $\beta_2$, and $\beta_3$ (drawn form their sampling
   distributions) are substituted into the systematic components, to
   find simulations of the predicted joint probabilities
   $\pi_{rs}=\Pr(Y_1=r, Y_2=s)$:


$$
    \begin{aligned}
        \pi_{11} &= \Pr(Y_1^* \geq 0 , Y_2^* \geq 0) &= \int_0^{\infty}
      \int_0^{\infty} \phi_2 (\mu_1, \mu_2, \rho) \, dY_2^*\, dY_1^* \\
      \pi_{10} &= \Pr(Y_1^* \geq 0 , Y_2^* < 0)  &= \int_0^{\infty}
      \int_{-\infty}^{0} \phi_2 (\mu_1, \mu_2, \rho) \, dY_2^*\, dY_1^*\\
      \pi_{01} &= \Pr(Y_1^* < 0 , Y_2^* \geq 0)  &= \int_{-\infty}^{0}
      \int_0^{\infty} \phi_2 (\mu_1, \mu_2, \rho) \, dY_2^*\, dY_1^*\\
      \pi_{11} &= \Pr(Y_1^* < 0 , Y_2^* < 0)  &= \int_{-\infty}^{0}
      \int_{-\infty}^{0} \phi_2 (\mu_1, \mu_2, \rho) \, dY_2^*\, dY_1^*\\
    \end{aligned}
$$

   where $r$ and $s$ may take a value of either 0 or 1,
   $\phi_2$ is the bivariate Normal density.

-  The predicted values (`qi$pr`) are draws from the multinomial
   distribution given the expected joint probabilities.

-  The first difference (`qi$fd`) in each of the predicted joint
   probabilities are given by


$$
    \textrm{FD}_{rs} = \Pr(Y_1=r, Y_2=s \mid x_1)-\Pr(Y_1=r, Y_2=s
        \mid x).
$$

-  The risk ratio (qi$rr) for each of the predicted joint probabilities
   are given by


$$
    \textrm{RR}_{rs} = \frac{\Pr(Y_1=r, Y_2=s \mid x_1)}{\Pr(Y_1=r, Y_2=s \mid x)}.
$$

-  In conditional prediction models, the average expected treatment
   effect (att.ev) for the treatment group is


$$
    \frac{1}{\sum_{i=1}^n t_i}\sum_{i:t_i=1}^n \left\{ Y_{ij}(t_i=1) -
        E[Y_{ij}(t_i=0)] \right\} \textrm{ for } j = 1,2,
$$

   where $t_i$ is a binary explanatory variable defining the
   treatment ($t_i=1$) and control ($t_i=0$) groups.
   Variation in the simulations are due to uncertainty in simulating
   $E[Y_{ij}(t_i=0)]$, the counterfactual expected value of
   $Y_{ij}$ for observations in the treatment group, under the
   assumption that everything stays the same except that the treatment
   indicator is switched to $t_i=0$.

-  In conditional prediction models, the average predicted treatment
   effect (att.pr) for the treatment group is


$$
    \frac{1}{\sum_{i=1}^n t_i}\sum_{i:t_i=1}^n \left\{ Y_{ij}(t_i=1) -
        \widehat{Y_{ij}(t_i=0)}\right\} \textrm{ for } j = 1,2,
$$

   where $t_i$ is a binary explanatory variable defining the
   treatment ($t_i=1$) and control ($t_i=0$) groups.
   Variation in the simulations are due to uncertainty in simulating
   $\widehat{Y_{ij}(t_i=0)}$, the counterfactual predicted value
   of $Y_{ij}$ for observations in the treatment group, under the
   assumption that everything stays the same except that the treatment
   indicator is switched to $t_i=0$.

Output Values
------

The output of each Zelig command contains useful information which you
may view. For example, if you run
`z.out <- zelig(y ~ x, model = bprobit, data)`, then you may examine
the available information in `z.out` by using `names(z.out)`, see
the coefficients by using z.out$coefficients, and obtain a default
summary of information through `summary(z.out)`. Other elements
available through the $ operator are listed below.

-  From the `zelig()` output object z.out, you may extract:

   -  coefficients: the named vector of coefficients.

   -  fitted.values: an $n \times 4$ matrix of the in-sample
      fitted values.

   -  predictors: an $n \times 3$ matrix of the linear predictors
      $x_j \beta_j$.

   -  residuals: an $n \times 3$ matrix of the residuals.

   -  df.residual: the residual degrees of freedom.

   -  df.total: the total degrees of freedom.

   -  rss: the residual sum of squares.

   -  y: an $n \times 2$ matrix of the dependent variables.

   -  zelig.data: the input data frame if `save.data = TRUE`.

-  From `summary(z.out)`, you may extract:

   -  `coef3`: a table of the coefficients with their associated standard
      errors and $t$-statistics.

   -  `cov.unscaled`: the variance-covariance matrix.

   -  `pearson.resid`: an $n \times 3$ matrix of the Pearson
      residuals.

See also
------

The bivariate probit function is part of the VGAM package by Thomas
Yee. In addition, advanced users may wish to refer to `help(vglm)`
in the VGAM library.

```{r, eval = TRUE, echo=FALSE, results = "asis"}
z5 <- zbprobit$new()
z5$references()
```
