---
title: "Zelig-exp"
output:
  html_document:
    number_sections: yes
    toc: yes
bibliography: zelig5.bib
---

<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Model: Zelig-exp}
-->

# Exponential Regression for Duration Dependent Variables

Use the exponential duration regression model if you have a dependent
variable representing a duration (time until an event). The model
assumes a constant hazard rate for all events. The dependent variable
may be censored (for observations have not yet been completed when data
were collected).

## Syntax

```{r, eval = FALSE}
z.out <- zelig(Surv(Y, C) ~ X, model = "exp", data = mydata)
x.out <- setx(z.out)
s.out <- sim(z.out, x = x.out)
```

Exponential models require that the dependent variable be in the form
`Surv(Y, C)`, where `Y` and `C` are
vectors of length $n$. For each observation $i$ in 1, …, $n$, the value
$y_i$ is the duration (lifetime, for example), and the associated $c_i$
is a binary variable such that $c_i = 1$ if the duration is not censored
(`*e.g.*`, the subject dies during the study) or $c_i = 0$ if
the duration is censored (`*e.g.*`, the subject is still
alive at the end of the study and is know to live at least as long as
$y_i$). If $c_i$ is omitted, all Y are assumed to be completed; that is,
time defaults to 1 for all observations.

## Input Values

In addition to the standard inputs, `zelig()` takes the
following additional options for exponential regression:

-   `robust`: defaults to `FALSE`. If
    `TRUE`, `zelig()` computes robust standard
    errors based on sandwich estimators (see @Huber81 and @White80) and
    the options selected in `cluster`.

-   `cluster`: if `robust = TRUE`, you may select
    a variable to define groups of correlated observations. Let
    `x3` be a variable that consists of either discrete
    numeric values, character strings, or factors that define strata.
    Then

        > z.out <- zelig(y ~ x1 + x2, robust = TRUE, cluster = "x3", 
                         model = "exp", data = mydata)

    means that the observations can be correlated within the strata
    defined by the variable `x3`, and that robust standard
    errors should be calculated according to those clusters. If
    `robust = TRUE` but `cluster` is not
    specified, `zelig()` assumes that each observation falls
    into its own cluster.

## Example

Attach the sample data:

```{r echo = FALSE, warning = FALSE, message = FALSE}
suppressPackageStartupMessages(library(Zelig5, quietly = TRUE))
data(macro)
```

```{r}
data(coalition)
```

Estimate the model:

```{r}
z.out <- zelig(Surv(duration, ciep12) ~ fract + numst2,
               model = "exp", data = coalition)
```

View the regression output:

```{r}
summary(z.out)
```

Set the baseline values (with the ruling coalition in the minority) and
the alternative values (with the ruling coalition in the majority) for
X:

```{r}
x.low <- setx(z.out, numst2 = 0)
x.high <- setx(z.out, numst2 = 1)
```


Simulate expected values (`qi$ev`) and first differences
(`qi$fd`):

```{r}
s.out <- sim(z.out, x = x.low, x1 = x.high)
```

Summarize quantities of interest and produce some plots:

```{r}
summary(s.out)
# plot(s.out)
```

## Model

Let $Y_i^*$ be the survival time for observation $i$. This variable
might be censored for some observations at a fixed time $y_c$ such that
the fully observed dependent variable, $Y_i$, is defined as

$$Y_i = \left\{ \begin{array}{ll}
      Y_i^* & \textrm{if }Y_i^* \leq y_c \\
      y_c & \textrm{if }Y_i^* > y_c \\
    \end{array} \right.$$

-   The *stochastic component* is described by the distribution of the
    partially observed variable $Y^*$. We assume $Y_i^*$ follows the
    exponential distribution whose density function is given by

    $$f(y_i^*\mid \lambda_i) = \frac{1}{\lambda_i} \exp\left(-\frac{y_i^*}{\lambda_i}\right)$$

    for $y_i^*\ge 0$ and $\lambda_i>0$. The mean of this distribution is
    $\lambda_i$.

    In addition, survival models like the exponential have three
    additional properties. The hazard function $h(t)$ measures the
    probability of not surviving past time $t$ given survival up to $t$.
    In general, the hazard function is equal to $f(t)/S(t)$ where the
    survival function $S(t) = 1 - \int_{0}^t f(s) ds$ represents the
    fraction still surviving at time $t$. The cumulative hazard function
    $H(t)$ describes the probability of dying before time $t$. In
    general, $H(t)= \int_{0}^{t} h(s) ds = -\log S(t)$. In the case of
    the exponential model,

    $$\begin{aligned}
    h(t) &=& \frac{1}{\lambda_i} \\
    S(t) &=& \exp\left( -\frac{t}{\lambda_i} \right) \\
    H(t) &=& \frac{t}{\lambda_i}\end{aligned}$$

    For the exponential model, the hazard function $h(t)$ is constant
    over time. The Weibull model and lognormal models allow the hazard
    function to vary as a function of elapsed time (see and
    respectively).

-   The *systematic component* $\lambda_i$ is modeled as

    $$\lambda_i = \exp(x_i \beta),$$

    where $x_i$ is the vector of explanatory variables, and $\beta$ is
    the vector of coefficients.

## Quantities of Interest

-   The expected values (`qi$ev`) for the exponential model
    are simulations of the expected duration given $x_i$ and draws of
    $\beta$ from its posterior, $$E(Y) = \lambda_i = \exp(x_i \beta).$$

-   The predicted values (`qi$pr`) are draws from the
    exponential distribution with rate equal to the expected value.

-   The first difference (or difference in expected values,
    `qi$ev.diff`), is

    $$\textrm{FD} \; = \; E(Y \mid x_1) - E(Y \mid x),$$

    where $x$ and $x_1$ are different vectors of values for the
    explanatory variables.

-   In conditional prediction models, the average expected treatment
    effect (`att.ev`) for the treatment group is

    $$\frac{1}{\sum_{i=1}^n t_i}\sum_{i:t_i=1}^n \left\{ Y_i(t_i=1) - E[Y_i(t_i=0)]
      \right\},$$

    where $t_i$ is a binary explanatory variable defining the treatment
    ($t_i=1$) and control ($t_i=0$) groups. When $Y_i(t_i=1)$ is
    censored rather than observed, we replace it with a simulation from
    the model given available knowledge of the censoring process.
    Variation in the simulations is due to two factors: uncertainty in
    the imputation process for censored $y_i^*$ and uncertainty in
    simulating $E[Y_i(t_i=0)]$, the counterfactual expected value of
    $Y_i$ for observations in the treatment group, under the assumption
    that everything stays the same except that the treatment indicator
    is switched to $t_i=0$.

-   In conditional prediction models, the average predicted treatment
    effect (`att.pr`) for the treatment group is

    $$\frac{1}{\sum_{i=1}^n t_i}\sum_{i:t_i=1}^n \left\{ Y_i(t_i=1) -
      \widehat{Y_i(t_i=0)} \right\},$$

    where $t_i$ is a binary explanatory variable defining the treatment
    ($t_i=1$) and control ($t_i=0$) groups. When $Y_i(t_i=1)$ is
    censored rather than observed, we replace it with a simulation from
    the model given available knowledge of the censoring process.
    Variation in the simulations is due to two factors: uncertainty in
    the imputation process for censored $y_i^*$ and uncertainty in
    simulating $\widehat{Y_i(t_i=0)}$, the counterfactual predicted
    value of $Y_i$ for observations in the treatment group, under the
    assumption that everything stays the same except that the treatment
    indicator is switched to $t_i=0$.

## Output Values

The output of each Zelig command contains useful information which you
may view. For example, if you run
`z.out <- zelig(Surv(Y, C) ~ X, model = exp, data)`, then you may
examine the available information in `z.out` by using `names(z.out)`,
see the `coefficients` by using
`z.out$coefficients`, and a default summary of information
through `summary(z.out)`. Other elements available through the
`$` operator are listed below.

-   From the `zelig()` output object `z.out`, you
    may extract:

    -   `coefficients`: parameter estimates for the
        explanatory variables.

    -   `icoef`: parameter estimates for the intercept and
        scale parameter. While the scale parameter varies for the
        Weibull distribution, it is fixed to 1 for the exponential
        distribution (which is modeled as a special case of the
        Weibull).

    -   `var`: the variance-covariance matrix for the
        estimates of $\beta$.

    -   `loglik`: a vector containing the log-likelihood for
        the model and intercept only (respectively).

    -   `linear.predictors`: the vector of $x_{i}\beta$.

    -   `df.residual`: the residual degrees of freedom.

    -   `df.null`: the residual degrees of freedom for the
        null model.

    -   `zelig.data`: the input data frame if `save.data
        = TRUE`.

-   Most of this may be conveniently summarized using
    `summary(z.out)`. From `summary(z.out)`, you
    may additionally extract:

    -   `table`: the parameter estimates with their
        associated standard errors, $p$-values, and $t$-statistics. For
        example, `summary(z.out)$table`

-   From the `sim()` output stored in `s.out`:

-   From the `sim()` output object `s.out`, you
    may extract quantities of interest arranged as matrices indexed by
    simulation $\times$ `x`-observation (for more than one
    `x`-observation). Available quantities are:

    -   `qi$ev`: the simulated expected values for the
        specified values of `x`.

    -   `qi$pr`: the simulated predicted values drawn from a
        distribution defined by the expected values.

    -   `qi$fd`: the simulated first differences between the
        simulated expected values for `x` and
        `x1`.

    -   `qi$att.ev`: the simulated average expected
        treatment effect for the treated from conditional prediction
        models.

    -   `qi$att.pr`: the simulated average predicted
        treatment effect for the treated from conditional prediction
        models.

How to Cite {#how-to-cite .unnumbered}
-----------

See also {#see-also .unnumbered}
--------

The exponential function is part of the survival library by Terry
Therneau, ported to R by Thomas Lumley. Advanced users may wish to refer
to `help(survfit)` in the survival library and @VenRip02. Sample data are
from @KinAltBur90.
