---
title: "Comparing Zelig 5 and Zelig 4 Syntax"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Zelig 5 vs Zelig 4}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

*Built using Zelig version `r packageVersion('Zelig')`*

```{r setup, include=FALSE}
knitr::opts_knit$set(
        stop_on_error = 2L
)
knitr::opts_chunk$set(eval=FALSE)
```

The release of Zelig 5.0 expands the set of models available, while simplifying the model wrapping process, and solving architectural problems by completely rewriting into R’s Reference Classes for a fully object-oriented architecture. Comparability wrappers are available so that you can still use pre-Zelig 5 syntax.

Zelig 5 uses [reference classes](http://adv-r.had.co.nz/R5.html) which work a bit differently from what you may expect in R. The big difference is that they are "mutable", i.e. assigning values to them does not overwrite the objects previous contents.

## Zelig 5 workflow overview

All models in Zelig 5 can be estimated and results explored presented using a five simple steps:

- Initialise the Zelig object, e.g with `z.out <- zls$new()`, for a least squares model. Then populate the object with:

- `zelig` to estimate the parameters,

- `setx` to set fitted values for which we want to find quantities of interest,

- `sim` to simulate the quantities of interest,

- `graph` to plot the simulation results.

Here is an example workflow with Zelig 5:

```{r zelig-5-pseudo}
z5 <- zls$new()
z5$zelig(Y ~ X1 + X ~ X, weights = w, data = mydata)
z5$setx()
z5$sim()
z5$graph()
```

## Wrappers

Here is the same set of operations as above using the Zelig 4 wrappers:

```{r zelig-4-pseudo}
z.out <- zelig(Y ~ X1 + X2, model = "ls", weights = w, data = mydata)
x.out <- setx(z.out)
s.out <- sim(z.out, x = x.out)
plot(s.out)
```

Note that all of the output objects from the Zelig 4 wrappers are Zelig 5 reference class objects, so you can mix and match which syntax you like.

Also note that when using the Zelig 5 methods for a range of fitted values
you will need to use `setrange` rather than `setx`. For first differences you
will need to call `$setx1` after `$setx`. Because of this, using the
`setx` compatibility wrapper can be more straightforward.
